### 2023-04-12 example_paper.tex
Dump abstract from proposal; can't figure out paper.yml → header.tex
Don't know how to the information from the paper.yml file to automatically
populate the header.tex file like the docs say. [1] I just modified it manually
for kicks.

[1]: https://juliacon.github.io/proceedings-guide/author/

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}

\begin{document}

\input{header}

\maketitle

\begin{abstract}

This is a guide for authors who are preparing papers for JuliaCon using the \LaTeX{} document
preparation system and the \verb|juliacon|  class file.

\end{abstract}

\section{Introduction}

The \LaTeX{} document preparation system is a special version of the
\TeX; typesetting program where-in a collection of \TeX; commands
are added to \LaTeX{} to simplify typesetting. Importantly, it allows
the author to concentrate on the logical structure of the document
rather than its visual layout.\vskip 6pt
Moreover, \LaTeX{} provides a consistent and comprehensive document
preparation interface. There are simple-to-use commands for
generating a table of contents, lists of figures and/or tables, and indexes.
\LaTeX{} can automatically number list entries, equations, figures,
tables, and footnotes, as well as articles, sections, and subsections.
Using this numbering system, bibliographic citations, page
references, and cross references to any other numbered entity (e.g.
article, section, equation, figure, list entry, etc.) become quite simple
and straightforward. The use of \LaTeX{} document classes allows
a simple change of class to transform the appearance of your document.\vskip 6pt
\LaTeX{} is a powerful tool for managing long and complex documents.
In particular, partial processing enables long documents to
be produced article by article without losing sequential information.
The use of document classes allows a simple change of style
(or style option) to transform the appearance of your document.

\section{The JuliaCon Article Class}
\label{sec:documentclass}
%
The juliacon class file preserves the standard LATEX{} interface such
that any document that can be produced using the standard LATEX{}
article class can also be produced with the class file.\vskip 6pt
It is likely that the make up will change after file submission. For
this reason, we ask you to ignore details such as slightly long lines,
page stretching, or figures falling out of synchronization, as these
details can be dealt with at a later stage.\vskip 6pt
Use should be made of symbolic references (\verb|\ref| ) in order to
protect against late changes of order, etc.

\section{USING THE JuliaCon Article CLASS FILE}

If the file \verb|juliacon.cls|  is not already in the appropriate system directory
for \LaTeX{} files, either arrange for it to be put there or copy
it to your working directory. The \verb|juliacon|  document class is implemented
as a complete class, not a document style option. In order to
use the \verb|juliacon  document class, replace \verb|article|  by \verb|juliacon|  in the
\verb|\documentclass|  command at the beginning of your document:
\vskip 6pt
\begin{centering}
    \verb|\documentclass{article}|  \end{centering}
\vskip 6pt
replace by
\vskip 6pt
 \verb|\documentclass{juliacon}|  \vskip 6pt
In general, the following standard document \verb|style|  options should
{ \itshape not} be used with the {\footnotesize \itshape article} class file:
\begin{enumerate}
\item[(1)] \verb|10pt|,  \verb|11pt|,  \verb|12pt|   ? unavailable;
\item[(2)] \verb|twoside|  (no associated style file) ? \verb|twoside|  is the default;
\item[(3)] \verb|fleqn|, \verb|leqno|, \verb|titlepage| ? should not be used;
\end{enumerate}

\section{Additional Document Style Options}
\label{sec:additional_doc}
%
The following additional style option is available with the \verb|juliacon|  class file:
\vskip 6pt
Please place any additional command definitions at the very start of
the \LaTeX{} file, before the \verb|\begin{document}|. For example, user-defined
\verb|\def|  and \verb|\newcommand|   commands that define macros for
technical expressions should be placed here. Other author-defined
macros should be kept to a minimum.
\vskip 6pt
Commands that differ from the standard \LaTeX{} interface, or that
are provided in addition to the standard interface, are explained in
this guide. This guide is not a substitute for the \LaTeX{} manual itself.
Authors planning to submit their papers in \LaTeX{} are advised to use
\verb|juliacon.cls|  as early as possible in the creation of their files.

%
%
%
%
\begin{table*}[t]
\tabcolsep22pt
\tbl{If necessary, the tables can be extended both columns.}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Test 1 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test 2 & Testing Data I & & 105\\\hline
Test 3 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\label{tab:symbols}
\begin{tabnote}
This is an example of table footnote.
\end{tabnote}
\end{table*}
% \begin{figure*}[t]
% \centerline{\includegraphics[width=11cm]{juliagraphs.png}}
% \caption{If necessary, the images can be extended both columns.}
%   \label{fig:sample_image}
% \end{figure*}

\section{Additional features}
\label{sec:additional_faci}
In addition to all the standard \LaTeX{} design elements, the \verb|juliacon|  class file includes the following features:
In general, once you have used the additional \verb|juliacon.cls| facilities
in your document, do not process it with a standard \LaTeX{} class
file.

\subsection{Titles, Author's Name, and Affiliation}
\label{subsub:title_auth}
The title of the article, author's name, and affiliation are used at the
beginning of the article (for the main title). These can be produced
using the following code:

\begin{verbatim}
\title{ This is an example of article title} }
\author{
   \large 1st Author \\[-3pt]
   \normalsize 1st author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	1st author's email address \\[-3pt]
  \and
   \large 2nd Author \\[-3pt]
   \normalsize 2nd author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	2nd author's email address \\[-3pt]
\and
   \large 3rd Author \\[-3pt]
   \normalsize 3rd author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	3rd author's email address \\[-3pt]
}
\maketitle
\end{verbatim}

\subsection{Writing Julia code}

A special environment is already defined for Julia code,
built on top of \textit{listings} and \textit{jlcode}.

\begin{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}
\end{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}

\subsection{Abstracts, Key words, term etc...}
\label{subsub:abs_key_etc}

At the beginning of your article, the title should be generated
in the usual way using the \verb|\maketitle|  command. For genaral tem and keywords use
\verb|\terms|,
\verb|\keywords|  commands respectively. The abstract should be enclosed
within an abstract environment, All these environment
can be produced using the following code:
\begin{verbatim}
\terms{Experimentation, Human Factors}

\keywords{Face animation, image-based modelling...}

\begin{abstract}
In this paper, we propose a new method for the
systematic determination of the model's base of
time varying delay system. This method based on
the construction of the classification data related
to the considered system. The number, the orders,
the time delay and the parameters of the local
models are generated automatically without any
knowledge about the full operating range of the
process. The parametric identification of the local
models is realized by a new recursive algorithm for
on line identification of systems with unknown time
delay. The proposed algorithm allows simultaneous
estimation of time delay and parameters of
discrete-time systems. The effectiveness of
the new method has been illustrated through
simulation.
\end{abstract}

\end{verbatim}

\section{Some guidelines}
\label{sec:some_guide}
The following notes may help you achieve the best effects with the
\verb|juliacon|  class file.

\subsection{Sections}
\label{subsub:sections}
\LaTeXe{}  provides four levels of section headings and they are all
defined in the \verb|juliacon|  class file:
\begin{itemize}
\item \verb|\section|
\item \verb|\subsection|
\item \verb|\subsubsection|
\item \verb|\paragraph|
\end{itemize}
Section headings are automatically converted to allcaps style.
\subsection{Lists}
\label{sec:lists}
%
The \verb|juliacon|   class file provides unnumbered lists using the
\verb|unnumlist|  environment for example,

\begin{unnumlist}
\item First unnumbered item which has no label and is indented from the
left margin.
\item Second unnumbered item.
\item Third unnumbered item.
\end{unnumlist}
The unnumbered list which has no label and is indented from the
left margin. was produced by:
\begin{verbatim}
\begin{unnumlist}
\item First unnumbered item...
\item Second unnumbered item...
\item Third unnumbered item...
\end{unnumlist}
\end{verbatim}

The \verb|juliacon|  class file also provides hyphen list using the
\verb|itemize|  environment for example,
\begin{itemize}
\item First unnumbered bulleted item which has no label and is indented
from the left margin.
\item Second unnumbered bulleted item.
\item Third unnumbered bulleted item which has no label and is indented
from the left margin.
\end{itemize}
was produced by:
\begin{verbatim}
\begin{itemize}
\item First item...
\item Second item...
\item Third item...
\end{itemize}
\end{verbatim}

Numbered list is also provided in acmtog class file using the
enumerate environment for example,
\begin{enumerate}
\item The attenuated and diluted stellar radiation.
\item Scattered radiation, and
\item Reradiation from other grains.
\end{enumerate}

was produced by:
\begin{verbatim}
\begin{enumerate}
\item The attenuated...
\item Scattered radiation, and...
\item Reradiation from other grains...
\end{enumerate}
\end{verbatim}
\subsection{Illustrations (or figures)}
\label{subsub:sec_Illus}
The \verb|juliacon|  class file will cope with most of the positioning of
your illustrations and you should not normally use the optional positional
qualifiers on the \verb|figure|  environment that would override
these decisions.
\vskip 6pt

%
\begin{figure}[t]
\centerline{\includegraphics[width=4cm]{juliagraphs.png}}
\caption{This is example of the image in a column.}
	\label{fig:sample_figure}
\end{figure}

The figure \ref{fig:sample_figure} is taken from the JuliaGraphs
organization \footnote{https://github.com/JuliaGraphs}.

Figure captions should be \emph{below} the figure itself, therefore the
\verb|\caption|  command should appear after the figure or space left for
an illustration. For example, Figure 1 is produced using the following
commands:

\begin{verbatim}
\begin{figure}
\centerline{\includegraphics[width=20pc]{Graphics.eps}}
\caption{An example of the testing process for a
binary tree. The globa null hypothesis is tested
first at level $\alpha$ (a), and the level of
individual variables is reached last (d). Note
that individual hypotheses can be tested at
level $\alpha/4$ and not $\alpha/8$ as one might
expect at first.}
\label{sample-figure_2}
\end{figure}
\end{verbatim}
Figures can be resized using first and second argument of
\verb|\includegraphics|   command. First argument is used for modifying
figure height and the second argument is used for modifying
figure width respectively.
\vskip 6pt
Cross-referencing of figures, tables, and numbered, displayed
equations using the \verb|\label|  and \verb|\ref|   commands is encouraged.
For example, in referencing Figure 1 above, we used
\verb|Figure~\ref{sample-figure}|

 \subsection{Tables}
\label{subsub:sec_Tab}
The \verb|juliacon|   class file will cope with most of the positioning of
your tables and you should not normally use the optional positional qualifiers on the table environment which would override these
decisions. Table captions should be at the top.
\begin{verbatim}
\begin{table}
\tbl{Tuning Set and Testing Set}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Train70 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test70 & Testing Data I & & 105\\\hline
Test30 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\end{table}
\end{verbatim}

\begin{table}
\tbl{Tuning Set and Testing Set}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Test 1 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test 2 & Testing Data I & & 105\\\hline
Test 3 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\end{table}
\subsection{Landscaping Pages}
\label{subsub:landscaping_pages}
If a table is too wide to fit the standard measure, it may be turned,
with its caption, to 90 degrees. Landscape tables cannot be produced
directly using the \verb|juliacon|   class file because \TeX{} itself cannot
turn the page, and not all device drivers provide such a facility.
The following procedure can be used to produce such pages.
\vskip 6pt
Use the package \verb|rotating|   in your document and change the coding
from
\begin{verbatim}
\begin{table}...\end{table}
to
\begin{sidewaystable}...\end{sidewaystable}
and for figures
\begin{figure}...\end{figure}
to
\begin{sidewaysfigure}...\end{sidewaysfigure}
\end{verbatim}

environments in your document to turn your table on the appropriate
page of your document. For instance, the following code prints
a page with the running head, a message half way down and the
table number towards the bottom.
\begin{verbatim}
\begin{sidewaystable}
\tbl{Landscape table caption to go here.}{...}
\label{landtab}
\end{sidewaystable}
\end{verbatim}

\subsection{Double Column Figure and Tables}
\label{subsub:double_fig_tab}
For generating the output of figures and tables in double column
we can use the following coding:

\begin{enumerate}
\item For Figures:
\begin{verbatim}
\begin{figure*}...\end{figure*}
\end{verbatim}
\item For landscape figures:
\begin{verbatim}
\begin{sidewaysfigure*}...\end{sidewaysfigure*}
\end{verbatim}
\item For Tables:
\begin{verbatim}
\begin{table*}...\end{table*}
\end{verbatim}
\item For landscape tables:
\begin{verbatim}
\begin{sidewaystable*}...\end{sidewaystable*}
\end{verbatim}
\end{enumerate}

\subsection{Typesetting Mathematics}
\label{subsub:type_math}
The \verb|juliacon| class file will set displayed mathematics with center to
the column width, provided that you use the \LaTeXe{} standard of
open and closed square brackets as delimiters.
The equation
\[
\sum_{i=1}^p \lambda_i = (S)
\]

was typeset using the acmtog class file with the commands

\begin{verbatim}
\[
\sum_{i=1}^p \lambda_i = (S)
\]
\end{verbatim}

For display equations, cross-referencing is encouraged. For example,
\begin{verbatim}
\begin{equation}
(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
\label{eq:samplevar}
\end{equation}
Equation~(\ref{eq:samplevar}) gives the formula for
sample variance.
\end{verbatim}
The following output is generated with the above coding:
\begin{equation}
(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
\label{eq:samplevar}
\end{equation}
Equation~(\ref{eq:samplevar}) gives the formula for
sample variance.

\subsection{Enunciations}
\label{subsub:enunciation}
The \verb|juliacon|   class file generates the enunciations with the help of
the following commands:
\begin{verbatim}
\begin{theorem}...\end{theorem}
\begin{strategy}...\end{strategy}
\begin{property}...\end{property}
\begin{proposition}...\end{proposition}
\begin{lemma}...\end{lemma}
\begin{example}...\end{example}
\begin{proof}...\end{proof}
\begin{definition}...\end{definition}
\begin{algorithm}...\end{algorithm}
\begin{remark}...\end{remark}
\end{verbatim}
The above-mentioned coding can also include optional arguments
such as
\begin{verbatim}
\begin{theorem}[...]. Example for theorem:
\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven ... created equal.
\end{theorem}
\end{verbatim}

\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven years ago our fathers brought forth,
upon this continent, a new nation, conceived in Liberty,
 and dedicated to the proposition that all men are
created equal.
\end{theorem}

\subsection{Extract}
\label{subsub:extract}
Extract environment should be coded within
\begin{verbatim}
\begin{extract}..\end{extract}
\end{verbatim}

\subsection{Balancing column at last page}
\label{subsub:Balance}
For balancing the both column length at last page use :
\begin{verbatim}
\vadjust{\vfill\pagebreak}
\end{verbatim}

%\vadjust{\vfill\pagebreak}

at appropriate place in your \TeX{} file or in bibliography file.

\section{Handling references}
\label{subsub:references}
References are most easily (and correctly) generated using the
BIBTEX, which is easily invoked via
\begin{verbatim}
\bibliographystyle{juliacon}
\bibliography{ref}
\end{verbatim}
When submitting the document source (.tex) file to external
parties, the ref.bib file should be sent with it.
\cite{bezanson2017julia}

\input{bib.tex}

\end{document}

% Inspired by the International Journal of Computer Applications template

### 2023-04-12 header.tex
Dump abstract from proposal; can't figure out paper.yml → header.tex
Don't know how to the information from the paper.yml file to automatically
populate the header.tex file like the docs say. [1] I just modified it manually
for kicks.

[1]: https://juliacon.github.io/proceedings-guide/author/

% **************GENERATED FILE, DO NOT EDIT**************

\title{My JuliaCon proceeding}

\author[1]{Taylor Allred}
\author[1]{Xinyi Li}
\author[1]{Ashton Wiersdorf}
\author[1]{Ben Greenman}
\author[1]{Ganesh Gopalakrishnan}
\affil[1]{University of Utah}

\keywords{Julia, Optimization, Game theory, Compiler}

\hypersetup{
pdftitle = {My JuliaCon proceeding},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {1st author, 2nd author, 3rd author},
pdfkeywords = {Julia, Optimization, Game theory, Compiler},
}

### 2023-04-18 paper.tex
Basic outline; really rough draft on FloatTracker and ODE case study

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}

\begin{document}

\input{header}
\maketitle

\begin{abstract}
  Reliable numerical computations are central to HPC and ML.
  We present FlowFPX: a Julia-based tool for tracking the onset and flow of IEEE Floating-Point exceptions that signal numerical defects.
  FlowFPX's design exploits Julia's operator overloading to trace exception flows and even inject exceptions to accelerate testing.
  We present intuitive visualizations of summarized exception flows including how they are generated, propagated and killed, thus helping with debugging and repair.
\end{abstract}

\section{Introduction}

\section{To Kill a Floating Point}
% If anyone has a better "kill" pun/reference than "To Kill a Mockingbird", I'm all ears.

% I think we will definitely need a clear, detailed section on how NaNs crop up and how they can be killed.

\section{FloatTracker Internals}

FloatTracker works by leveraging Julia's type-based method dispatch system to capture everything from calls to the standard library down to basic arithmetic operations.
We do this by creating custom types \texttt{TrackedFloat16}, \texttt{TrackedFloat32}, and \texttt{TrackedFloat64} that wrap their corresponding \texttt{Float16}, \texttt{Float32} etc. counterparts.
We then defined methods for every function in the \texttt{Base} module, including new implementations for operators like \texttt{+} and \texttt{*}.
These new methods had to be defined for every combination of tracked/untracked arguments;
fortunately Julia's metaprogramming capabilities make this relatively simple.

The surrogate methods in place, every operation on a TrackedFloat can be intercepted.
We have three primary goals when intercepting a float:

\begin{enumerate}
  \item Monitor the propagation of NaNs through the program
  \item Catch instances of NaN kills
  \item Optionally \emph{inject} NaNs to fuzz the program under scrutiny against NaN-kills
\end{enumerate}

% Good things to put in this section:
% - start with the max example, perhaps?
% - trace examples
% - CSTGs

\section{Case Studies}

To illustrate what FloatTracker is capable of, we applied our tool to three different Julia libraries to help us track down strange behavior:

\begin{description}
\item[Surprise NaNs] We found that otherwise innocuous values for a time step parameter in the \texttt{ShallowWaters} library can produce NaNs in the result.
  We used FloatTracker to track down where the NaNs arose.
\item[NaN Fuzzing] We injected NaNs in a N-Body simulation and were able to find a bug in the widely-used \texttt{OrdinaryDiffEq} library.
\item[NaN Kills] We observed issues with NaN kills inside of the \texttt{Finch} library.
  % FIXME: more needed here
\end{description}

All three libraries are available from the JuliaHub package archive.
The \texttt{OrdinaryDiffEq} library in particular enjoys widespread use in the scientific computing community.

\subsection{Surprise NaNs: ShallowWaters}

\subsection{NaN Fuzzing: OrdinaryDiffEq}

The second thing we looked at evaluating was an N-Body simulator that we found.
We wanted to test the fuzzing abilities of FloatTracker:
we configured FloatTracker so that it would inject one NaN at random into the program as it ran, and also to record the injection so that we could reproduce any interesting runs later.
During one run, the N-Body simulation gave us a warning about a NaN causing instability, and then a notice that the program would exit.
However, the program went into an infinite loop.

With the recording in hand, we were able to reliably reproduce this behavior.
Even though the loop only manifested on certain runs of the simulator, the recording made the NaN-injection deterministic.
We used both the stack trace from terminating the loop as well as the injection and kill logs provided by FloatTracker to trace where the loop occurred and how the NaN we injected caused the problem.

% Note for Ashton:
% file here: ~/Research/ode_debug/dev/OrdinaryDiffEq/src/solve.jl
% see line 514

The logs lead us to a routine in the widely-used \texttt{OrdinaryDiffEq} library.
During the initialization of a differential equation problem, a NaN in the wrong place causes the setup to loop forever.
Indecently, this is where we see a NaN kill; the critical code in question looks something like this:

\begin{lstlisting}[language = Julia]
while !isempty(time_stops)
  while tdir * t < first(time_stops)
    # do integration work
  end
  pop_if_work_done(time_stops)
end
\end{lstlisting}

The problem here is on line 2 when \texttt{tdir} is NaN: multiplication propagates NaNs, and comparison kills it, so the condition on the inner \texttt{while} loop is \emph{always} false.
No work got done in the inner loop, and so the conditional \texttt{pop} routine never reduced the size of the \texttt{time\_stops} vector.

Our kill logs led us right to this line; this is a perfect example of how NaN kills can influence control flow to go awry.
In this case, the problem was apparent besides what our logs told us and it manifested as an infinite loop.
More dangerous cases can occur when the influence of a bad comparison is not as readily observable.

\subsection{NaN Kills: Finch}

% Advection examples go here, right?

\section{Evaluation}

\section{Discussion}

\section{Related Work}

\section{Acknowledgments}

\input{bib.tex}

\end{document}

% Below is some Emacs-specific stuff. If you're using Emacs, you should try
% using the new jinx package for spell-checking!

% Local Variables:
% jinx-local-words: "CSTGs JuliaHub OrdinaryDiffEq ShallowWaters TrackedFloat"
% End:

### 2023-04-20 paper.tex
Rework & expand ODE case study: add injection and kill log examples

@@
 % - CSTGs
+%
+% Don't forget to talk about trace recording
 
@@
 
-The second thing we looked at evaluating was an N-Body simulator that we found.
-We wanted to test the fuzzing abilities of FloatTracker:
-we configured FloatTracker so that it would inject one NaN at random into the program as it ran, and also to record the injection so that we could reproduce any interesting runs later.
-During one run, the N-Body simulation gave us a warning about a NaN causing instability, and then a notice that the program would exit.
-However, the program went into an infinite loop.
-
-With the recording in hand, we were able to reliably reproduce this behavior.
-Even though the loop only manifested on certain runs of the simulator, the recording made the NaN-injection deterministic.
-We used both the stack trace from terminating the loop as well as the injection and kill logs provided by FloatTracker to trace where the loop occurred and how the NaN we injected caused the problem.
+The second thing we evaluated was an N-Body simulation library.
+We configured FloatTracker to inject a single NaN to see if we could find any NaN kills.
+On some runs we noticed that we would get a kill and the program would warn about a NaN and prepare to exit.
+Contrary to the error message's claim, however, the program went into an infinite loop.
 
@@
 The logs lead us to a routine in the widely-used \texttt{OrdinaryDiffEq} library.
-During the initialization of a differential equation problem, a NaN in the wrong place causes the setup to loop forever.
-Indecently, this is where we see a NaN kill; the critical code in question looks something like this:
+During initialization, a NaN got injected during the execution of \texttt{-} between two \texttt{TrackedFloat}s:
+
+\begin{verbatim}
+…
+- at FloatTracker/src/TrackedFloat.jl:89
+#__init#628 at OrdinaryDiffEq/src/solve.jl:106
+…
+\end{verbatim}
+
+That injection point is here on third line of this snippet, which corresponds to line 106 from the \texttt{solve.jl} file in the \texttt{OrdinaryDiffEq} library:
+
+\begin{lstlisting}[language = Julia]
+tType = eltype(prob.tspan)
+tspan = prob.tspan
+tdir = sign(tspan[end] - tspan[1])
+
+t = tspan[1]
+\end{lstlisting}
+
+\texttt{sign} correctly propagates NaNs, so \texttt{tdir} now contains a NaN.
+
+We \emph{could} trace this variable through the code, but that would be a lot of grunt work.
+Instead, we can look at the NaN kill logs to get a clue as to where we should look next:
+
+\begin{verbatim}
+…
+< at FloatTracker/src/TrackedFloat.jl:193
+solve! at OrdinaryDiffEq/src/solve.jl:515
+…
+\end{verbatim}
+
+We get a very large kill file from this run, and the two lines from above show up repeatedly.
+Thus, this is a good candidate location for where the cause of the loop is.
+
+The relevant part of \texttt{solve.jl} looks like this:
 

### 2023-04-21 paper.tex
Add a sentence more
FILL: TF items need a subsection each, eventually
Write a bit about killing floating point NaNs

@@
 \setcounter{page}{1}
+\usepackage{amsmath}
 
@@
 
+The IEEE 754 specification for floating-point numbers \cite{IEEEStandardBinary1985} governs how NaNs are handled as arguments to arithmetic functions.
+Most of the time, when any or all of the arguments are NaN, the result is also a NaN.
+This is good and desirable, because it means that an arithmetic error---even if buried deep in an expression---will still be observable, allowing a user to find and debug it.
+
+However, there are a few places where NaNs flow into an expression, but not out.
+We call these instances ``NaN-kills'', and these can lead to subtle logic bugs.
+Consider the following example:
+
+\begin{lstlisting}[language = Julia]
+function bad_max(lst)
+  max_seen = 0.0
+  for x in lst
+    if ! (x <= max_seen)
+      # swap if new val greater
+      max_seen = x
+    end
+  end
+  max_seen
+end
+
+function good_max(lst)
+  foldl(max, lst)
+end
+\end{lstlisting}
+
+Both \texttt{bad\_max} and \texttt{good\_max} compute the maximum in a list.
+However, the \texttt{<=} operator used in the \texttt{bad\_max} implementation does \emph{not} propagate NaNs, so you might not get the result you expected:
+
+\begin{lstlisting}[language = Julia]
+bad_max([1, 5, NaN, 4])     # returns 4
+good_max([1, 5, NaN, 4])    # returns NaN
+\end{lstlisting}
+
+Not only is the result from \texttt{bad\_max} problematic for obscuring the fact that there was a NaN in the list, the result is also wrong!
+The built-in \texttt{max} function correctly propagates NaNs, so we can see from its return that we got a NaN somewhere.
+It's not entirely obvious \emph{where} that NaN came from---but more on that later.   % this promise answered in "NaN propagation monitoring"
+
+\subsection{Where do NaNs come from?}
+
+IEEE 754 \cite{IEEEStandardBinary1985} lists the following operations (among a few others) where NaNs result:
+
+\begin{itemize}
+\item $\frac{0}{0}$ or $\frac{\infty}{\infty}$
+\item $0 * \infty$
+\item $x \% y$ when $x = \infty$ or $y = 0$
+\item $+\infty + -\infty$ or $\infty - \infty$
+\item The square root or logarithm of a negative number
+\end{itemize}
+
+% Yadda yadda yadda...
+
+\subsection{NaN kills}
+
+Most mathematical operations preserve NaNs.
+However, there are some instances where NaNs can disappear, but not before potentially altering the correctness of the program!
+
+% \begin{itemize}
+% \item $x \; \texttt{cmp} \; y \rightarrow false$ where \texttt{cmp} is a binary comparison like \texttt{<}, \texttt{>}, \texttt{<=}, etc. and $x$ or $y$ is NaN.
+% \item $1^{NaN} \rightarrow 1$
+% \item $NaN^{0} \rightarrow 1$
+% \end{itemize}
+
+\begin{align*}
+x \; \texttt{cmp} \; y &\rightarrow false &\text{where \texttt{cmp} is \texttt{<}, \texttt{>}, \texttt{<=}, etc.} \\
+1^{\text{\texttt{NaN}}} &\rightarrow 1 \\
+\text{\texttt{NaN}}^{0} &\rightarrow 1
+\end{align*}
+
+% Yadda yadda yadda...
+
+% What else do we need to talk about?
+
 \section{FloatTracker Internals}
@@
 \begin{enumerate}
+  \item Catch instances of NaN kills
   \item Monitor the propagation of NaNs through the program
-  \item Catch instances of NaN kills
   \item Optionally \emph{inject} NaNs to fuzz the program under scrutiny against NaN-kills
@@
 % Don't forget to talk about trace recording
+
+\subsection{Catching NaN kills}
+
+\subsection{NaN propagation monitoring}
+
+Remember in our first example with the \texttt{bad\_max} and \texttt{good\_max} functions how we didn't have a way to tell \emph{where} the NaNs originally came from?
+That can be a tedious task.
+However, FloatTracker makes this a lot easier by watching the life cycle of a NaN.
+
+% FIXME: talk about CSTGs
+
+\subsection{NaN injection}
 
@@
 The second thing we evaluated was an N-Body simulation library.
+We didn't find any sensible configuration of input parameters that lead to NaN kills like in the ShallowWaters library.
+We turned to the NaN injection capabilities of FloatTracker to fuzz the library to find any lurking bugs.
+% FIXME: add hyper-ref to § 4.1: ShallowWaters
+
 We configured FloatTracker to inject a single NaN to see if we could find any NaN kills.
@@
 Contrary to the error message's claim, however, the program went into an infinite loop.
-
-% Note for Ashton:
-% file here: ~/Research/ode_debug/dev/OrdinaryDiffEq/src/solve.jl
-% see line 514
 
@@
 The relevant part of \texttt{solve.jl} looks like this:
+
+% Note for Ashton:
+% file here: ~/Research/ode_debug/dev/OrdinaryDiffEq/src/solve.jl
+% see line 514
 

### 2023-05-12 paper.tex
Add note
Update bibliography; begin draft of ShallowWaters case study

@@
 \begin{abstract}
-  Reliable numerical computations are central to HPC and ML.
+  Reliable numerical computations are central to high-performance computing and machine learning.
   We present FlowFPX: a Julia-based tool for tracking the onset and flow of IEEE Floating-Point exceptions that signal numerical defects.
@@
 % If anyone has a better "kill" pun/reference than "To Kill a Mockingbird", I'm all ears.
-
 % I think we will definitely need a clear, detailed section on how NaNs crop up and how they can be killed.
@@
 
-% What else do we need to talk about?
+% What else do we need to talk about? What other problems do we need to set up so that FloatTracker's solutions are clear?
 
@@
 
-\subsection{NaN Fuzzing: OrdinaryDiffEq}
+Most physics simulations of this sort include a parameter called the Courant-Friedrichs-Lewy (``CFL'') parameter, which roughly describes the size of the time step to take in running the simulation.
+Normally this value sits between 0 and 1---smaller values mean the simulation runs slower but with higher accuracy, whereas higher values mean the simulation takes bigger time steps and consequently runs faster, but results might be inaccurate.
+The primary reason for problems arising comes from information not having enough time to propagate through the simulation before the next time step is taken.
+
+% This is the story that I think we want to tell: you're going to run into NaNs,
+% so instead of dialing the the value down, we should look for where NaNs first
+% start to appear and fix those places.
+
+It is not uncommon to encounter cases where setting the CFL parameter too high results in NaNs in the result.
+Driving the CFL parameter low enough to excise the NaNs can make the simulation too long and unwieldy to be comfortable to use.
+Identifying where NaNs begin to arise can help point the way to where some mathematical refinement can be used to make the simulation both fast and accurate enough for practical use.
+
+For our first case study, we looked at the \texttt{ShallowWaters} library,\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019,klowerLowprecisionClimateComputing2021} which runs a simulates the flow of water over a seabed.
+This is not a general shallow waters simulation library---it is designed primarily to showcase numerical methods involving 16-bit floats.
+That said, it works for our purposes as a proving ground for FloatTracker.
+We dialed the CFL parameter up to high values (as high as 10) and we started seeing NaNs in the result.
+
+\textbf{FIXME: add pictures here}
+
+The holes in the plots are regions where NaNs crept into the computation and destroyed the results.
+In one regard, this is not as bad as a NaN kill---knowing that there are NaNs in the computation means that we have a chance to fix it.
+NaN kills might silently give us the wrong answer, but we know that there were no NaN kills because our kill logs are empty.
+That said, finding the source of these NaNs can be difficult.
+This is where we employed FloatTracker.
+
+We ran ShallowWaters with a CFL value known to make bad outputs and recorded NaN \texttt{gen} events:
+
+\begin{lstlisting}[language = Julia]
+set_logger(buffersize=20, cstg=true, cstgArgs=true, cstgLineNum=true)
+
+P = run_model(T=TrackedFloat32,
+              cfl=10, Ndays=100, nx=100, L_ratio=1,
+              bc="nonperiodic", wind_forcing_x="double_gyre",
+              topography="seamount")
+\end{lstlisting}
+
+This produced the following in the log file: (formatted for clarity)
+
+\begin{verbatim}
+…
+-                 at FloatTracker/src/TrackedFloat.jl:106
+momentum_u!       at ShallowWaters/src/rhs.jl:246
+rhs_nonlinear!    at ShallowWaters/src/rhs.jl:50
+rhs!              at ShallowWaters/src/rhs.jl:14 [inlined]
+time_integration  at ShallowWaters/src/time_integration.jl:77
+run_model         at ShallowWaters/src/run_model.jl:37
+…
+\end{verbatim}
+
+% WORKING HERE
+
+\subsection{NaN fuzzing to find a kill: OrdinaryDiffEq}
 
@@
 % Local Variables:
-% jinx-local-words: "CSTGs JuliaHub OrdinaryDiffEq ShallowWaters TrackedFloat"
+% jinx-local-words: "CFL CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq ShallowWaters TrackedFloat"
 % End:

### 2023-05-15 paper.tex
Some formatting fixes; more about ShallowWaters and catching kills

@@
 
+\begin{figure}[h]
+  \label{fig:nan-gens}
   \begin{itemize}
@@
   \end{itemize}
+  \caption{NaN-producing operations}
+\end{figure}
 
@@
 
+With a floating point value wrapped in our \texttt{TrackedFloat} type, it is easy to monitor when a NaN gets killed.
+Julia dispatches all arithmetic operations on that value to flow through FloatTracker, where we can inspect the arguments.
+Kills are relatively simple: if one or more NaNs flow into a function but don't flow out (i.e. the answer isn't a NaN) then that's a kill.
+For example:
+
+\begin{lstlisting}[language = Julia]
+  NaN + 42     -> NaN    # not a kill
+  Inf - Inf    -> NaN    # gen, but not a kill
+  min(NaN, 42) -> NaN    # not a kill
+  NaN < 42     -> false  # NaN kill!
+\end{lstlisting}
+
+There are a few exceptions to the rule of ``NaN in and no NaN out → kill''---some instances that we report as being a NaN kill might be instead a \emph{NaN-repair}, i.e.
+where the user explicitly checks for a NaN and gets some non-NaN value as a result.
+Once clear instance of this is the \texttt{isnan} function.
+Calling \texttt{isnan} and getting a boolean is expected behavior, and is usually used in error handling or in \emph{explicit} handling of unexpected values.
+FloatTracker still reports these cases in the kill logs so as to be complete---the user must not blindly assume that every instance of a NaN kill is bad and might be a case of NaN repair.
+
+% FIXME: we don't (yet) talk much about NaN repair in the rest of the paper.
+
 \subsection{NaN propagation monitoring}
@@
 
-% FIXME: talk about CSTGs
+% FIXME: don't forget to talk about CSTGs
 
@@
 That said, finding the source of these NaNs can be difficult.
-This is where we employed FloatTracker.
+
+Fortunately in this case, one of the key features \texttt{ShallowWaters} is its \emph{type flexibility}:
+\texttt{ShallowWaters} was designed to showcase physics simulations using variable levels of floating-point precision.
+Upon model construction, \texttt{ShallowWaters} allows you to specify a type for floats in the model.
+We simply asked it to use our \texttt{TrackedFloat32} type, and then the entire simulation ran under the supervision of FloatTracker.
 
@@
 \begin{lstlisting}[language = Julia]
-set_logger(buffersize=20, cstg=true, cstgArgs=true, cstgLineNum=true)
+set_logger(buffersize=20, cstg=true, cstgArgs=true,
+           cstgLineNum=true)
 
@@
               cfl=10, Ndays=100, nx=100, L_ratio=1,
-              bc="nonperiodic", wind_forcing_x="double_gyre",
+              bc="nonperiodic",
+              wind_forcing_x="double_gyre",
               topography="seamount")
@@
 …
--                 at FloatTracker/src/TrackedFloat.jl:106
-momentum_u!       at ShallowWaters/src/rhs.jl:246
-rhs_nonlinear!    at ShallowWaters/src/rhs.jl:50
-rhs!              at ShallowWaters/src/rhs.jl:14 [inlined]
-time_integration  at ShallowWaters/src/time_integration.jl:77
-run_model         at ShallowWaters/src/run_model.jl:37
+-                 FloatTracker/TrackedFloat.jl:106
+momentum_u!       ShallowWaters/rhs.jl:246
+rhs_nonlinear!    ShallowWaters/rhs.jl:50
+rhs!              ShallowWaters/rhs.jl:14 [inlined]
+time_integration  ShallowWaters/time_integration.jl:77
+run_model         ShallowWaters/run_model.jl:37
 …
@@
 
-% WORKING HERE
+We can see that the NaN appeared from subtraction (first line) and that it was used in the \texttt{momentum\_u!} function on line 246 of the \texttt{rhs.jl} file (second line).
+In figure~\ref{fig:nan-gens} we can tell that this must be the result of $\infty - \infty$.
+Where did the $\infty$ come from?
+
+% What do we say here?!
 
@@
 % Local Variables:
-% jinx-local-words: "CFL CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq ShallowWaters TrackedFloat"
+% jinx-local-words: "CFL CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq ShallowWaters TrackedFloat isnan"
 % End:

### 2023-05-25 paper.tex
Throw some pictures in

@@
 \usepackage{amsmath}
+
+% If using Emacs, put cursor on last closing paren and call "eval-last-sexp" (usually `C-x C-e') to regenerate
+\iffalse
+(save-excursion
+  (let* ((toc (funcall imenu-create-index-function))
+         (headers (mapcar (pcase-lambda (`(,str . ,loc)) str)
+                          toc)))
+    (beginning-of-buffer)
+    (search-forward (concat "%% " "--- autogenerated TOC ---"))
+    (next-line)
+    (beginning-of-line)
+    (let ((start (point)))
+      (search-forward (concat "%% " "--- end autogenerated TOC ---"))
+      (beginning-of-line)
+      (delete-region start (point))
+      (goto-char start))
+    (cl-loop for header in headers
+             do (insert (format "%% %s\n" header)))
+    (next-line)))
+\fi
+
+%% --- autogenerated TOC ---
+% juliacon
+% document
+%   Introduction
+%   To Kill a Floating Point
+%     Where do NaNs come from?
+%     NaN kills
+%   FloatTracker Internals
+%     Catching NaN kills
+%     NaN propagation monitoring
+%     NaN injection
+%   Case Studies
+%     Surprise NaNs: ShallowWaters
+%     NaN fuzzing to find a kill: OrdinaryDiffEq
+%     NaN Kills: Finch
+%   Evaluation
+%   Discussion
+%   Related Work
+%   Acknowledgments
+% document
+%% --- end autogenerated TOC ---
 
@@
 
+% TODO?
+
 \subsection{NaN propagation monitoring}
@@
 
+% TODO
+
 % FIXME: don't forget to talk about CSTGs
@@
 \subsection{NaN injection}
+
+% TODO
 
@@
 
-% This is the story that I think we want to tell: you're going to run into NaNs,
-% so instead of dialing the the value down, we should look for where NaNs first
-% start to appear and fix those places.
-
 It is not uncommon to encounter cases where setting the CFL parameter too high results in NaNs in the result.
@@
 We dialed the CFL parameter up to high values (as high as 10) and we started seeing NaNs in the result.
+Note that the program did not crash: instead, it began producing bad output.
 
-\textbf{FIXME: add pictures here}
+\begin{figure}
+  \centering
+  \includegraphics[width=3in]{./fig/shallow_waters_cfl_diff.png}
+  \caption{Effect of high CFL parameters: on the left, normal output. On the right: NaNs cause white holes and gaps in the result.}
+  \label{fig:sw_nans}
+\end{figure}
 
 The holes in the plots are regions where NaNs crept into the computation and destroyed the results.
-In one regard, this is not as bad as a NaN kill---knowing that there are NaNs in the computation means that we have a chance to fix it.
-NaN kills might silently give us the wrong answer, but we know that there were no NaN kills because our kill logs are empty.
-That said, finding the source of these NaNs can be difficult.
+In one regard, this is not as bad as a NaN kill---knowing that there are NaNs in the computation means that we have a chance to fix it, whereas a NaN kill might silently give us the wrong answer.
+Incidentally, we know that there were no NaN kills in this test because our kill logs are empty.
+But finding the sources of NaNs can be tricky.
+We expected NaNs in this case because we set the CFL parameter up so high, but tracking down the sources of NaNs generally is a difficult and time-consuming problem.
 

### 2023-05-31 paper.tex
Minor notes on RxInfer

@@
 %     NaN Kills: Finch
+%     A Bug-Hunting Safari
 %   Evaluation
@@
 
+\subsection{A Bug-Hunting Safari}
+
+% Notes on RxInfer
+
+We searched GitHub for open issues in Julia libraries involving difficult-to-find NaNs.
+In our first batch of issues, we found the ``RxInfer.jl'' library with an issue\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}} revolving around NaN detection.
+We offered our tool for the package authors to use, and within a day the authors were able to track down the location of a NaN and fix the issue---all with minimal guidance.
+
 \section{Evaluation}
@@
 \section{Acknowledgments}
+
+We would also like to thank the developers of RxInfer.jl for providing us with feedback on how we could make our tools more useful.
 

### 2023-06-02 paper.tex
change unicode -> to tex

@@
 
-There are a few exceptions to the rule of ``NaN in and no NaN out → kill''---some instances that we report as being a NaN kill might be instead a \emph{NaN-repair}, i.e.
+There are a few exceptions to the rule of ``NaN in and no NaN out $\rightarrow$ kill''---some instances that we report as being a NaN kill might be instead a \emph{NaN-repair}, i.e.
 where the user explicitly checks for a NaN and gets some non-NaN value as a result.

### 2023-06-02 header.tex
add latexmkrc, rebuild header etc

@@
 
-\title{My JuliaCon proceeding}
+\title{FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions}
 
@@
 
-\keywords{Julia, Optimization, Game theory, Compiler}
+\keywords{Julia, Floating point, Numerical computing}
 
 \hypersetup{
-pdftitle = {My JuliaCon proceeding},
-pdfsubject = {JuliaCon 2019 Proceedings},
-pdfauthor = {1st author, 2nd author, 3rd author},
-pdfkeywords = {Julia, Optimization, Game theory, Compiler},
+pdftitle = {FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions},
+pdfsubject = {JuliaCon 2022 Proceedings},
+pdfauthor = {Taylor Allred, Xinyi Li, Ashton Wiersdorf, Ben Greenman, Ganesh Gopalakrishnan},
+pdfkeywords = {Julia, Floating point, Numerical computing},
 }

### 2023-06-02 paper.tex
cite

@@
 
-The IEEE 754 specification for floating-point numbers \cite{IEEEStandardBinary1985} governs how NaNs are handled as arguments to arithmetic functions.
+The IEEE 754 specification for floating-point numbers governs how NaNs are handled as arguments to arithmetic functions~\cite{IEEEStandardBinary1985}.
 Most of the time, when any or all of the arguments are NaN, the result is also a NaN.
@@
 
-IEEE 754 \cite{IEEEStandardBinary1985} lists the following operations (among a few others) where NaNs result:
+IEEE 754~\cite{IEEEStandardBinary1985} lists the following operations (among a few others) where NaNs result:
 
@@
 
-For our first case study, we looked at the \texttt{ShallowWaters} library,\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019,klowerLowprecisionClimateComputing2021} which runs a simulates the flow of water over a seabed.
+For our first case study, we looked at the \texttt{ShallowWaters} library~\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019,klowerLowprecisionClimateComputing2021}, which runs a simulates the flow of water over a seabed.
 This is not a general shallow waters simulation library---it is designed primarily to showcase numerical methods involving 16-bit floats.

### 2023-06-02 paper.tex
cleanup

@@
 \end{abstract}
+
+\newcommand{\code}[1]{\texttt{#1}}
+\newcommand{\FlowFPX}{FlowFPX}
 
@@
 We searched GitHub for open issues in Julia libraries involving difficult-to-find NaNs.
-In our first batch of issues, we found the ``RxInfer.jl'' library with an issue\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}} revolving around NaN detection.
+In our first batch of issues, we found the \code{RxInfer.jl} library with an issue revolving around NaN detection.\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}}
 We offered our tool for the package authors to use, and within a day the authors were able to track down the location of a NaN and fix the issue---all with minimal guidance.
@@
 
-We would also like to thank the developers of RxInfer.jl for providing us with feedback on how we could make our tools more useful.
+We would also like to thank the developers of \code{RxInfer.jl} for feedback on \FlowFPX{}.
 

### 2023-06-07 paper.tex
finch case study, slicing future work

@@
 \usepackage{amsmath}
+\usepackage{cleveref}
 
@@
 \newcommand{\FlowFPX}{FlowFPX}
+\newcommand{\FloatTracker}{FloatTracker}
+\newcommand{\FT}{\FloatTracker}
+\newcommand{\CSTG}{stack graph}
+\newcommand{\CPP}{\code{C++}}
+\newcommand{\Dendro}{\textsc{Dendro}}
+\newcommand{\urlaccess}[2]{\url{#1}}
 
@@
 
-To illustrate what FloatTracker is capable of, we applied our tool to three different Julia libraries to help us track down strange behavior:
-
-\begin{description}
-\item[Surprise NaNs] We found that otherwise innocuous values for a time step parameter in the \texttt{ShallowWaters} library can produce NaNs in the result.
-  We used FloatTracker to track down where the NaNs arose.
-\item[NaN Fuzzing] We injected NaNs in a N-Body simulation and were able to find a bug in the widely-used \texttt{OrdinaryDiffEq} library.
-\item[NaN Kills] We observed issues with NaN kills inside of the \texttt{Finch} library.
-  % FIXME: more needed here
-\end{description}
-
-All three libraries are available from the JuliaHub package archive.
-The \texttt{OrdinaryDiffEq} library in particular enjoys widespread use in the scientific computing community.
+\FT{} is an effective tool for debugging all sorts of NaN-related issues.
+We have used it to reveal silent NaNs, uncover the root cause of NaNs,
+and fuzz for weak points in several case studies.
 
@@
 
-\subsection{NaN Kills: Finch}
+\subsection{Finch}
+\label{s:finch}
 
-% Advection examples go here, right?
+Finch is a domain-specific language for specifying
+PDEs~\cite{heislerFinchDomainSpecific2022}.\footnote{Not to be confused
+with the Finch loop optimizer~\cite{adka-cgo-2023}.}
+In the spirit of FEniCS~\cite{fenics} and related
+tools~\cite{freefem,openfoam,dune,firedrake},
+Finch helps scientists quickly convert math into code.
+What sets Finch apart is its flexibility.
+It supports multiple discretization methods (finite element and finite
+volume) and multiple backends (Julia, \CPP{}, \Dendro{}~\cite{dendro}).
+Furthermore it strives to output code that humans can easily fine-tune.
+
+Fuzzing with \FT{} revealed two places where Finch needed
+protection against user input.
+The first was when reading an input mesh.\footnote{\urlaccess{https://github.com/paralab/Finch/issues/16}{2023-06-06}}
+A NaN injected in the mesh led to a crash further on:
+\begin{verbatim}
+  BoundsError: attempt to access 1-element
+   Vector{Int64} at index [2]
+\end{verbatim}
+Defensive programming solves the issue.
+The second place was in setting bounds for the
+solver.\footnote{\urlaccess{https://github.com/paralab/Finch/issues/17}{2023-06-06}}
+Here, a NaN could leave bounds uninitialized, leading to a bounds error.
+Additionally, \FT{} and \CSTG{}s have been useful for identifying NaNs that
+appear in unstable systems.
+%% advection2d, high cfl and high T (time) led to NaN during simulation
 
@@
 \section{Discussion}
+
+\subsection{Enhanced Fuzzing}
+
+While fuzzing is useful for discovering issues, its success rate
+is low because \emph{every} floating-point operation is a candidate
+for injection.
+Even operations that are already well-defended against NaNs are candidates.
+\FT{} could use two sorts of tools for improving injection.
+First, fine-grained control to let users decide where not to inject.
+Second, tools for understanding the context of an injection point
+after the fact.
+Program slicing is especially relevant to the latter point (and effectively
+what we did by hand when fuzzing Finch~\cref{s:finch}).
+For each operation, an expert needs to study the values that feed into it to
+decide whether they are protected or not.
+
+%% TODO example, start here at glbvertex https://github.com/paralab/Finch/blob/master/src/grid.jl#L360
 

### 2023-06-07 paper.tex
Add section to discussion, round out shallow waters section

@@
 %     NaN fuzzing to find a kill: OrdinaryDiffEq
-%     NaN Kills: Finch
+%     Finch
 %     A Bug-Hunting Safari
@@
 %   Discussion
+%     Tracking more than NaNs
+%     Enhanced Fuzzing
 %   Related Work
@@
 In figure~\ref{fig:nan-gens} we can tell that this must be the result of $\infty - \infty$.
-Where did the $\infty$ come from?
 
-% What do we say here?!
+Thus, with little effort we were able to track down where the NaNs were originating.
+More work is needed to follow the $\infty$ back to wherever \emph{it} came from.
+Fortunately, we will be able to extend FloatTracker to handle more than just NaNs---watching and logging \texttt{Inf}s should be completely tractable.
+See~\cref{discussion} for more on future plans around this.
+
+% TODO: if we get some handling with Inf working, then we should keep tracing this back.
+% Where did the $\infty$ come from?
 
@@
 \section{Discussion}
+\label{s:discussion}
+
+\subsection{Tracking more than NaNs}
+
+% This is mostly a stand-in in case we don't get around to adding Inf tracking.
+% It should be easy to do, but whether or not we have enough time to do it is
+% another matter.
+FloatTracker is capable of monitoring all sorts of events in a floating point value's lifetime---not just a value going to NaN.
+Extending FloatTracker to handle \texttt{Inf} is just a matter of engineering.
+Other values of interest may follow.
 
@@
 % Local Variables:
-% jinx-local-words: "CFL CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq ShallowWaters TrackedFloat isnan"
+% jinx-local-words: "CFL CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq RxInfer ShallowWaters TrackedFloat isnan"
 % End:

### 2023-06-08 paper.tex
Minor restructure, talk about Inf

@@
 %     To Kill a Floating Point
+%     FloatTracker
+%   Floating Point Exception Primer
 %     Where do NaNs come from?
-%     NaN kills
+%     Infinity
 %   FloatTracker Internals
@@
 
-\section{To Kill a Floating Point}
+% Here is a short, illustrative example about the problem.
+\subsection{To Kill a Floating Point}
+\label{s:to-kill-a-fp}
 % If anyone has a better "kill" pun/reference than "To Kill a Mockingbird", I'm all ears.
-% I think we will definitely need a clear, detailed section on how NaNs crop up and how they can be killed.
 
@@
 
+\subsection{FloatTracker}
+% I'd love to find some Sherlock Holmes title for this section, as it would be a subtle nod to Sherlogs.jl and would be a nice literary reference to compliment "To Kill a Mockingbird"
+
+\section{Floating Point Exception Primer}
+
+In our work we focused on two of the most common problematic values in the floating point space: NaN and Inf.
+We leave the possibility of tracking other kinds of troublesome values (e.g. subnormal numbers) open, but for our work we restrict ourselves to just these values.
+
 \subsection{Where do NaNs come from?}
@@
 
-% Yadda yadda yadda...
-
-\subsection{NaN kills}
+When one of these events happens, we call it a NaN \emph{generation}.
 
 Most mathematical operations preserve NaNs.
-However, there are some instances where NaNs can disappear, but not before potentially altering the correctness of the program!
+We call these instances NaN \emph{propagation}, as the NaN is propagated from the source in the arguments to the result.
 
-% \begin{itemize}
-% \item $x \; \texttt{cmp} \; y \rightarrow false$ where \texttt{cmp} is a binary comparison like \texttt{<}, \texttt{>}, \texttt{<=}, etc. and $x$ or $y$ is NaN.
-% \item $1^{NaN} \rightarrow 1$
-% \item $NaN^{0} \rightarrow 1$
-% \end{itemize}
+There are, however, some instances where the IEEE standard requires that NaNs to \emph{not} propagate from arguments to the result; we call these NaN \emph{kills}.
+Sometimes this can be good---an \texttt{isnan} check should clearly return a boolean and not a NaN---but kills are often harmful in two ways:
+
+\begin{enumerate}
+  \item Users no longer see that there was some errant operation: NaNs are an opportunity to closely examine data for problems and code for unsoundness.
+  \item NaN kills can influence the control of a program leading to incorrect results, as we saw in~\cref{to-kill-a-fp}.
+\end{enumerate}
+
+Here are cases where, according to the IEEE standard, a NaN can be killed:
 
@@
 
-% Yadda yadda yadda...
+FloatTracker monitors these and all other operations for interesting events.
 
-% What else do we need to talk about? What other problems do we need to set up so that FloatTracker's solutions are clear?
+\subsection{Infinity}
+
+There are two infinities, $+\infty$ and $-\infty$, but they behave the same for our purposes, so we will refer to them just as ``Inf''.
+Like NaNs, Inf can crop up when bad math happens: $\frac{x}{0} \rightarrow \infty$ when $x \ne 0$.
+Inf can also arise when numbers get too big or too small for the choice of representation.
+
+% TODO
 

### 2023-06-12 paper.tex
Add notes on tracking down Inf, add CSTG diagrams

@@
 
-\subsection{Catching NaN kills}
+\subsection{Catching a NaN Killer}
 
@@
 …
--                 FloatTracker/TrackedFloat.jl:106
+-                 FT/TrackedFloat.jl:106
 momentum_u!       ShallowWaters/rhs.jl:246
@@
 In figure~\ref{fig:nan-gens} we can tell that this must be the result of $\infty - \infty$.
+Thus, with little effort we were able to track down where the NaNs were originating.
+Now the question is: where did the $\infty$ come from?
 
-Thus, with little effort we were able to track down where the NaNs were originating.
-More work is needed to follow the $\infty$ back to wherever \emph{it} came from.
-Fortunately, we will be able to extend FloatTracker to handle more than just NaNs---watching and logging \texttt{Inf}s should be completely tractable.
-See~\cref{discussion} for more on future plans around this.
+FloatTracker is not limited to NaNs---we can track any kind of value that we want.
+We have built-in support for tracking NaN and $\infty$; we can look at the file of the $\infty$ gens where we see the following: (formatted for clarity)
 
-% TODO: if we get some handling with Inf working, then we should keep tracing this back.
-% Where did the $\infty$ come from?
+\begin{verbatim}
+check_error([-1.515f31, 2])  FT/TrackedFloat.jl:10
+^(::TrackedFloat32, ::Int64) FT/TrackedFloat.jl:138
+…
+materialize(^)               broadcast.jl:860
+top-level getproperty(P, :u) examples/sw_nan_tf.jl:14
+…
+\end{verbatim}
+
+It looks like the $-\infty$ came from $-1.515e31^2$---no wonder a Float32 type couldn't handle something that large.
+We can tell that because the first line shows us the arguments\footnote{There is a configuration option that allows us to control this behavior.} and the second line shows us what function got called.
+Moreover, we get the top-level line: we can see that it came from line 14 in our ShallowWaters file.
+
+Tracking down why that value got so large is beyond the scope of FloatTracker right now.
+Most likely we will need some domain-expertise to find a solution to this problem.
+The point is, we were able to quickly and automatically arrive at this excellent starting point with FloatTracker.
+
+\subsubsection{CSTGs help us get a bigger picture}
+
+Running FloatTracker can produce a lot of logs.
+To get a quick overview of the most common paths to problems, we can use CSTG to get an overview of what stack frames appear the most often.
+% FIXME: cite CSTG
+
+We can produce a CSTG of both the Inf as well as the NaN logs.
+Here is what the two graphs look like:\footnote{As with the call stack logs, these have been edited for clarity. We only remove a few boxes in between—mostly due to space constraints in this paper.}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=3in]{fig/sw_nan_cstg_clean.png}
+  \caption{CSTG of NaN generation logs from ShallowWaters.}
+  \label{fig:sw_nan_cstg}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=3in]{fig/sw_inf_cstg_clean.png}
+  \caption{CSTG of Inf generation logs from the same run as~\cref{fig:sw_nan_cstg}.}
+  \label{fig:sw_inf_cstg}
+\end{figure}
 

### 2023-06-13 paper.tex
More talk draft

@@
 
-\subsection{Tracking more than NaNs}
-
-% This is mostly a stand-in in case we don't get around to adding Inf tracking.
-% It should be easy to do, but whether or not we have enough time to do it is
-% another matter.
-FloatTracker is capable of monitoring all sorts of events in a floating point value's lifetime---not just a value going to NaN.
-Extending FloatTracker to handle \texttt{Inf} is just a matter of engineering.
-Other values of interest may follow.
+\subsection{Tracking more than NaNs and Infs}
+FloatTracker is capable of monitoring all sorts of events in a floating point value's lifetime---not just a value going to NaN or Inf.
+We spent most of the time engineering and testing NaN tracking and injection.
+Adding Inf tracking took a few hours.
+Adding other values of interest would be issues of relatively simple engineering.
 

### 2023-06-13 fp_examples.jl

@@ -1,0 +1,17 @@
+n = 0
+i::Float32 = 0.0
+while i < 1
+  global i += 0.1
+  global n += 1
+end
+
+println("With float32: $n iterations")
+
+n = 0
+j::Float64 = 0.0
+while j < 1
+  global j += 0.1
+  global n += 1
+end
+
+println("With float64: $n iterations")

### 2023-06-16 paper.tex
rearrange outline

@@
   Reliable numerical computations are central to high-performance computing and machine learning.
-  We present FlowFPX: a Julia-based tool for tracking the onset and flow of IEEE Floating-Point exceptions that signal numerical defects.
+  We present FlowFPX: a Julia-based tool for tracking the onset and flow of IEEE floating point exceptions.
   FlowFPX's design exploits Julia's operator overloading to trace exception flows and even inject exceptions to accelerate testing.
@@
 \newcommand{\FlowFPX}{FlowFPX}
+\newcommand{\GPUFPX}{GPU-FPX}
 \newcommand{\FloatTracker}{FloatTracker}
@@
 
-% Here is a short, illustrative example about the problem.
-\subsection{To Kill a Floating Point}
-\label{s:to-kill-a-fp}
-% If anyone has a better "kill" pun/reference than "To Kill a Mockingbird", I'm all ears.
-
-The IEEE 754 specification for floating-point numbers governs how NaNs are handled as arguments to arithmetic functions~\cite{IEEEStandardBinary1985}.
-Most of the time, when any or all of the arguments are NaN, the result is also a NaN.
-This is good and desirable, because it means that an arithmetic error---even if buried deep in an expression---will still be observable, allowing a user to find and debug it.
-
-However, there are a few places where NaNs flow into an expression, but not out.
-We call these instances ``NaN-kills'', and these can lead to subtle logic bugs.
-Consider the following example:
-
-\begin{lstlisting}[language = Julia]
-function bad_max(lst)
-  max_seen = 0.0
-  for x in lst
-    if ! (x <= max_seen)
-      # swap if new val greater
-      max_seen = x
-    end
-  end
-  max_seen
-end
-
-function good_max(lst)
-  foldl(max, lst)
-end
-\end{lstlisting}
-
-Both \texttt{bad\_max} and \texttt{good\_max} compute the maximum in a list.
-However, the \texttt{<=} operator used in the \texttt{bad\_max} implementation does \emph{not} propagate NaNs, so you might not get the result you expected:
-
-\begin{lstlisting}[language = Julia]
-bad_max([1, 5, NaN, 4])     # returns 4
-good_max([1, 5, NaN, 4])    # returns NaN
-\end{lstlisting}
-
-Not only is the result from \texttt{bad\_max} problematic for obscuring the fact that there was a NaN in the list, the result is also wrong!
-The built-in \texttt{max} function correctly propagates NaNs, so we can see from its return that we got a NaN somewhere.
-It's not entirely obvious \emph{where} that NaN came from---but more on that later.   % this promise answered in "NaN propagation monitoring"
-
-\subsection{FloatTracker}
-% I'd love to find some Sherlock Holmes title for this section, as it would be a subtle nod to Sherlogs.jl and would be a nice literary reference to compliment "To Kill a Mockingbird"
+kk
 
@@
 
-\section{FloatTracker Internals}
+% Here is a short, illustrative example about the problem.
+\subsection{To Kill a Floating Point}
+\label{s:to-kill-a-fp}
+% If anyone has a better "kill" pun/reference than "To Kill a Mockingbird", I'm all ears.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{fig/genpropkill-outline.png}
+  \caption{Gen, Prop, Kill: Lifetime of a NaN value}
+  \label{f:gpk}
+\end{figure}
+
+The IEEE 754 specification for floating-point numbers governs how NaNs are handled as arguments to arithmetic functions~\cite{IEEEStandardBinary1985}.
+Most of the time, when any or all of the arguments are NaN, the result is also a NaN.
+This is good and desirable, because it means that an arithmetic error---even if buried deep in an expression---will still be observable, allowing a user to find and debug it.
+
+However, there are a few places where NaNs flow into an expression, but not out.
+We call these instances ``NaN-kills'', and these can lead to subtle logic bugs.
+Consider the following example:
+
+\begin{lstlisting}[language = Julia]
+function bad_max(lst)
+  max_seen = 0.0
+  for x in lst
+    if ! (x <= max_seen)
+      # swap if new val greater
+      max_seen = x
+    end
+  end
+  max_seen
+end
+
+function good_max(lst)
+  foldl(max, lst)
+end
+\end{lstlisting}
+
+Both \texttt{bad\_max} and \texttt{good\_max} compute the maximum in a list.
+However, the \texttt{<=} operator used in the \texttt{bad\_max} implementation does \emph{not} propagate NaNs, so you might not get the result you expected:
+
+\begin{lstlisting}[language = Julia]
+bad_max([1, 5, NaN, 4])     # returns 4
+good_max([1, 5, NaN, 4])    # returns NaN
+\end{lstlisting}
+
+Not only is the result from \texttt{bad\_max} problematic for obscuring the fact that there was a NaN in the list, the result is also wrong!
+The built-in \texttt{max} function correctly propagates NaNs, so we can see from its return that we got a NaN somewhere.
+It's not entirely obvious \emph{where} that NaN came from---but more on that later.   % this promise answered in "NaN propagation monitoring"
+
+\section{FlowFPX Toolkit}
 
@@
 
-\FT{} is an effective tool for debugging all sorts of NaN-related issues.
+\FlowFPX{} is effective for debugging all sorts of NaN-related issues.
 We have used it to reveal silent NaNs, uncover the root cause of NaNs,
@@
 
-\section{Evaluation}
+\subsection{\GPUFPX{}: Oceananigans}
+
+FILL
+
+% * what about GPUs? companion project at Utah GPU-FPX
+% * architecture diagram, CPU + GPU, show Ocean source code
+% * replace CPU() with GPU() in code, get output like below & attached
+% 
+% \begin{verbatim}
+% using Oceananigans
+% grid = RectilinearGrid(GPU(), size=(128, 128), x=(0, 2π), y=(0, 2π), topology=(Periodic, Periodic, Flat))
+% model = NonhydrostaticModel(; grid, advection=WENO())
+% ϵ(x, y, z) = 2rand() - 1
+% set!(model, u=ϵ, v=ϵ)
+% simulation = Simulation(model; Δt=0.01, stop_time=4)
+% run!(simulation)
+% \end{verbatim}
+% 
+% ... xinyi ran gpu-fpx on this code ... on the binary it created?
+% 
+% 
+% - - -
+% 
+% 
+% #GPU-FPX LOC-EXCEP INFO: in kernel
+% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
+% DIV0 found @ ./float.jl:335 [FP64]
+% #GPU-FPX LOC-EXCEP INFO: in kernel
+% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
+% NaN found @ ./float.jl:335 [FP64]
+% #GPU-FPX LOC-EXCEP INFO: in kernel
+% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
+% NaN found @ ./float.jl:335 [FP32]
+% #GPU-FPX LOC-EXCEP INFO: in kernel
+% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
+% DIV0 found @ ./int.jl:261 [FP64]
+% #GPU-FPX LOC-EXCEP INFO: in kernel
+% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
+% NaN found @ ./int.jl:261 [FP64]
+% #GPU-FPX LOC-EXCEP INFO: in kernel
+% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
+% INF found @ ./int.jl:261 [FP64]
+% 
+% 
+% \begin{verbatim}
+% --- GPU-FPX Report ---
+% 
+% --- FP64 Operations ---
+% Total NaN found: 2
+% Total INF found: 1
+% Total underflow (subnormal): 0
+% Total division by zero: 2
+% --- FP32 Operations ---
+% Total NaN found: 1
+% Total INF found: 0
+% Total underflow (subnormal): 0
+% Total division by zero: 0
+% --- Other Stats ---
+% Kernels: 21
+% \end{verbatim}
+
+\section{Related Work}
 
@@
 
-\section{Related Work}
-
 \section{Acknowledgments}

### 2023-06-16 paper.tex
intro, GPK fig

@@
   Reliable numerical computations are central to high-performance computing and machine learning.
-  We present FlowFPX: a Julia-based tool for tracking the onset and flow of IEEE floating point exceptions.
+  We present FlowFPX: a Julia-based tool for tracking the onset and flow of floating-point exceptions.
   FlowFPX's design exploits Julia's operator overloading to trace exception flows and even inject exceptions to accelerate testing.
@@
 \newcommand{\FT}{\FloatTracker}
+\newcommand{\fp}{floating-point} % hyphen or no?
 \newcommand{\CSTG}{stack graph}
@@
 \newcommand{\urlaccess}[2]{\url{#1}}
+\newcommand{\Nan}{\code{NaN}}
+\newcommand{\NaN}{\Nan}
+\newcommand{\Inf}{\code{Inf}}
 
@@
 
-kk
+Reliable numeric computations are central to high-performance computing,
+machine learning, and scientific applications.
+Yet, the \fp{} arithmetic that powers these computations is fundamentally
+unreliable~(\cref{s:background}).
+Exceptional values, namely Not a Number (\Nan{}) and infinity (\Inf{}),
+can and often do arise thanks to culprits such as roundoff error,
+catastrophic cancellation, singular matrices, and vanishing
+derivatives~\cite{sdjmrstp-pc-2022,ddghlllprr-correctness-2022,gllprt-correctness-2021,fpchecker-reports,llg-soap-2022,bllmg-xloop-2022}.
+When exceptions happen, domain experts must take time away from their main
+focus and trek down a rabbit hole to find the source of the error.
+There is little tool support to assist in this difficult task,
+thus (unsurpringly) a quick GitHub search reports over 4,000 open issues
+related to \NaN{} or \Inf{} exceptions~\cite{github-issues}.
+
+We introduce \FlowFPX{}, a toolkit for debugging
+\fp{} exceptions in an easy and intuitive way~(\cref{s:flowfpx}):
+\begin{itemize}
+  \item
+    The centerpiece of \FlowFPX{} is \FT{}, a dynamic analysis tool that uses
+    Julia's operator overloading to selectively monitor for exceptions and to
+    fuzz code for vulnerabilities.
+  \item
+    To visualize results, \FT{} adapts coalesced stack-trace graphs 
+    (CSTGs)~\cite{hmbdrg-cse-2014} for a whole-program view on the generation,
+    propagation, and death of exceptional values.
+  \item \GPUFPX{}~[CITE] provides fine-grained insights for GPU kernels.
+\end{itemize}
+%
+Together, the \FlowFPX{} tools have helped debug exceptions in a variety
+of applications: from ocean simulations to heat models~(\cref{s:casestudies}).
+That said, there is much more to explore.
+\FlowFPX{} is a growing project and we welcome feedback from domain experts
+on how to best suit their workflow.
+
+This paper concludes with related work~(\cref{s:related}) and a
+brief discussion~(\cref{s:discussion}).
 
 \section{Floating Point Exception Primer}
+\label{s:background}
 
-In our work we focused on two of the most common problematic values in the floating point space: NaN and Inf.
+In our work we focused on two of the most common problematic values in the \fp{} space: NaN and Inf.
 We leave the possibility of tracking other kinds of troublesome values (e.g. subnormal numbers) open, but for our work we restrict ourselves to just these values.
@@
   \item Users no longer see that there was some errant operation: NaNs are an opportunity to closely examine data for problems and code for unsoundness.
-  \item NaN kills can influence the control of a program leading to incorrect results, as we saw in~\cref{to-kill-a-fp}.
+  \item NaN kills can influence the control of a program leading to incorrect results, as we saw in~\cref{s:to-kill-a-fp}.
 \end{enumerate}
@@
 \subsection{Infinity}
+
+%%ben: shouldn't this come before NaN b/c infs lead to NaNs?
 
@@
 
-The IEEE 754 specification for floating-point numbers governs how NaNs are handled as arguments to arithmetic functions~\cite{IEEEStandardBinary1985}.
+The IEEE 754 specification for \fp{} numbers governs how NaNs are handled as arguments to arithmetic functions~\cite{IEEEStandardBinary1985}.
 Most of the time, when any or all of the arguments are NaN, the result is also a NaN.
@@
 
-\section{FlowFPX Toolkit}
+\section{\FlowFPX{}}
+\label{s:flowfpx}
 
-FloatTracker works by leveraging Julia's type-based method dispatch system to capture everything from calls to the standard library down to basic arithmetic operations.
+FILL intro
+
+\subsection{\FT{}}
+\label{s:floattracker}
+
+\FT{} works by leveraging Julia's type-based method dispatch system to capture everything from calls to the standard library down to basic arithmetic operations.
 We do this by creating custom types \texttt{TrackedFloat16}, \texttt{TrackedFloat32}, and \texttt{TrackedFloat64} that wrap their corresponding \texttt{Float16}, \texttt{Float32} etc. counterparts.
@@
 
-With a floating point value wrapped in our \texttt{TrackedFloat} type, it is easy to monitor when a NaN gets killed.
+With a \fp{} value wrapped in our \texttt{TrackedFloat} type, it is easy to monitor when a NaN gets killed.
 Julia dispatches all arithmetic operations on that value to flow through FloatTracker, where we can inspect the arguments.
@@
 \section{Case Studies}
+\label{s:casestudies}
 
@@
 Fortunately in this case, one of the key features \texttt{ShallowWaters} is its \emph{type flexibility}:
-\texttt{ShallowWaters} was designed to showcase physics simulations using variable levels of floating-point precision.
+\texttt{ShallowWaters} was designed to showcase physics simulations using variable levels of \fp{} precision.
 Upon model construction, \texttt{ShallowWaters} allows you to specify a type for floats in the model.
@@
 \section{Related Work}
+\label{s:related}
 
@@
 \label{s:discussion}
+% future work
 
 \subsection{Tracking more than NaNs and Infs}
-FloatTracker is capable of monitoring all sorts of events in a floating point value's lifetime---not just a value going to NaN or Inf.
+FloatTracker is capable of monitoring all sorts of events in a \fp{} value's lifetime---not just a value going to NaN or Inf.
 We spent most of the time engineering and testing NaN tracking and injection.
@@
 While fuzzing is useful for discovering issues, its success rate
-is low because \emph{every} floating-point operation is a candidate
+is low because \emph{every} \fp{} operation is a candidate
 for injection.
@@
 %% TODO example, start here at glbvertex https://github.com/paralab/Finch/blob/master/src/grid.jl#L360
+
+\subsection{Static FPX, Gradual Types}
+
+FILL
 

### 2023-06-16 paper.tex
Update helper TOC; add note
NaN → \NaN{}

@@
 %   Introduction
-%     To Kill a Floating Point
-%     FloatTracker
 %   Floating Point Exception Primer
@@
 %     Infinity
-%   FloatTracker Internals
-%     Catching NaN kills
+%     To Kill a Floating Point
+%   \FlowFPX{
+%     \FT{
+%     Catching a NaN Killer
 %     NaN propagation monitoring
@@
 %     Surprise NaNs: ShallowWaters
+%       CSTGs help us get a bigger picture
 %     NaN fuzzing to find a kill: OrdinaryDiffEq
@@
 %     A Bug-Hunting Safari
-%   Evaluation
+%     \GPUFPX{
+%   Related Work
 %   Discussion
-%     Tracking more than NaNs
+%     Tracking more than NaNs and Infs
 %     Enhanced Fuzzing
-%   Related Work
+%     Static FPX, Gradual Types
 %   Acknowledgments
@@
 
-In our work we focused on two of the most common problematic values in the \fp{} space: NaN and Inf.
+% TODO: talk about how floating point != real numbers → errors appear and build up
+% (see other tutorials; Art of Computer Programming, Ganesh, slides in Google Drive from guy in England)
+% 2 paragraphs or so: floating point primer
+
+In our work we focused on two of the most common problematic values in the \fp{} space: \NaN{} and \Inf{}.
 We leave the possibility of tracking other kinds of troublesome values (e.g. subnormal numbers) open, but for our work we restrict ourselves to just these values.
 
-\subsection{Where do NaNs come from?}
+\subsection{Where do \NaN{}s come from?}
 
-IEEE 754~\cite{IEEEStandardBinary1985} lists the following operations (among a few others) where NaNs result:
+IEEE 754~\cite{IEEEStandardBinary1985} lists the following operations (among a few others) where \NaN{}s result:
 
@@
   \end{itemize}
-  \caption{NaN-producing operations}
+  \caption{\NaN{}-producing operations}
 \end{figure}
 
-When one of these events happens, we call it a NaN \emph{generation}.
+When one of these events happens, we call it a \NaN{} \emph{generation}.
 
-Most mathematical operations preserve NaNs.
-We call these instances NaN \emph{propagation}, as the NaN is propagated from the source in the arguments to the result.
+Most mathematical operations preserve \NaN{}s.
+We call these instances \NaN{} \emph{propagation}, as the \NaN{} is propagated from the source in the arguments to the result.
 
-There are, however, some instances where the IEEE standard requires that NaNs to \emph{not} propagate from arguments to the result; we call these NaN \emph{kills}.
-Sometimes this can be good---an \texttt{isnan} check should clearly return a boolean and not a NaN---but kills are often harmful in two ways:
+There are, however, some instances where the IEEE standard requires that \NaN{}s to \emph{not} propagate from arguments to the result; we call these \NaN{} \emph{kills}.
+Sometimes this can be good---an \texttt{isnan} check should clearly return a boolean and not a \NaN{}---but kills are often harmful in two ways:
 
 \begin{enumerate}
-  \item Users no longer see that there was some errant operation: NaNs are an opportunity to closely examine data for problems and code for unsoundness.
-  \item NaN kills can influence the control of a program leading to incorrect results, as we saw in~\cref{s:to-kill-a-fp}.
+  \item Users no longer see that there was some errant operation: \NaN{}s are an opportunity to closely examine data for problems and code for unsoundness.
+  \item \NaN{} kills can influence the control of a program leading to incorrect results, as we saw in~\cref{s:to-kill-a-fp}.
 \end{enumerate}
 
-Here are cases where, according to the IEEE standard, a NaN can be killed:
+Here are cases where, according to the IEEE standard, a \NaN{} can be killed:
 
@@
 x \; \texttt{cmp} \; y &\rightarrow false &\text{where \texttt{cmp} is \texttt{<}, \texttt{>}, \texttt{<=}, etc.} \\
-1^{\text{\texttt{NaN}}} &\rightarrow 1 \\
-\text{\texttt{NaN}}^{0} &\rightarrow 1
+1^{\text{\texttt{\NaN{}}}} &\rightarrow 1 \\
+\text{\texttt{\NaN{}}}^{0} &\rightarrow 1
 \end{align*}
@@
 
-There are two infinities, $+\infty$ and $-\infty$, but they behave the same for our purposes, so we will refer to them just as ``Inf''.
-Like NaNs, Inf can crop up when bad math happens: $\frac{x}{0} \rightarrow \infty$ when $x \ne 0$.
-Inf can also arise when numbers get too big or too small for the choice of representation.
-
-% TODO
+There are two infinities, $+\infty$ and $-\infty$, but they behave the same for our purposes, so we will refer to them just as ``\Inf{}''.
+Like \NaN{}s, \Inf{} can crop up when bad math happens: $\frac{x}{0} \rightarrow \infty$ when $x \ne 0$.
+\Inf{} can also arise when numbers get too big or too small for the choice of representation.
 
@@
   \includegraphics[width=\columnwidth]{fig/genpropkill-outline.png}
-  \caption{Gen, Prop, Kill: Lifetime of a NaN value}
+  \caption{Gen, Prop, Kill: Lifetime of a \NaN{} value}
   \label{f:gpk}
@@
 
-The IEEE 754 specification for \fp{} numbers governs how NaNs are handled as arguments to arithmetic functions~\cite{IEEEStandardBinary1985}.
-Most of the time, when any or all of the arguments are NaN, the result is also a NaN.
+The IEEE 754 specification for \fp{} numbers governs how \NaN{}s are handled as arguments to arithmetic functions~\cite{IEEEStandardBinary1985}.
+Most of the time, when any or all of the arguments are \NaN{}, the result is also a \NaN{}.
 This is good and desirable, because it means that an arithmetic error---even if buried deep in an expression---will still be observable, allowing a user to find and debug it.
 
-However, there are a few places where NaNs flow into an expression, but not out.
-We call these instances ``NaN-kills'', and these can lead to subtle logic bugs.
+However, there are a few places where \NaN{}s flow into an expression, but not out.
+We call these instances ``\NaN{}-kills'', and these can lead to subtle logic bugs.
 Consider the following example:
@@
 Both \texttt{bad\_max} and \texttt{good\_max} compute the maximum in a list.
-However, the \texttt{<=} operator used in the \texttt{bad\_max} implementation does \emph{not} propagate NaNs, so you might not get the result you expected:
+However, the \texttt{<=} operator used in the \texttt{bad\_max} implementation does \emph{not} propagate \NaN{}s, so you might not get the result you expected:
 
@@
 
-Not only is the result from \texttt{bad\_max} problematic for obscuring the fact that there was a NaN in the list, the result is also wrong!
-The built-in \texttt{max} function correctly propagates NaNs, so we can see from its return that we got a NaN somewhere.
-It's not entirely obvious \emph{where} that NaN came from---but more on that later.   % this promise answered in "NaN propagation monitoring"
+Not only is the result from \texttt{bad\_max} problematic for obscuring the fact that there was a \NaN{} in the list, the result is also wrong!
+The built-in \texttt{max} function correctly propagates \NaN{}s, so we can see from its return that we got a \NaN{} somewhere.
+It's not entirely obvious \emph{where} that \NaN{} came from---but more on that later.   % this promise answered in "NaN propagation monitoring"
 
@@
 \begin{enumerate}
-  \item Catch instances of NaN kills
-  \item Monitor the propagation of NaNs through the program
-  \item Optionally \emph{inject} NaNs to fuzz the program under scrutiny against NaN-kills
+  \item Catch instances of \NaN{} kills
+  \item Monitor the propagation of \NaN{}s through the program
+  \item Optionally \emph{inject} \NaN{}s to fuzz the program under scrutiny against \NaN{}-kills
 \end{enumerate}
@@
 
-\subsection{Catching a NaN Killer}
+\subsection{Catching a \NaN{} Killer}
 
-With a \fp{} value wrapped in our \texttt{TrackedFloat} type, it is easy to monitor when a NaN gets killed.
+With a \fp{} value wrapped in our \texttt{TrackedFloat} type, it is easy to monitor when a \NaN{} gets killed.
 Julia dispatches all arithmetic operations on that value to flow through FloatTracker, where we can inspect the arguments.
-Kills are relatively simple: if one or more NaNs flow into a function but don't flow out (i.e. the answer isn't a NaN) then that's a kill.
+Kills are relatively simple: if one or more \NaN{}s flow into a function but don't flow out (i.e. the answer isn't a \NaN{}) then that's a kill.
 For example:
@@
   NaN + 42     -> NaN    # not a kill
-  Inf - Inf    -> NaN    # gen, but not a kill
+  \Inf{} - \Inf{}    -> NaN    # gen, but not a kill
   min(NaN, 42) -> NaN    # not a kill
@@
 
-There are a few exceptions to the rule of ``NaN in and no NaN out $\rightarrow$ kill''---some instances that we report as being a NaN kill might be instead a \emph{NaN-repair}, i.e.
-where the user explicitly checks for a NaN and gets some non-NaN value as a result.
+There are a few exceptions to the rule of ``\NaN{} in and no \NaN{} out $\rightarrow$ kill''---some instances that we report as being a \NaN{} kill might be instead a \emph{\NaN{}-repair}, i.e.
+where the user explicitly checks for a \NaN{} and gets some non-\NaN{} value as a result.
 Once clear instance of this is the \texttt{isnan} function.
 Calling \texttt{isnan} and getting a boolean is expected behavior, and is usually used in error handling or in \emph{explicit} handling of unexpected values.
-FloatTracker still reports these cases in the kill logs so as to be complete---the user must not blindly assume that every instance of a NaN kill is bad and might be a case of NaN repair.
+FloatTracker still reports these cases in the kill logs so as to be complete---the user must not blindly assume that every instance of a \NaN{} kill is bad and might be a case of \NaN{} repair.
 
@@
 
-\subsection{NaN propagation monitoring}
+\subsection{\NaN{} propagation monitoring}
 
-Remember in our first example with the \texttt{bad\_max} and \texttt{good\_max} functions how we didn't have a way to tell \emph{where} the NaNs originally came from?
+Remember in our first example with the \texttt{bad\_max} and \texttt{good\_max} functions how we didn't have a way to tell \emph{where} the \NaN{}s originally came from?
 That can be a tedious task.
-However, FloatTracker makes this a lot easier by watching the life cycle of a NaN.
+However, FloatTracker makes this a lot easier by watching the life cycle of a \NaN{}.
 
@@
 
-\subsection{NaN injection}
+\subsection{\NaN{} injection}
 
@@
 
-\FlowFPX{} is effective for debugging all sorts of NaN-related issues.
-We have used it to reveal silent NaNs, uncover the root cause of NaNs,
+\FlowFPX{} is effective for debugging all sorts of \NaN{}-related issues.
+We have used it to reveal silent \NaN{}s, uncover the root cause of \NaN{}s,
 and fuzz for weak points in several case studies.
 
-\subsection{Surprise NaNs: ShallowWaters}
+\subsection{Surprise \NaN{}s: ShallowWaters}
 
@@
 
-It is not uncommon to encounter cases where setting the CFL parameter too high results in NaNs in the result.
-Driving the CFL parameter low enough to excise the NaNs can make the simulation too long and unwieldy to be comfortable to use.
-Identifying where NaNs begin to arise can help point the way to where some mathematical refinement can be used to make the simulation both fast and accurate enough for practical use.
+It is not uncommon to encounter cases where setting the CFL parameter too high results in \NaN{}s in the result.
+Driving the CFL parameter low enough to excise the \NaN{}s can make the simulation too long and unwieldy to be comfortable to use.
+Identifying where \NaN{}s begin to arise can help point the way to where some mathematical refinement can be used to make the simulation both fast and accurate enough for practical use.
 
@@
 That said, it works for our purposes as a proving ground for FloatTracker.
-We dialed the CFL parameter up to high values (as high as 10) and we started seeing NaNs in the result.
+We dialed the CFL parameter up to high values (as high as 10) and we started seeing \NaN{}s in the result.
 Note that the program did not crash: instead, it began producing bad output.
@@
   \includegraphics[width=3in]{./fig/shallow_waters_cfl_diff.png}
-  \caption{Effect of high CFL parameters: on the left, normal output. On the right: NaNs cause white holes and gaps in the result.}
+  \caption{Effect of high CFL parameters: on the left, normal output. On the right: \NaN{}s cause white holes and gaps in the result.}
   \label{fig:sw_nans}
@@
 
-The holes in the plots are regions where NaNs crept into the computation and destroyed the results.
-In one regard, this is not as bad as a NaN kill---knowing that there are NaNs in the computation means that we have a chance to fix it, whereas a NaN kill might silently give us the wrong answer.
-Incidentally, we know that there were no NaN kills in this test because our kill logs are empty.
-But finding the sources of NaNs can be tricky.
-We expected NaNs in this case because we set the CFL parameter up so high, but tracking down the sources of NaNs generally is a difficult and time-consuming problem.
+The holes in the plots are regions where \NaN{}s crept into the computation and destroyed the results.
+In one regard, this is not as bad as a \NaN{} kill---knowing that there are \NaN{}s in the computation means that we have a chance to fix it, whereas a \NaN{} kill might silently give us the wrong answer.
+Incidentally, we know that there were no \NaN{} kills in this test because our kill logs are empty.
+But finding the sources of \NaN{}s can be tricky.
+We expected \NaN{}s in this case because we set the CFL parameter up so high, but tracking down the sources of \NaN{}s generally is a difficult and time-consuming problem.
 
@@
 
-We ran ShallowWaters with a CFL value known to make bad outputs and recorded NaN \texttt{gen} events:
+We ran ShallowWaters with a CFL value known to make bad outputs and recorded \NaN{} \texttt{gen} events:
 
@@
 
-We can see that the NaN appeared from subtraction (first line) and that it was used in the \texttt{momentum\_u!} function on line 246 of the \texttt{rhs.jl} file (second line).
+We can see that the \NaN{} appeared from subtraction (first line) and that it was used in the \texttt{momentum\_u!} function on line 246 of the \texttt{rhs.jl} file (second line).
 In figure~\ref{fig:nan-gens} we can tell that this must be the result of $\infty - \infty$.
-Thus, with little effort we were able to track down where the NaNs were originating.
+Thus, with little effort we were able to track down where the \NaN{}s were originating.
 Now the question is: where did the $\infty$ come from?
 
-FloatTracker is not limited to NaNs---we can track any kind of value that we want.
-We have built-in support for tracking NaN and $\infty$; we can look at the file of the $\infty$ gens where we see the following: (formatted for clarity)
+FloatTracker is not limited to \NaN{}s---we can track any kind of value that we want.
+We have built-in support for tracking \NaN{} and $\infty$; we can look at the file of the $\infty$ gens where we see the following: (formatted for clarity)
 
@@
 
-We can produce a CSTG of both the Inf as well as the NaN logs.
+We can produce a CSTG of both the \Inf{} as well as the \NaN{} logs.
 Here is what the two graphs look like:\footnote{As with the call stack logs, these have been edited for clarity. We only remove a few boxes in between—mostly due to space constraints in this paper.}
@@
   \includegraphics[width=3in]{fig/sw_nan_cstg_clean.png}
-  \caption{CSTG of NaN generation logs from ShallowWaters.}
+  \caption{CSTG of \NaN{} generation logs from ShallowWaters.}
   \label{fig:sw_nan_cstg}
@@
   \includegraphics[width=3in]{fig/sw_inf_cstg_clean.png}
-  \caption{CSTG of Inf generation logs from the same run as~\cref{fig:sw_nan_cstg}.}
+  \caption{CSTG of \Inf{} generation logs from the same run as~\cref{fig:sw_nan_cstg}.}
   \label{fig:sw_inf_cstg}
@@
 
-\subsection{NaN fuzzing to find a kill: OrdinaryDiffEq}
+\subsection{\NaN{} fuzzing to find a kill: OrdinaryDiffEq}
 
 The second thing we evaluated was an N-Body simulation library.
-We didn't find any sensible configuration of input parameters that lead to NaN kills like in the ShallowWaters library.
-We turned to the NaN injection capabilities of FloatTracker to fuzz the library to find any lurking bugs.
+We didn't find any sensible configuration of input parameters that lead to \NaN{} kills like in the ShallowWaters library.
+We turned to the \NaN{} injection capabilities of FloatTracker to fuzz the library to find any lurking bugs.
 % FIXME: add hyper-ref to § 4.1: ShallowWaters
 
-We configured FloatTracker to inject a single NaN to see if we could find any NaN kills.
-On some runs we noticed that we would get a kill and the program would warn about a NaN and prepare to exit.
+We configured FloatTracker to inject a single \NaN{} to see if we could find any \NaN{} kills.
+On some runs we noticed that we would get a kill and the program would warn about a \NaN{} and prepare to exit.
 Contrary to the error message's claim, however, the program went into an infinite loop.
@@
 The logs lead us to a routine in the widely-used \texttt{OrdinaryDiffEq} library.
-During initialization, a NaN got injected during the execution of \texttt{-} between two \texttt{TrackedFloat}s:
+During initialization, a \NaN{} got injected during the execution of \texttt{-} between two \texttt{TrackedFloat}s:
 
@@
 
-\texttt{sign} correctly propagates NaNs, so \texttt{tdir} now contains a NaN.
+\texttt{sign} correctly propagates \NaN{}s, so \texttt{tdir} now contains a \NaN{}.
 
 We \emph{could} trace this variable through the code, but that would be a lot of grunt work.
-Instead, we can look at the NaN kill logs to get a clue as to where we should look next:
+Instead, we can look at the \NaN{} kill logs to get a clue as to where we should look next:
 
@@
 
-The problem here is on line 2 when \texttt{tdir} is NaN: multiplication propagates NaNs, and comparison kills it, so the condition on the inner \texttt{while} loop is \emph{always} false.
+The problem here is on line 2 when \texttt{tdir} is \NaN{}: multiplication propagates \NaN{}s, and comparison kills it, so the condition on the inner \texttt{while} loop is \emph{always} false.
 No work got done in the inner loop, and so the conditional \texttt{pop} routine never reduced the size of the \texttt{time\_stops} vector.
 
-Our kill logs led us right to this line; this is a perfect example of how NaN kills can influence control flow to go awry.
+Our kill logs led us right to this line; this is a perfect example of how \NaN{} kills can influence control flow to go awry.
 In this case, the problem was apparent besides what our logs told us and it manifested as an infinite loop.
@@
 The first was when reading an input mesh.\footnote{\urlaccess{https://github.com/paralab/Finch/issues/16}{2023-06-06}}
-A NaN injected in the mesh led to a crash further on:
+A \NaN{} injected in the mesh led to a crash further on:
 \begin{verbatim}
@@
 solver.\footnote{\urlaccess{https://github.com/paralab/Finch/issues/17}{2023-06-06}}
-Here, a NaN could leave bounds uninitialized, leading to a bounds error.
-Additionally, \FT{} and \CSTG{}s have been useful for identifying NaNs that
+Here, a \NaN{} could leave bounds uninitialized, leading to a bounds error.
+Additionally, \FT{} and \CSTG{}s have been useful for identifying \NaN{}s that
 appear in unstable systems.
@@
 
-We searched GitHub for open issues in Julia libraries involving difficult-to-find NaNs.
-In our first batch of issues, we found the \code{RxInfer.jl} library with an issue revolving around NaN detection.\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}}
-We offered our tool for the package authors to use, and within a day the authors were able to track down the location of a NaN and fix the issue---all with minimal guidance.
+We searched GitHub for open issues in Julia libraries involving difficult-to-find \NaN{}s.
+In our first batch of issues, we found the \code{RxInfer.jl} library with an issue revolving around \NaN{} detection.\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}}
+We offered our tool for the package authors to use, and within a day the authors were able to track down the location of a \NaN{} and fix the issue---all with minimal guidance.
 
@@
 
-\subsection{Tracking more than NaNs and Infs}
-FloatTracker is capable of monitoring all sorts of events in a \fp{} value's lifetime---not just a value going to NaN or Inf.
-We spent most of the time engineering and testing NaN tracking and injection.
-Adding Inf tracking took a few hours.
+\subsection{Tracking more than \NaN{}s and \Inf{}s}
+FloatTracker is capable of monitoring all sorts of events in a \fp{} value's lifetime---not just a value going to \NaN{} or \Inf{}.
+We spent most of the time engineering and testing \NaN{} tracking and injection.
+Adding \Inf{} tracking took a few hours.
 Adding other values of interest would be issues of relatively simple engineering.
@@
 for injection.
-Even operations that are already well-defended against NaNs are candidates.
+Even operations that are already well-defended against \NaN{}s are candidates.
 \FT{} could use two sorts of tools for improving injection.

### 2023-06-16 paper.tex
sec headers for viz, gpu

@@
 % If using Emacs, put cursor on last closing paren and call "eval-last-sexp" (usually `C-x C-e') to regenerate
-\iffalse
-(save-excursion
-  (let* ((toc (funcall imenu-create-index-function))
-         (headers (mapcar (pcase-lambda (`(,str . ,loc)) str)
-                          toc)))
-    (beginning-of-buffer)
-    (search-forward (concat "%% " "--- autogenerated TOC ---"))
-    (next-line)
-    (beginning-of-line)
-    (let ((start (point)))
-      (search-forward (concat "%% " "--- end autogenerated TOC ---"))
-      (beginning-of-line)
-      (delete-region start (point))
-      (goto-char start))
-    (cl-loop for header in headers
-             do (insert (format "%% %s\n" header)))
-    (next-line)))
-\fi
+%\iffalse
+%(save-excursion
+%  (let* ((toc (funcall imenu-create-index-function))
+%         (headers (mapcar (pcase-lambda (`(,str . ,loc)) str)
+%                          toc)))
+%    (beginning-of-buffer)
+%    (search-forward (concat "%% " "--- autogenerated TOC ---"))
+%    (next-line)
+%    (beginning-of-line)
+%    (let ((start (point)))
+%      (search-forward (concat "%% " "--- end autogenerated TOC ---"))
+%      (beginning-of-line)
+%      (delete-region start (point))
+%      (goto-char start))
+%    (cl-loop for header in headers
+%             do (insert (format "%% %s\n" header)))
+%    (next-line)))
+%\fi
 
@@
     To visualize results, \FT{} adapts coalesced stack-trace graphs 
-    (CSTGs)~\cite{hmbdrg-cse-2014} for a whole-program view on the generation,
+    (CSTGs, or \CSTG{}s)~\cite{hmbdrg-cse-2014} for a whole-program view on the generation,
     propagation, and death of exceptional values.
@@
 
-FILL intro
+FILL intro: float tracker, viz tools, gpu
 
@@
 
-\subsection{Catching a \NaN{} Killer}
+\subsubsection{Catching a \NaN{} Killer}
 
@@
 
-\subsection{\NaN{} propagation monitoring}
+\subsubsection{\NaN{} propagation monitoring}
 
@@
 
-\subsection{\NaN{} injection}
+\subsubsection{\NaN{} injection}
 
 % TODO
+
+\subsection{Visualization}
+\label{s:cstg}
+
+FILL cstg output, how to read, float tracker interface,
+
+\subsection{\GPUFPX{}}
+\label{s:gpufpx}
+
+FILL basic idea, loose coupling, typical overhead,
 

### 2023-06-16 paper.tex
Short draft of fp primer

@@
 \newcommand{\FT}{\FloatTracker}
+\newcommand{\Fp}{Floating-point} % hyphen or no?
 \newcommand{\fp}{floating-point} % hyphen or no?
@@
 % 2 paragraphs or so: floating point primer
+
+\Fp{} numbers are a way of representing real numbers in a computer.
+However, since we limit our precision to a fixed, finite set of bits, \fp{} is \emph{inherently approximate} and necessarily incurs some error during computation.
+This is not bad; it is a trade-off: we get fast and compact representations that are good enough for most use cases.
+However, the downside is that \fp{} numbers \emph{do not} behave like real numbers.
+
+There are many excellent resources\cite{knuthArtComputerProgramming1997} on understanding the theory behind \fp{};
+we will explain the most essential aspects here.
+We represent \fp{} values $r \in \mathbb{R}$ in scientific notation $r = f \times b^{e-q}$,
+for some exponent $e$, a fixed value called an \emph{excess} $q$,\footnote{The purpose of the excess $q$ is so that with value of $e > q$ we can represent positive exponents, and $e < q$ we can represent negative exponents.} and fraction $f$.\footnote{Another name for the fraction part commonly seen in the literature is ``mantissa''.
+  You might occasionally also see ``characteristic'' referring to the exponent.
+  We will avoid those terms however.}
+
+We round and pack $e$ and $f$ into bit fields.
+This gives us a compact representation for both very small and very large values.
+
+Things get tricky when we start to do math when there is a big disparity between exponent values.
+For example, when adding two numbers $u$ and $v$, if the exponents differ by too much, the result of the addition will simply be the larger value.
+
+% TODO illustrative example: e.g. 1,234,567 + 0.0000001 = 1,234,567
+
+The gradual accumulation of these errors leads to surprising mathematical results!
+% FIXME: more examples of a NaN popping up out of nowhere, e.g. with bad addition
 

### 2023-06-17 paper.tex
Edit primer on floating point

--- 2023-06-16 paper.tex
@@
 % 2 paragraphs or so: floating point primer
+
+\Fp{} numbers are a way of representing real numbers in a computer.
+However, since we limit our precision to a fixed, finite set of bits, \fp{} values are \emph{inherently approximate} and necessarily incur some error during computation.
+This is not bad; it is a trade-off: we get fast and compact representations that are good enough for most use cases.
+However, the downside is that \fp{} numbers \emph{do not} behave like real numbers.
+
+% FIXME: more resources!
+There are many excellent resources\cite{knuthArtComputerProgramming1997,torontoPracticallyAccurateFloatingPoint2014} on understanding the theory behind \fp{};
+we will explain the most essential aspects here briefly.
+
+A \fp value approximates real number $r \in \mathbb{R}$ by expressing $r$ in scientific notation $r = s \times f \times b^{e-q}$,
+for some base $b$,
+an integer exponent $e \in \mathbb{N}$,
+a fixed value called an \emph{excess} $q$,\footnote{The purpose of the excess $q$ is so that with value of $e > q$ we can represent positive exponents, and $e < q$ we can represent negative exponents.}
+a \emph{sign} $s \in \{-1, 1\}$,
+and a \emph{fraction} $f$ such that $0 \leq f < b$.\footnote{Another name for the fraction part commonly seen in the literature is ``mantissa'' or ``significand''.
+  You might occasionally also see ``characteristic'' referring to the exponent.
+  We will stick to ``fraction'' and ``exponent'' however.}
+
+We pack $s$, $e$, and $f$ into bit fields.
+This gives us a compact representation for both very small and very large values: because the exponent is variable, we can represent numbers as big as Avogadro's number or as small as Planck's constant with the same number of bits.
+However, since we have only a finite number of bits $p$, we have to round $f$ and the range of $e$ is limited.
+Moreover, there is some error between the real number $r$ and the value we can express with a float.
+This is why $0.1 + 0.2 = 0.30000000000000004$---$0.3$ is not precisely representable.
+
+To make matters worse, things get tricky when we do math.
+Adding e.g. Avogadro's number and Planck's constant together is beyond 16-, 32-, and 64-bit floats to represent.
+In fact, whenever the exponents differ by more than $p+2$, the result of the addition will just be the same as the larger number.
+In the case of Avogadro's number, adding any value less than $FIXME$ will just result in Avogadro's number again.
+
+The gradual accumulation of these errors and other representation issues leads to surprising results!
+% FIXME: more examples of a NaN popping up out of nowhere, e.g. with bad addition
 

### 2023-06-18 paper.tex
Fix bibliography and makefile

@@
   NaN + 42     -> NaN    # not a kill
-  \Inf{} - \Inf{}    -> NaN    # gen, but not a kill
+  Inf - Inf    -> NaN    # gen, but not a kill
   min(NaN, 42) -> NaN    # not a kill

### 2023-06-18 paper.tex
cite gpu fpx
oceananigans output table
ocean example

@@
 \newcommand{\Inf}{\code{Inf}}
+\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
+\newcommand{\zerocode}[1]{\zerowidth{\code{#1}}}
 
@@
     propagation, and death of exceptional values.
-  \item \GPUFPX{}~[CITE] provides fine-grained insights for GPU kernels.
+  \item A companion tool \GPUFPX{}~\cite{llsflg-hpdc-2023} provides fine-grained insights for GPU kernels.
 \end{itemize}
@@
 
+% * what about GPUs? companion project at Utah GPU-FPX
+% * architecture diagram, CPU + GPU, show Ocean source code
+% * replace CPU() with GPU() in code, get output like below & attached
+
 \section{Case Studies}
@@
 
-FILL
+Oceananigans~\cite{OceananigansJOSS} is simulation package for incompressible
+fluid dynamics that can generate code for Nvidia GPUs.
+For example, the following program (from the project readme) simulates turbulance:
 
-% * what about GPUs? companion project at Utah GPU-FPX
-% * architecture diagram, CPU + GPU, show Ocean source code
-% * replace CPU() with GPU() in code, get output like below & attached
-% 
-% \begin{verbatim}
-% using Oceananigans
-% grid = RectilinearGrid(GPU(), size=(128, 128), x=(0, 2π), y=(0, 2π), topology=(Periodic, Periodic, Flat))
-% model = NonhydrostaticModel(; grid, advection=WENO())
-% ϵ(x, y, z) = 2rand() - 1
-% set!(model, u=ϵ, v=ϵ)
-% simulation = Simulation(model; Δt=0.01, stop_time=4)
-% run!(simulation)
-% \end{verbatim}
-% 
-% ... xinyi ran gpu-fpx on this code ... on the binary it created?
-% 
-% 
-% - - -
-% 
-% 
-% #GPU-FPX LOC-EXCEP INFO: in kernel
-% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
-% DIV0 found @ ./float.jl:335 [FP64]
-% #GPU-FPX LOC-EXCEP INFO: in kernel
-% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
-% NaN found @ ./float.jl:335 [FP64]
-% #GPU-FPX LOC-EXCEP INFO: in kernel
-% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
-% NaN found @ ./float.jl:335 [FP32]
-% #GPU-FPX LOC-EXCEP INFO: in kernel
-% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
-% DIV0 found @ ./int.jl:261 [FP64]
-% #GPU-FPX LOC-EXCEP INFO: in kernel
-% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
-% NaN found @ ./int.jl:261 [FP64]
-% #GPU-FPX LOC-EXCEP INFO: in kernel
-% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
-% INF found @ ./int.jl:261 [FP64]
-% 
-% 
-% \begin{verbatim}
-% --- GPU-FPX Report ---
-% 
-% --- FP64 Operations ---
-% Total NaN found: 2
-% Total INF found: 1
-% Total underflow (subnormal): 0
-% Total division by zero: 2
-% --- FP32 Operations ---
-% Total NaN found: 1
-% Total INF found: 0
-% Total underflow (subnormal): 0
-% Total division by zero: 0
-% --- Other Stats ---
-% Kernels: 21
-% \end{verbatim}
+\begin{lstlisting}[language = Julia]
+using Oceananigans
+grid = RectilinearGrid(GPU(),
+  size=(128, 128), x=(0, 2π), y=(0, 2π),
+  topology=(Periodic, Periodic, Flat))
+model = NonhydrostaticModel(; grid,
+  advection=WENO())
+ϵ(x, y, z) = 2rand() - 1
+set!(model, u=ϵ, v=ϵ)
+simulation = Simulation(model;
+  Δt=0.01, stop_time=4)
+run!(simulation)
+\end{lstlisting}
+
+\GPUFPX{} provides detailed feedback on this program.
+The output in~\cref{f:gpufpx} shows that 21 kernels
+appear and generate six floating-point exceptions.
+There are three \NaN{}s, one \Inf{}, and two division
+by zero errors.
+The report is a starting point for further investigation
+of the reliability of the example.
+%% TODO exactly how?
+
+\begin{figure}[t]\centering
+  \begin{tabular}[t]{ll}
+    \begin{tabular}[t]{lr}
+      \zerocode{-{}- FP64 Operations -{}-} \\
+      \code{Total NaN:} & \code{2} \\
+      \code{Total INF:} & \code{1} \\
+      \code{Total subnormal:} & \code{0} \\
+      \code{Total div0:} & \code{2} \\
+    \end{tabular}
+    &
+    \begin{tabular}[t]{lr}
+      \zerocode{-{}- FP32 Operations -{}-} \\
+      \code{Total NaN:} & \code{1} \\
+      \code{Total INF:} & \code{0} \\
+      \code{Total subnormal:} & \code{0} \\
+      \code{Total div0:} & \code{0} \\
+    \end{tabular}
+  \end{tabular}
+  \begin{tabular}{lr}
+    \zerocode{-{}- Other Stats -{}-} \\
+    \code{Kernels:} & \code{21}
+  \end{tabular}
+
+  \caption{Example \GPUFPX{} output}
+  \label{f:gpufpx}
+\end{figure}
 

### 2023-06-23 paper.tex
ganesh note: FPtracker + GPUFPX
ieee 754 quiet nan, retrospective analysis

@@
 
+\subsubsection{TODO good quote, try to work this in}
+
+From IEEE 754. Section 6.2 here: \url{http://www.dsc.ufcg.edu.br/~cnum/modulos/Modulo2/IEEE754\_2008.pdf}:
+
+\begin{quote}
+  Quiet NaNs should, by means left to the implementer’s discretion, afford
+  retrospective diagnostic information inherited from invalid or unavailable
+  data and results.
+\end{quote}
+
+Found via \url{https://anniecherkaev.com/the-secret-life-of-nan}
+
+TODO what other tools are there for retrospective diagnosis?
+
 \subsection{Infinity}
@@
 % * replace CPU() with GPU() in code, get output like below & attached
+
+%% Ganesh
+% 3) These are conceptually usable together (Julia calling GPU kernels below).
+% The usage is simple: Start the Julia application that calls a GPU kernel under
+% LD_PRELOAD as with GPU-FPX launch. When this launch transfers control to a GPU
+% kernel, the GPU-FPX instrumentation starts notifying the exceptions on the
+% console of the CPU. When the GPU launch returns, the exceptions go in the
+% return call path of the Julia execution. Ashton and Xinyi plz validate this
+% scenario
 

### 2023-07-19 paper.tex
ganesh orcid, start rel work
more rw notes

@@
 
+Error analysis, typical answer.
+Difficult.
+CITE
+
+Toronto McCarthy lightweight plotting approach.
+No tool support.
+
+Herbie, improving accuracy automatically.
+ParetoHerbie
+
+%% http://eprints.maths.manchester.ac.uk/2873/1/fami22.pdf
+%% https://drive.google.com/drive/folders/1HkgMpxi6hsqLSZj9tSnuCxnFExPPjmvx?usp=sharing
+%% https://docs.google.com/presentation/d/1ueqzb5QY9YHmtY2DfY_hpN9oufSOcXErLWsJOZQxgzo/edit?usp=sharing
+%% https://ieeexplore.ieee.org/document/8916392
+%% https://link.springer.com/chapter/10.1007/978-3-319-73721-8_24
+
+%% Demmel, somewhere!
+%% https://arxiv.org/pdf/2207.09281.pdf
+
+Sherlogs, ShallowWaters
+%% https://github.com/milankl/Sherlogs.jl
+%% sherlogs
+
+FPSpy x86 binaries [10]
+Dinda etal hpdc 2020
+
+FPChecker, " considers CPU OpenMP and MPI codes, again using LLVM
+instrumentation"~\cite{ltlg-iiswc-2022}
+
+6-8,24,25
+Arnab Das scalable yet rigorous
+Marc Daumas Certification of Bounds Expres
+David Delmas Towards an Industrial Use of
+Alexey Solovyev FPTaylor Results table
+Titolo absint
+
+CSTG~\cite{hmbdrg-cse-2014}.
+What else???
+
+GPU-FPX state of art, companion project~\cite{llsflg-hpdc-2023}.
+Latest in line of work.
+BinFPE exn checking, slower than GPU-FPX, 
+\cite{llg-soap-2022}.
+FPChecker, early pioneer, fpx in gpus, llvm-level instrumentation
+if compiled with Clang, but most gpus use closed NVCC compiler~\cite{l-ase-2019}.
+
 \section{Discussion}

### 2023-08-01 paper.tex
quick notes + links
sec2 note: gpu-fpx has a great fp intro
Bunch of SVGs for the talk—update rough TOC for writing
Lots of work on §3

@@
 
+%% extra links:
+%% - GPU Julia: https://cuda.juliagpu.org/stable/tutorials/introduction/
+%% - GPU x2: https://nextjournal.com/sdanisch/julia-gpu-programming
+%% - NaN vs missing https://discourse.julialang.org/t/is-there-any-reason-to-use-nan-instead-of-missing/84396
+
 % If using Emacs, put cursor on last closing paren and call "eval-last-sexp" (usually `C-x C-e') to regenerate
-%\iffalse
-%(save-excursion
-%  (let* ((toc (funcall imenu-create-index-function))
-%         (headers (mapcar (pcase-lambda (`(,str . ,loc)) str)
-%                          toc)))
-%    (beginning-of-buffer)
-%    (search-forward (concat "%% " "--- autogenerated TOC ---"))
-%    (next-line)
-%    (beginning-of-line)
-%    (let ((start (point)))
-%      (search-forward (concat "%% " "--- end autogenerated TOC ---"))
-%      (beginning-of-line)
-%      (delete-region start (point))
-%      (goto-char start))
-%    (cl-loop for header in headers
-%             do (insert (format "%% %s\n" header)))
-%    (next-line)))
-%\fi
+%% (save-excursion
+%%  (let* ((toc (funcall imenu-create-index-function))
+%%         (headers (mapcar (pcase-lambda (`(,str . ,loc)) str)
+%%                          toc)))
+%%    (beginning-of-buffer)
+%%    (search-forward (concat "%% " "--- autogenerated TOC ---"))
+%%    (next-line)
+%%    (beginning-of-line)
+%%    (let ((start (point)))
+%%      (search-forward (concat "%% " "--- end autogenerated TOC ---"))
+%%      (beginning-of-line)
+%%      (delete-region start (point))
+%%      (goto-char start))
+%%    (cl-loop for header in headers
+%%             do (insert (format "%% %s\n" header)))
+%%    (next-line)))
 
 %% --- autogenerated TOC ---
-% juliacon
-% document
+% --
+% --
+% <<header.tex
 % Introduction
 % Floating Point Exception Primer
-%     Where do NaNs come from?
-%     Infinity
-%     To Kill a Floating Point
-%   \FlowFPX{
-%     \FT{
-%     Catching a NaN Killer
-%     NaN propagation monitoring
-%     NaN injection
+% . Where do \NaN{}s come from?
+% . . TODO good quote, try to work this in
+% . Infinity
+% . To Kill a Floating Point
+% \FlowFPX{}
+% . \FT{}
+% . . Catching a \NaN{} Killer
+% . . \NaN{} propagation monitoring
+% . . \NaN{} injection
+% . Visualization
+% . \GPUFPX{}
 % Case Studies
-%     Surprise NaNs: ShallowWaters
-%       CSTGs help us get a bigger picture
-%     NaN fuzzing to find a kill: OrdinaryDiffEq
-%     Finch
-%     A Bug-Hunting Safari
-%     \GPUFPX{
+% . Surprise \NaN{}s: ShallowWaters
+% . . CSTGs help us get a bigger picture
+% . \NaN{} fuzzing to find a kill: OrdinaryDiffEq
+% . Finch
+% . A Bug-Hunting Safari
+% . \GPUFPX{}: Oceananigans
 % Related Work
 % Discussion
-%     Tracking more than NaNs and Infs
-%     Enhanced Fuzzing
-%     Static FPX, Gradual Types
+% . Tracking more than \NaN{}s and \Inf{}s
+% . Enhanced Fuzzing
+% . Static FPX, Gradual Types
 % Acknowledgments
-% document
+% <<bib.tex.tex
+% --
 %% --- end autogenerated TOC ---
@@
 \label{s:background}
+
+%% TODO use the GPU-FPX paper as a guide for the background. Give a little more detail than that, but not much!
 
@@
 
-FILL intro: float tracker, viz tools, gpu
+\FlowFPX{} is a growing set of tools to aid in tracking down \fp{} exceptions.
+The two primary tools we cover in this paper are \FT{} and CSTG:
+
+\begin{itemize}
+\item \FT{}'s role is twofold: to trace the lifetime of exceptional \fp{} values and to enable inside-out fuzzing.
+\item CSTG's role is to visualize the tracing output from \FT{} in compact form.
+\end{itemize}
+
+\GPUFPX\cite{llsflg-hpdc-2023} is another component of \FlowFPX{} that can track \fp{} exceptions inside a GPU kernel.
+Please see the cited paper for more details on \GPUFPX.
+Work to use \GPUFPX in tandem with \FT{} is ongoing.
+
+%% TODO julia proposed consistent NaN handling,
+%%  do we propagate the exact NaN
+%%  https://github.com/JuliaLang/julia/issues/48523
+%%
+%% Ashton: @Ben we simply propagate whatever we got from the operation. If
+%% Julia's behavior around an operator change, that should be preserved even
+%% when using FloatTracker
 
@@
 
-\FT{} works by leveraging Julia's type-based method dispatch system to capture everything from calls to the standard library down to basic arithmetic operations.
-We do this by creating custom types \texttt{TrackedFloat16}, \texttt{TrackedFloat32}, and \texttt{TrackedFloat64} that wrap their corresponding \texttt{Float16}, \texttt{Float32} etc. counterparts.
-We then defined methods for every function in the \texttt{Base} module, including new implementations for operators like \texttt{+} and \texttt{*}.
-These new methods had to be defined for every combination of tracked/untracked arguments;
-fortunately Julia's metaprogramming capabilities make this relatively simple.
-
-The surrogate methods in place, every operation on a TrackedFloat can be intercepted.
-We have three primary goals when intercepting a float:
+\FT{} has two roles:
 
 \begin{enumerate}
-  \item Catch instances of \NaN{} kills
-  \item Monitor the propagation of \NaN{}s through the program
-  \item Optionally \emph{inject} \NaN{}s to fuzz the program under scrutiny against \NaN{}-kills
+  \item Track exceptional values
+  \item Fuzz \fp{} code from the inside-out
 \end{enumerate}
 
-% Good things to put in this section:
-% - start with the max example, perhaps?
-% - trace examples
-% - CSTGs
-%
-% Don't forget to talk about trace recording
+Building the mechanism to track exceptional values also gave us a way to inject exceptional values wherever \fp{} operations occur.
+We will examine each of these roles, how they are used, and how \FT{} realizes them.
 
-\subsubsection{Catching a \NaN{} Killer}
+%% What it does, then how it works
 
-With a \fp{} value wrapped in our \texttt{TrackedFloat} type, it is easy to monitor when a \NaN{} gets killed.
-Julia dispatches all arithmetic operations on that value to flow through FloatTracker, where we can inspect the arguments.
-Kills are relatively simple: if one or more \NaN{}s flow into a function but don't flow out (i.e. the answer isn't a \NaN{}) then that's a kill.
-For example:
+\subsubsection{Tracking Excptional Values}
+
+\FT{}'s first purpose is to monitor \fp{} operations for exceptional values.
+\FT{} interposes on every arithmetic operation and standard library call that involves a \fp{} and logs a stack trace to that point whenever an exception is generated, propagated, or killed.
+A programmer can then use these stack traces to pinpoint where an exceptional value like \NaN{} came into being or where a guard against a \NaN{} kill is needed.
+\FT{} can also log the arguments to the suspect call, so you can get \emph{normal, concrete} values to help you further discover what is going wrong.
+
+All the logs that \FT{} captures are straight from Julia's \texttt{stacktrace()} function.
+\FT{} writes these logs to file; one file for each kind of event.
+This way, if a user of \FT{} is attempting to determine where a \NaN{} or an \Inf{} came from, they can look in the file for all the gen events.
+Likewise, if the user wants to find where \NaN{}s get killed, then they can look in the kills file.
+
+While other tools like profilers might accomplish this tracking by writing extensions to the language's compiler or runtime, we were able to implement \FT{} entirely in user land.
+We created custom types \texttt{TrackedFloat16}, \texttt{TrackedFloat32}, and \texttt{TrackedFloat64} that wrap their corresponding \texttt{Float16}, \texttt{Float32}, and \texttt{Float64} counterparts.
+We then defined methods for every function in the \texttt{Base} module, including new implementations for operators like \texttt{+} and \texttt{*}.
+This lets us intercept all \fp{} operations working on our \texttt{TrackedFloat} types.
 
 \begin{lstlisting}[language = Julia]
-  NaN + 42     -> NaN    # not a kill
-  Inf - Inf    -> NaN    # gen, but not a kill
-  min(NaN, 42) -> NaN    # not a kill
-  NaN < 42     -> false  # NaN kill!
+function Base.+(x::TrackedFloat64, y::TrackedFloat64)
+  result = x.val + y.val
+  check_error(+, result, x.val, y.val)
+  TrackedFloat64(result)
+end
 \end{lstlisting}
 
-There are a few exceptions to the rule of ``\NaN{} in and no \NaN{} out $\rightarrow$ kill''---some instances that we report as being a \NaN{} kill might be instead a \emph{\NaN{}-repair}, i.e.
-where the user explicitly checks for a \NaN{} and gets some non-\NaN{} value as a result.
-Once clear instance of this is the \texttt{isnan} function.
-Calling \texttt{isnan} and getting a boolean is expected behavior, and is usually used in error handling or in \emph{explicit} handling of unexpected values.
-FloatTracker still reports these cases in the kill logs so as to be complete---the user must not blindly assume that every instance of a \NaN{} kill is bad and might be a case of \NaN{} repair.
+These new methods had to be defined for every combination of tracked/untracked arguments;
+fortunately Julia's metaprogramming abilities made it simple for us to scale without the code becoming unwieldy.
+We made a macro that first loops through all the \texttt{TrackedFloat} types that we wished to implement, and then looped through all of the operators that we wished to support.
+This approached saved us over 10x the lines a non-macro approach would have required.\footnote{We implemented overrides for 645 function/type combinations.
+At roughly 5 lines per variant, this is 3225 lines of code.
+Our implementation at time of writing weighs in at 218 lines of code, which is \emph{including} whitespace and comments.}
+This approach was inspired by \texttt{Sherlogs.jl}.\cite{kMilanklSherlogsJl2021}
 
-% FIXME: we don't (yet) talk much about NaN repair in the rest of the paper.
+With the overrides in place, we can examine the arguments to a \fp{} operation to determine if an interesting event occurred.
+Here is how we classify events around a \NaN{}:
 
-% TODO?
+\begin{tabular}{ll}
+  Arguments & Result & Event Type \\
+  \hrule
+  No \NaN{} & \NaN{} & Gen \\
+  \NaN{} & \NaN{} & Prop \\
+  \NaN{} & No \NaN{} & Kill \\
+\end{tabular}
 
-\subsubsection{\NaN{} propagation monitoring}
+The same classification rules apply to \Inf{} and would apply to any other value of interest.
 
-Remember in our first example with the \texttt{bad\_max} and \texttt{good\_max} functions how we didn't have a way to tell \emph{where} the \NaN{}s originally came from?
-That can be a tedious task.
-However, FloatTracker makes this a lot easier by watching the life cycle of a \NaN{}.
+\subsubsection{Fuzzing}
 
-% TODO
+The second role of \FT{} is to fuzz code from the inside out.
+While it is wholly possible to fuzz a program by giving it random inputs, this might not help you find \emph{all} the places where an exceptional value could crop up during a calculation.\cite{ddghlllprr-correctness-2022}
 
-% FIXME: don't forget to talk about CSTGs
+The exception tracking infrastructure that we have gives us a simple way to inject exceptional values wherever we want in our code.
+We have a global configuration system that allows us to turn on fuzzing.
+Then, each time we intercept a \fp{} operation, we check to see if we should inject a \NaN{}.\footnote{At time of writing, \FT{} only injects \NaN{}, but it is trivial to extend the set of values to inject.}
+Our full decision process is as follows:
 
-\subsubsection{\NaN{} injection}
-
-% TODO
+\begin{enumerate}
+\end{enumerate}
 
@@
 % Local Variables:
-% jinx-local-words: "CFL CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq RxInfer ShallowWaters TrackedFloat isnan"
+% jinx-local-words: "CFL CSTG's CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq RxInfer ShallowWaters TrackedFloat isnan"
 % End:

### 2023-08-02 paper.tex
Fix typos; add paragraph indent so it doesn't look awful
Write a bunch about injection decision

@@
 \usepackage{cleveref}
+\usepackage{microtype}
+% \usepackage{fontspec}
 
@@
 % If using Emacs, put cursor on last closing paren and call "eval-last-sexp" (usually `C-x C-e') to regenerate
-%% (save-excursion
-%%  (let* ((toc (funcall imenu-create-index-function))
-%%         (headers (mapcar (pcase-lambda (`(,str . ,loc)) str)
-%%                          toc)))
-%%    (beginning-of-buffer)
-%%    (search-forward (concat "%% " "--- autogenerated TOC ---"))
-%%    (next-line)
-%%    (beginning-of-line)
-%%    (let ((start (point)))
-%%      (search-forward (concat "%% " "--- end autogenerated TOC ---"))
-%%      (beginning-of-line)
-%%      (delete-region start (point))
-%%      (goto-char start))
-%%    (cl-loop for header in headers
-%%             do (insert (format "%% %s\n" header)))
-%%    (next-line)))
+% (save-excursion
+%  (let* ((toc (funcall imenu-create-index-function))
+%         (headers (mapcar (pcase-lambda (`(,str . ,loc)) str)
+%                          toc)))
+%    (beginning-of-buffer)
+%    (search-forward (concat "%% " "--- autogenerated TOC ---"))
+%    (next-line)
+%    (beginning-of-line)
+%    (let ((start (point)))
+%      (search-forward (concat "%% " "--- end autogenerated TOC ---"))
+%      (beginning-of-line)
+%      (delete-region start (point))
+%      (goto-char start))
+%    (cl-loop for header in headers
+%             do (insert (format "%% %s\n" header)))
+%    (next-line)))
 
 %% --- autogenerated TOC ---
-% --
-% --
-% <<header.tex
+% juliacon
+% document
 %   Introduction
 %   Floating Point Exception Primer
-% . Where do \NaN{}s come from?
-% . . TODO good quote, try to work this in
-% . Infinity
-% . To Kill a Floating Point
-% \FlowFPX{}
-% . \FT{}
-% . . Catching a \NaN{} Killer
-% . . \NaN{} propagation monitoring
-% . . \NaN{} injection
-% . Visualization
-% . \GPUFPX{}
+%     Where do \NaN{
+%       TODO good quote, try to work this in
+%     Infinity
+%     To Kill a Floating Point
+%   \FlowFPX{
+%     \FT{
+%       Tracking Excptional Values
+%       Fuzzing
+%     Visualization
+%     \GPUFPX{
 %   Case Studies
-% . Surprise \NaN{}s: ShallowWaters
-% . . CSTGs help us get a bigger picture
-% . \NaN{} fuzzing to find a kill: OrdinaryDiffEq
-% . Finch
-% . A Bug-Hunting Safari
-% . \GPUFPX{}: Oceananigans
+%     Surprise \NaN{
+%       CSTGs help us get a bigger picture
+%     \NaN{
+%     Finch
+%     A Bug-Hunting Safari
+%     \GPUFPX{
 %   Related Work
 %   Discussion
-% . Tracking more than \NaN{}s and \Inf{}s
-% . Enhanced Fuzzing
-% . Static FPX, Gradual Types
+%     Tracking more than \NaN{
+%     Enhanced Fuzzing
+%     Static FPX, Gradual Types
 %   Acknowledgments
-% <<bib.tex.tex
-% --
+% document
 %% --- end autogenerated TOC ---
@@
 \begin{document}
+
+% Misc typographical tweaks.
+\setlength{\parindent}{1.2em}
+% \setmonofont{JuliaMono}[Extension = .ttf, Path = ./, UprightFont = *-Regular, ItalicFont = *-Italic]
 
@@
   Reliable numerical computations are central to high-performance computing and machine learning.
-  We present FlowFPX: a Julia-based tool for tracking the onset and flow of floating-point exceptions.
-  FlowFPX's design exploits Julia's operator overloading to trace exception flows and even inject exceptions to accelerate testing.
+  We present FlowFPX: a tool set for debugging floating-point code.
+  FlowFPX is composed of two primary parts: FloatTracker, a Julia-based tool for tracking the onset and flow of floating-point exceptions; and CSTG, a stack trace visualizer.
+  FlowTracker's design exploits Julia's operator overloading to trace exception flows and even inject exceptions to accelerate testing.
   We present intuitive visualizations of summarized exception flows including how they are generated, propagated and killed, thus helping with debugging and repair.
@@
 
-\subsubsection{Tracking Excptional Values}
+\subsubsection{Tracking Exceptional Values}
 
@@
 \begin{lstlisting}[language = Julia]
-function Base.+(x::TrackedFloat64, y::TrackedFloat64)
+function +(x::TrackedFloat64, y::TrackedFloat64)
   result = x.val + y.val
@@
 
-\begin{tabular}{ll}
+\begin{center}
+  \begin{tabular}{ccc}
     Arguments & Result & Event Type \\
-  \hrule
+    \hline
     No \NaN{} & \NaN{} & Gen \\
@@
   \end{tabular}
+\end{center}
 
@@
 The exception tracking infrastructure that we have gives us a simple way to inject exceptional values wherever we want in our code.
-We have a global configuration system that allows us to turn on fuzzing.
-Then, each time we intercept a \fp{} operation, we check to see if we should inject a \NaN{}.\footnote{At time of writing, \FT{} only injects \NaN{}, but it is trivial to extend the set of values to inject.}
-Our full decision process is as follows:
+We have a global configuration system that allows us to turn on fuzzing and configure where, when, and how many \NaN{}s\footnote{At time of writing, \FT{} only injects \NaN{}, but it is trivial to extend the set of values to inject.} to inject.
+
+\FT{} exposes these configuration options:
+
+\begin{description}
+\item[\texttt{active::Bool}] Inject only if this is true.
+\item[\texttt{n\_inject::Int64}] Inject at most this many \NaN{}s.
+\item[\texttt{odds::Int64}] Inject if \texttt{rand(odds) == 1}.
+\item[\texttt{functions::Array}] Inject if we are within the dynamic extent of any of these functions.
+\item[\texttt{libraries::Array}] Inject if the top-most stack frame\footnote{After stripping off the stack frames that come from \FT{} itself.} comes from any of these libraries.
+\item[\texttt{replay::String}] Name of a file to read an injection recording from.
+\item[\texttt{record::String}] Name of a file to write injection events to for later replay.
+\end{description}
+
+Then, each time we intercept a \fp{} operation, we check to see if we should inject a \NaN{}.
+\cref{fig:injection_decision} illustrates how we decide whether or not to inject a \NaN{}, which is as follows:
 
 \begin{enumerate}
+  \item Are we replaying an injection recording? If yes, inject iff the recording says to inject.
+  \item Is injection turned on? Don't inject if no.
+  \item Are there any NaNs left for us to inject? Don't inject if we're out.
+  \item Roll a \texttt{odds}-sided die---if we get anything other than 1, don't inject.
+  \item Check the stack trace: are we within the dynamic extent of any of the functions mentioned in \texttt{functions}?
+I.e.
+if we have a set of functions that we have expressed interest in via the \texttt{functions} configuration option, do any of them show up anywhere on the stack?
+If we have some functions, but none of them are on the stack, don't inject.
+  \item Check the top-most stack frame: (modulo the stack frames incurred by \FT{} intercepting the function call) is it in the list of modules mentioned in \texttt{modules}?
+If we have some modules we're interested in, but the top frame is \emph{not} from the set of interesting modules, don't inject.
+  \item If we've made it this far, inject and decrement \texttt{n\_inject}.
 \end{enumerate}
+
+In our experience, requesting the stack trace from Julia is the most expensive part of this operation.
+Therefore, we defer all the way until step 5 to request a stack trace.
+This expensive region is show with the purple box in \cref{fig:injection_decision}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=3in]{./fig/injection_decision_white_labels.png}
+  \caption{Decision flowchart for injecting NaNs. Purple box indicates expensive checks.}
+  \label{fig:injection_decision}
+\end{figure}
 
@@
 % Local Variables:
-% jinx-local-words: "CFL CSTG's CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq RxInfer ShallowWaters TrackedFloat isnan"
+% jinx-local-words: "CFL CSTG's CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq RxInfer ShallowWaters Sherlogs TrackedFloat isnan"
 % End:

### 2023-08-02 paper.tex
Start section on CSTG; add little bit about perf in §Discussion

@@
 \subsubsection{Tracking Exceptional Values}
+\label{s:trackingexceptionalvalues}
 
@@
 Likewise, if the user wants to find where \NaN{}s get killed, then they can look in the kills file.
+
+% note: there's a section on performance in the "discussion" section at the end, so we can leave it out here I think
 
@@
 Then, each time we intercept a \fp{} operation, we check to see if we should inject a \NaN{}.
-\cref{fig:injection_decision} illustrates how we decide whether or not to inject a \NaN{}, which is as follows:
-
-\begin{enumerate}
-  \item Are we replaying an injection recording? If yes, inject iff the recording says to inject.
-  \item Is injection turned on? Don't inject if no.
-  \item Are there any NaNs left for us to inject? Don't inject if we're out.
-  \item Roll a \texttt{odds}-sided die---if we get anything other than 1, don't inject.
-  \item Check the stack trace: are we within the dynamic extent of any of the functions mentioned in \texttt{functions}?
-I.e.
-if we have a set of functions that we have expressed interest in via the \texttt{functions} configuration option, do any of them show up anywhere on the stack?
-If we have some functions, but none of them are on the stack, don't inject.
-  \item Check the top-most stack frame: (modulo the stack frames incurred by \FT{} intercepting the function call) is it in the list of modules mentioned in \texttt{modules}?
-If we have some modules we're interested in, but the top frame is \emph{not} from the set of interesting modules, don't inject.
-  \item If we've made it this far, inject and decrement \texttt{n\_inject}.
-\end{enumerate}
+First we check if we're replaying a recording: if we are, just do as the recording says.
+Otherwise, we check to make sure we have \NaN{}s remaining to inject and we roll a die to see if we should inject.
+We also check the current stack trace against the lists of functions and libraries to inject in.
+If all looks good, we inject.
 
 In our experience, requesting the stack trace from Julia is the most expensive part of this operation.
-Therefore, we defer all the way until step 5 to request a stack trace.
-This expensive region is show with the purple box in \cref{fig:injection_decision}.
+Therefore, we defer the stack checks as late as we can.
+Once the fuzzer runs out of \NaN{}s, it effectively turns off, so the rest of the fuzzing session proceeds much more quickly.
+That said, gathering stack traces still incurs a heavy overhead during logging when interesting events occur;
+we recommend limiting the number and kinds of events logged for optimal performance.
 
-\begin{figure}
-  \centering
-  \includegraphics[width=3in]{./fig/injection_decision_white_labels.png}
-  \caption{Decision flowchart for injecting NaNs. Purple box indicates expensive checks.}
-  \label{fig:injection_decision}
-\end{figure}
-
-\subsection{Visualization}
+\subsection{CSTG}
 \label{s:cstg}
 
-FILL cstg output, how to read, float tracker interface,
+Coalesced Stack Trace Graphs, or CSTGs\cite{humphreySystematicDebuggingMethods2014} provide a way to visualize large amounts of stack traces in a compact form.
+\FT{} produces
 
@@
 
+\subsection{Performance}
+\ref{s:discussion-performance}
+
+% TODO: citations!!!!
+% TODO: real numbers!!!
+\FT{} incurs significant overhead on the order of 100x slower than a non-instrumented run of the same program.
+To put this number into context, Valgrind runs with a similar level of slowdown.
+
+In addition to the cost incurred by intercepting \fp{} operations, gathering stack traces is expensive---we measured a significant slowdown on the order of 10x relative to a non-logging setup of \FT{}.
+We recommend that users of \FT{} make use of the \texttt{maxLogs} and \texttt{exclude_stacktrace} configuration options to limit the number and kind of logs gathered, and thereby reduce the number of calls to \texttt{stacktrace()}.
+
 \subsection{Tracking more than \NaN{}s and \Inf{}s}
@@
 Adding \Inf{} tracking took a few hours.
-Adding other values of interest would be issues of relatively simple engineering.
+Adding other values of interest presents no fundamental difficulties.
 
@@
 % jinx-local-words: "CFL CSTG's CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq RxInfer ShallowWaters Sherlogs TrackedFloat isnan"
+% citar-bibliography: ("./ref.bib")
 % End:

### 2023-08-04 paper.tex
Cleanup bibliography
Rough draft of section 3 complete!

@@
 %   Floating Point Exception Primer
-%     Where do \NaN{
+%     Where do \NaN
 %       TODO good quote, try to work this in
@@
 %     To Kill a Floating Point
-%   \FlowFPX{
-%     \FT{
+%   \FlowFPX
+%     \FT
 %       Tracking Excptional Values
@@
 %     Visualization
-%     \GPUFPX{
+%     \GPUFPX
 %   Case Studies
-%     Surprise \NaN{
+%     Surprise \NaN
 %       CSTGs help us get a bigger picture
-%     \NaN{
+%     \NaN
 %     Finch
 %     A Bug-Hunting Safari
-%     \GPUFPX{
+%     \GPUFPX
 %   Related Work
 %   Discussion
-%     Tracking more than \NaN{
+%     Tracking more than \NaN
 %     Enhanced Fuzzing
@@
 
+\FT{} can produce copious amounts of log files which can be challenging to sift through manually.
 Coalesced Stack Trace Graphs, or CSTGs\cite{humphreySystematicDebuggingMethods2014} provide a way to visualize large amounts of stack traces in a compact form.
-\FT{} produces
+CSTG is a stand-alone tool and was not designed with \FT{} specifically in mind.
+However, it pairs well with \FT{} and makes coming through the log files easier.
+
+CSTG takes a set of stack traces and makes a graph: for each stack trace, it creates a node for every frame, and then creates or increments an edge between two adjacent frames, as seen in \cref{fig:cstg_demo}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=3in]{./fig/cstg_static_diagram.png}
+  \caption{CSTG makes a graph by connecting adjacent stack frames}
+  \label{fig:cstg_demo}
+\end{figure}
+
+CSTG's output looks like the output seen in \cref{fig:sw_nan_cstg}, which was generated by running \FT{} on a library called ShallowWaters.
+We will go into more detail about ShallowWaters in \cref{s:sw}.
+This graph has been edited for clarity and to fit into the size constraints of this paper.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=3in]{fig/sw_nan_cstg_clean.png}
+  \caption{CSTG of \NaN{} generation logs from ShallowWaters.}
+  \label{fig:sw_nan_cstg}
+\end{figure}
+
+These graphs are meant to be read bottom-up: the bottom nodes are the roots where exceptional values manifested.
+Higher nodes in the graph show function calls progressively closer to the top-level call that spawned them.
+Thick lines and high numbers indicate flows with the highest volume leading to errors---though thin areas are still worth investigating.
+CSTG provides a map to aid in contextualizing a stack trace in a log file.
+
+\subsubsection{Graph diffing}
+
+CSTG can diff two graphs to show how the flows changed.
+For example, we took the \FT{} logs from running ShallowWaters\cite{klowerNumberFormatsError2020} \footnote{\Cref{s:sw}}, and split the logs at the 10\% mark.
+We used CSTG to diff the flow graphs between the first 10\% of the run of the program and the remaining 90\%.
+The diff can be seen in \cref{fig:cstg_diff_demo}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=3in]{./fig/cstg_diff_pretty.png}
+  \caption{Diff between the early and latter stages of ShallowWaters}
+  \label{fig:cstg_diff_demo}
+\end{figure}
+
+The green lines indicate flows that were \emph{new} in the latter 90\% of the program's run.
+The red lines indicate flows that were \emph{only in} the first 10\% of the program's run.
+What this means depends on the particular program; a domain expert can use these clues to e.g.
+find where instability started in the first part of the program, and where that instability got reinforced later on.
 
@@
 
-FILL basic idea, loose coupling, typical overhead,
+% Acknowledge that apps get offloaded to GPU; FT doesn't do that, but there's this companion tool…
+% You run it on a kernel; catches things not caught by GPUs; see Oceananigans as an example below
+% TODO: add note in future work/discussion about this as well; other languages too—not just GPU
+% FloatTracker could supply an API that wraps calls to a C library—lib is still a black box, but FT gives you a big picture
+
+Many programs offload work onto GPUs, which are no less susceptible to \fp{} exceptions than CPUs.
+In some regards they are \emph{more} susceptible because GPUs do not have the same exception-handling mechanisms that CPUs do.
+Thus, tools to catch \fp{} exceptions on the GPU are essential.
+\FT{} is constrained to CPU workloads; however, there is a companion tool called \GPUFPX{}\cite{llsflg-hpdc-2023} that can help diagnose \fp{} exceptions on the GPU.
+\GPUFPX{} can detect floating-point exceptions on the GPU that GPU code cannot do on its own.
+\GPUFPX{} reports these exceptions back to the CPU for further analysis.
+
+\FT{} and \GPUFPX{} can cover the entire computational stack, as seen in \cref{fig:ft_and_gpu_fpx}.
+Work remains to tighten the integration between \FT{} and \GPUFPX{} to reduce manual work to employ both tools for heterogeneous computation.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=3in]{./fig/float_tracker_context_boxes.png}
+  \caption{FloatTracker and GPU-FPX compliment each other}
+  \label{fig:ft_and_gpu_fpx}
+\end{figure}
+
+See \cref{s:ocean} for an example of \GPUFPX{} in action, and \ref{f:gpufpx} for an example of its output.
 
@@
 \subsection{Surprise \NaN{}s: ShallowWaters}
+\label{s:sw}
 
@@
 \begin{verbatim}
-…
 -                 FT/TrackedFloat.jl:106
@@
 run_model         ShallowWaters/run_model.jl:37
-…
+top-level
 \end{verbatim}
@@
 Here is what the two graphs look like:\footnote{As with the call stack logs, these have been edited for clarity. We only remove a few boxes in between—mostly due to space constraints in this paper.}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=3in]{fig/sw_nan_cstg_clean.png}
-  \caption{CSTG of \NaN{} generation logs from ShallowWaters.}
-  \label{fig:sw_nan_cstg}
-\end{figure}
 
@@
 \subsection{\GPUFPX{}: Oceananigans}
+\label{s:ocean}
 
@@
 
-%% Demmel, somewhere!
+Demmel, somewhere!
 %% https://arxiv.org/pdf/2207.09281.pdf
@@
 instrumentation"~\cite{ltlg-iiswc-2022}
+% FPChecker has a nice table of tools; some already mentioned here
 
@@
 
-CSTG~\cite{hmbdrg-cse-2014}.
+CSTG~\cite{humphreySystematicDebuggingMethods2014}.
 What else???
@@
 In addition to the cost incurred by intercepting \fp{} operations, gathering stack traces is expensive---we measured a significant slowdown on the order of 10x relative to a non-logging setup of \FT{}.
-We recommend that users of \FT{} make use of the \texttt{maxLogs} and \texttt{exclude_stacktrace} configuration options to limit the number and kind of logs gathered, and thereby reduce the number of calls to \texttt{stacktrace()}.
+We recommend that users of \FT{} make use of the \texttt{maxLogs} and \texttt{exclude\_stacktrace} configuration options to limit the number and kind of logs gathered, and thereby reduce the number of calls to \texttt{stacktrace()}.
 

### 2023-08-04 paper.tex
fix broken cite

@@
     To visualize results, \FT{} adapts coalesced stack-trace graphs 
-    (CSTGs, or \CSTG{}s)~\cite{hmbdrg-cse-2014} for a whole-program view on the generation,
+    (CSTGs, or \CSTG{}s)~\cite{humphreySystematicDebuggingMethods2014} for a whole-program view on the generation,
     propagation, and death of exceptional values.

### 2023-08-04 paper.tex
Cleanup section 4

@@
 
-\begin{description}
-\item[\texttt{active::Bool}] Inject only if this is true.
-\item[\texttt{n\_inject::Int64}] Inject at most this many \NaN{}s.
-\item[\texttt{odds::Int64}] Inject if \texttt{rand(odds) == 1}.
-\item[\texttt{functions::Array}] Inject if we are within the dynamic extent of any of these functions.
-\item[\texttt{libraries::Array}] Inject if the top-most stack frame\footnote{After stripping off the stack frames that come from \FT{} itself.} comes from any of these libraries.
-\item[\texttt{replay::String}] Name of a file to read an injection recording from.
-\item[\texttt{record::String}] Name of a file to write injection events to for later replay.
-\end{description}
+\begin{itemize}
+\item \textbf{\texttt{active::Bool} Inject only if this is true.
+\item \texttt{n\_inject::Int64} Inject at most this many \NaN{}s.
+\item \texttt{odds::Int64} Inject if \texttt{rand(odds) == 1}.
+\item \texttt{functions::Array} Inject if we are within the dynamic extent of any of these functions.
+\item \texttt{libraries::Array} Inject if the top-most stack frame\footnote{After stripping off the stack frames that come from \FT{} itself.} comes from any of these libraries.
+\item \texttt{replay::String} Name of a file to read an injection recording from.
+\item \texttt{record::String} Name of a file to write injection events to for later replay.
+\end{itemize}
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=3in]{fig/sw_nan_cstg_clean.png}
-  \caption{CSTG of \NaN{} generation logs from ShallowWaters.}
-  \label{fig:sw_nan_cstg}
-\end{figure}
-
 These graphs are meant to be read bottom-up: the bottom nodes are the roots where exceptional values manifested.
@@
 CSTG provides a map to aid in contextualizing a stack trace in a log file.
-
-\subsubsection{Graph diffing}
-
-CSTG can diff two graphs to show how the flows changed.
-For example, we took the \FT{} logs from running ShallowWaters\cite{klowerNumberFormatsError2020} \footnote{\Cref{s:sw}}, and split the logs at the 10\% mark.
-We used CSTG to diff the flow graphs between the first 10\% of the run of the program and the remaining 90\%.
-The diff can be seen in \cref{fig:cstg_diff_demo}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=3in]{./fig/cstg_diff_pretty.png}
-  \caption{Diff between the early and latter stages of ShallowWaters}
-  \label{fig:cstg_diff_demo}
-\end{figure}
-
-The green lines indicate flows that were \emph{new} in the latter 90\% of the program's run.
-The red lines indicate flows that were \emph{only in} the first 10\% of the program's run.
-What this means depends on the particular program; a domain expert can use these clues to e.g.
-find where instability started in the first part of the program, and where that instability got reinforced later on.
 
@@
 
-\FlowFPX{} is effective for debugging all sorts of \NaN{}-related issues.
-We have used it to reveal silent \NaN{}s, uncover the root cause of \NaN{}s,
-and fuzz for weak points in several case studies.
+We have used \FlowFPX{} to debug and harden various programs against exceptional values, which we now present in the following case studies.
 
-\subsection{Surprise \NaN{}s: ShallowWaters}
+\begin{itemize}
+\item \Cref{s:sw} uses the \FT{} and CSTG to uncover the source of \NaN{} and \Inf{} in a simulation.
+\item \Cref{s:ode} uses the fuzzing abilities of \FT{} to harden an existing library against unexpected exceptional values.
+\item \Cref{s:finch} is another case where fuzzing uncovered a hard-to-find bug.
+\item \Cref{s:ocean} showcases using \GPUFPX{} in a simulation.
+\end{itemize}
+
+Besides case studies that we ourselves did, \cref{s:safari} is an ``in the wild'' usage of \FT{} that helped researchers uncover the source of a \NaN{} in their own code.
+
+\subsection{ShallowWaters}
 \label{s:sw}
 
-Most physics simulations of this sort include a parameter called the Courant-Friedrichs-Lewy (``CFL'') parameter, which roughly describes the size of the time step to take in running the simulation.
+For our first case study, we looked at the \texttt{ShallowWaters} library~\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019,klowerLowprecisionClimateComputing2021}, which runs a simulates the flow of water over a seabed.
+This is not a general shallow waters simulation library---it is designed primarily to showcase numerical methods involving 16-bit floats.
+That said, it works for our purposes as a proving ground for FloatTracker.
+
+Many physics simulations of this sort include a parameter called the Courant-Friedrichs-Lewy (``CFL'') parameter, which roughly describes the size of the time step to take in running the simulation.
 Normally this value sits between 0 and 1---smaller values mean the simulation runs slower but with higher accuracy, whereas higher values mean the simulation takes bigger time steps and consequently runs faster, but results might be inaccurate.
@@
 
-For our first case study, we looked at the \texttt{ShallowWaters} library~\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019,klowerLowprecisionClimateComputing2021}, which runs a simulates the flow of water over a seabed.
-This is not a general shallow waters simulation library---it is designed primarily to showcase numerical methods involving 16-bit floats.
-That said, it works for our purposes as a proving ground for FloatTracker.
-We dialed the CFL parameter up to high values (as high as 10) and we started seeing \NaN{}s in the result.
-Note that the program did not crash: instead, it began producing bad output.
+We dialed the CFL parameter up to extraordinarily high values and we started seeing \NaN{}s in the result.
+\code{ShallowWaters} detected the instability, but not before producing some bad output.
 
@@
 To get a quick overview of the most common paths to problems, we can use CSTG to get an overview of what stack frames appear the most often.
-% FIXME: cite CSTG
+
+\begin{figure}
+  \centering
+  \includegraphics[width=3in]{fig/sw_nan_cstg_clean.png}
+  \caption{CSTG of \NaN{} generation logs from ShallowWaters.}
+  \label{fig:sw_nan_cstg}
+\end{figure}
 
@@
 
+Another one of CSTGs abilities is \emph{graph diffing}.
+We took the \FT{} logs from running ShallowWaters and split the logs at the 10\% mark.
+We used CSTG to diff the flow graphs between the first 10\% of the run of the program and the remaining 90\%.
+The diff can be seen in \cref{fig:cstg_diff_demo}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=3in]{./fig/cstg_diff_pretty.png}
+  \caption{Diff between the early and latter stages of ShallowWaters}
+  \label{fig:cstg_diff_demo}
+\end{figure}
+
+The green lines indicate flows that were \emph{new} in the latter 90\% of the program's run.
+The red lines indicate flows that were \emph{only in} the first 10\% of the program's run.
+What this means depends on the particular program; a domain expert can use these clues to e.g.
+find where instability started in the first part of the program, and where that instability got reinforced later on.
+
 \subsection{\NaN{} fuzzing to find a kill: OrdinaryDiffEq}
+\label{s:ode}
 
@@
 
-\subsection{A Bug-Hunting Safari}
-
-% Notes on RxInfer
-
-We searched GitHub for open issues in Julia libraries involving difficult-to-find \NaN{}s.
-In our first batch of issues, we found the \code{RxInfer.jl} library with an issue revolving around \NaN{} detection.\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}}
-We offered our tool for the package authors to use, and within a day the authors were able to track down the location of a \NaN{} and fix the issue---all with minimal guidance.
-
 \subsection{\GPUFPX{}: Oceananigans}
@@
 \end{figure}
+
+\subsection{A Bug-Hunting Safari}
+\ref{s:safari}
+
+\FT{} has proven to be useful for more than its creators---we discovered an open issue in the \code{RxInfer.jl} library with an issue revolving around \NaN{} detection.\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}}
+Even though we were unable to provide hands-on guidance as the code in question was proprietary, the authors of \code{RxInfer.jl} were able use \FT{} themselves, and within a day they tracked down the location of a \NaN{} gen.
+The \code{RxInfer.jl} authors provided us with valuable feedback on \FT{}'s API.
 

### 2023-08-08 paper.tex
Clean up some test cases; start work on related work section

@@
 \usepackage{microtype}
-% \usepackage{fontspec}
 
@@
 \begin{itemize}
-\item \textbf{\texttt{active::Bool} Inject only if this is true.
+\item \texttt{active::Bool} Inject only if this is true.
 \item \texttt{n\_inject::Int64} Inject at most this many \NaN{}s.
@@
 
-\subsection{\NaN{} fuzzing to find a kill: OrdinaryDiffEq}
+\subsection{OrdinaryDiffEq}
 \label{s:ode}
 
-The second thing we evaluated was an N-Body simulation library.
-We didn't find any sensible configuration of input parameters that lead to \NaN{} kills like in the ShallowWaters library.
-We turned to the \NaN{} injection capabilities of FloatTracker to fuzz the library to find any lurking bugs.
-% FIXME: add hyper-ref to § 4.1: ShallowWaters
+To exercise the fuzzing abilities of \FT{}, we tried fuzzing \texttt{NBodySimulator}\footnote{\url{https://github.com/SciML/NBodySimulator.jl}} from the SciML team, which simulates the gravitational interactions between two or more massive bodies.
+However, when we used \FT{}, we got no \fp{} operations.
+We discovered that \texttt{NBodySimulator} merely sets up the system of equations to solve, then hands all the work of running the simulation off to the \texttt{OrdinaryDiffEq}\footnote{\url{https://github.com/SciML/OrdinaryDiffEq.jl}} library from the same team.
 
-We configured FloatTracker to inject a single \NaN{} to see if we could find any \NaN{} kills.
-On some runs we noticed that we would get a kill and the program would warn about a \NaN{} and prepare to exit.
-Contrary to the error message's claim, however, the program went into an infinite loop.
+We configured \FT{} to inject a single \NaN{} somewhere within the scope of the \texttt{NBodySimulator} and \texttt{OrdinaryDiffEq} libraries.
+We found a run where something curious happened: the library reported that it had detected a \NaN{} and would then exit.
+However, after printing that message, the program went into an infinite loop.
 
-The logs lead us to a routine in the widely-used \texttt{OrdinaryDiffEq} library.
-During initialization, a \NaN{} got injected during the execution of \texttt{-} between two \texttt{TrackedFloat}s:
-
-\begin{verbatim}
-…
--            at FloatTracker/src/TrackedFloat.jl:89
-#__init#628  at OrdinaryDiffEq/src/solve.jl:106
-…
-\end{verbatim}
-
-That injection point is here on third line of this snippet, which corresponds to line 106 from the \texttt{solve.jl} file in the \texttt{OrdinaryDiffEq} library:
-
-\begin{lstlisting}[language = Julia]
-tType = eltype(prob.tspan)
-tspan = prob.tspan
-tdir = sign(tspan[end] - tspan[1])
-
-t = tspan[1]
-\end{lstlisting}
-
-\texttt{sign} correctly propagates \NaN{}s, so \texttt{tdir} now contains a \NaN{}.
-
-We \emph{could} trace this variable through the code, but that would be a lot of grunt work.
-Instead, we can look at the \NaN{} kill logs to get a clue as to where we should look next:
+% FIXME: should I show that CSTG here in the paper?
+We used CSTGs to get an idea of what the important flows were in the run of this errant behavior.
+With an idea of what was going on, we took a look at the logs and found a \NaN{} kill inside the \texttt{solve.jl} file in the \texttt{OrdinaryDiffEq} library.
 
@@
 
-We get a very large kill file from this run, and the two lines from above show up repeatedly.
-Thus, this is a good candidate location for where the cause of the loop is.
-
 The relevant part of \texttt{solve.jl} looks like this:
 
-% Note for Ashton:
 % file here: ~/Research/ode_debug/dev/OrdinaryDiffEq/src/solve.jl
@@
     # do integration work
+    # pop_off(time_stops)
   end
-  pop_if_work_done(time_stops)
 end
@@
 The problem here is on line 2 when \texttt{tdir} is \NaN{}: multiplication propagates \NaN{}s, and comparison kills it, so the condition on the inner \texttt{while} loop is \emph{always} false.
-No work got done in the inner loop, and so the conditional \texttt{pop} routine never reduced the size of the \texttt{time\_stops} vector.
+No work got done in the inner loop, and so the size of the \texttt{time\_stops} vector never decreases.
 
@@
 More dangerous cases can occur when the influence of a bad comparison is not as readily observable.
+In any case, \FT{} can help monitor for kills and help fuzz to find kills before they arise in production.
+
+We were able to use the injection recording and the gen logs to track down where the first \NaN{} was injected to cause the infinite loop.
+This made it simple for us to create an example to reproduce the behavior, which we included in the GitHub issue that we opened.\footnote{\url{https://github.com/SciML/OrdinaryDiffEq.jl/issues/1939}}
 
@@
 
-Toronto McCarthy lightweight plotting approach.
-No tool support.
+Toronto and McCarthy\cite{torontoPracticallyAccurateFloatingPoint2014} developed a methodology for assessing where arithmetic operations are liable to magnify \fp{} error.
+They do not provide any tooling to find \emph{where} the error is causing problems, but they do provide guidance on what to do to mitigate error.
 
-Herbie, improving accuracy automatically.
-ParetoHerbie
+Herbie\cite{panchekhaAutomaticallyImprovingAccuracy2015} is a tool for automatically rewriting arithmetic expressions to reduce \fp{} error.
+\FT{} tracks down exceptional \fp{} values; once the location of a gen is found, Herbie could be employed to rewrite the arithmetic expression, if \fp{} error is the cause of the exceptional value.
+Odyssey\cite{misbackOdysseyInteractiveWorkbench2023} provides a helpful interactive interface to the tooling that Herbie provides.
+
+% Herbie, improving accuracy automatically.
+% ParetoHerbie
 
@@
 
-Demmel, somewhere!
-%% https://arxiv.org/pdf/2207.09281.pdf
+Demmel et al.\cite{ddghlllprr-correctness-2022} examine \fp{} exceptional value handling in the BLAS and LAPACK libraries and propose solutions to inconsistencies.
+They also suggest a fuzzing technique that we build upon with \FT{}:
+instead of fuzzing the \emph{inputs} to a program, we fuzz by randomly injecting exceptional values at \emph{any} point where there is an arithmetic operation.
+Demmel suggests that this approach is useful since any \fp{} operation could lead to an exceptional value if the inputs are in dangerous ranges.
 
-Sherlogs, ShallowWaters
-%% https://github.com/milankl/Sherlogs.jl
-%% sherlogs
+\texttt{Sherlogs}\footnote{\url{https://github.com/milankl/Sherlogs.jl}} by Milan Klöwer inspired the use of a custom number type to intercept operations on a number.
+In contrast to \FT{} which monitors for exceptional values and logs stack traces at interesting points in their lifetime, \texttt{Sherlogs} tracks and reports the range of values seen over the course of a computation.
+This is intended to provide insight into whether or not a library could tolerate a lower-precision \fp{} format.
 

### 2023-08-09 paper.tex
Lots more work in related work section

@@
 
-Error analysis, typical answer.
-Difficult.
-CITE
+\subsection{Floating-point error analysis}
+
+Demmel et al.\cite{ddghlllprr-correctness-2022} examine \fp{} exceptional value handling in the BLAS and LAPACK libraries and propose solutions to inconsistencies.
+Their solution includes a characterization equivalent to our gen-prop-kill framework. % pg. 18---good way to cite this page specifically?
+They also suggest a fuzzing technique that we realize in \FT{}:
+instead of fuzzing the \emph{inputs} to a program, we fuzz by randomly injecting exceptional values at \emph{any} point where there is an arithmetic operation.
+Demmel suggests that this approach is useful since any \fp{} operation could lead to an exceptional value if the inputs are in dangerous ranges.
 
 Toronto and McCarthy\cite{torontoPracticallyAccurateFloatingPoint2014} developed a methodology for assessing where arithmetic operations are liable to magnify \fp{} error.
-They do not provide any tooling to find \emph{where} the error is causing problems, but they do provide guidance on what to do to mitigate error.
+Their work includes error visualization and work on how to avoid error-magnifying ``badlands''.
+They focus on rewriting arithmetic expressions to avoid error which could lead to exceptional values, rather than on exceptional values themselves.
 
-Herbie\cite{panchekhaAutomaticallyImprovingAccuracy2015} is a tool for automatically rewriting arithmetic expressions to reduce \fp{} error.
+In a similar vein, Herbie\cite{panchekhaAutomaticallyImprovingAccuracy2015} is a tool for automatically rewriting arithmetic expressions to reduce \fp{} error.
 \FT{} tracks down exceptional \fp{} values; once the location of a gen is found, Herbie could be employed to rewrite the arithmetic expression, if \fp{} error is the cause of the exceptional value.
@@
 
-% Herbie, improving accuracy automatically.
-% ParetoHerbie
+\subsection{Diagnosing floating-point exceptions}
 
-%% http://eprints.maths.manchester.ac.uk/2873/1/fami22.pdf
-%% https://drive.google.com/drive/folders/1HkgMpxi6hsqLSZj9tSnuCxnFExPPjmvx?usp=sharing
-%% https://docs.google.com/presentation/d/1ueqzb5QY9YHmtY2DfY_hpN9oufSOcXErLWsJOZQxgzo/edit?usp=sharing
-%% https://ieeexplore.ieee.org/document/8916392
-%% https://link.springer.com/chapter/10.1007/978-3-319-73721-8_24
-
-Demmel et al.\cite{ddghlllprr-correctness-2022} examine \fp{} exceptional value handling in the BLAS and LAPACK libraries and propose solutions to inconsistencies.
-They also suggest a fuzzing technique that we build upon with \FT{}:
-instead of fuzzing the \emph{inputs} to a program, we fuzz by randomly injecting exceptional values at \emph{any} point where there is an arithmetic operation.
-Demmel suggests that this approach is useful since any \fp{} operation could lead to an exceptional value if the inputs are in dangerous ranges.
-
-\texttt{Sherlogs}\footnote{\url{https://github.com/milankl/Sherlogs.jl}} by Milan Klöwer inspired the use of a custom number type to intercept operations on a number.
+The Julia library \texttt{Sherlogs}\footnote{\url{https://github.com/milankl/Sherlogs.jl}} by Milan Klöwer inspired the use of a custom number type to intercept operations on a number.
 In contrast to \FT{} which monitors for exceptional values and logs stack traces at interesting points in their lifetime, \texttt{Sherlogs} tracks and reports the range of values seen over the course of a computation.
@@
 
-FPSpy x86 binaries [10]
-Dinda etal hpdc 2020
+% TODO: make sure that what I said about what FPSpy can track is true.
+FPSpy~\cite{dindaSpyingFloatingPoint2020} is an \texttt{LD_PRELOAD} shared library that works on unmodified x86 binaries.
+It can monitor a program during execution for any \fp{} arithmetic issues, such as division by zero, denormalization, underflow, and overflow.
+\FT{} in contrast focuses on the lifetime of exceptional values; \FT{} can notice when a \NaN{} gets killed, for instance, while that is not an event that FPSpy appears to be able to track.
+FPSpy has the advantage of being lightweight enough to run on production code for certain loads.
 
-FPChecker, " considers CPU OpenMP and MPI codes, again using LLVM
-instrumentation"~\cite{ltlg-iiswc-2022}
-% FPChecker has a nice table of tools; some already mentioned here
+\subsection{Stack trace graphs}
 
-6-8,24,25
-Arnab Das scalable yet rigorous
-Marc Daumas Certification of Bounds Expres
-David Delmas Towards an Industrial Use of
-Alexey Solovyev FPTaylor Results table
-Titolo absint
+CSTG~\cite{humphreySystematicDebuggingMethods2014} is built on the work of and related to the STAT~\cite{arnoldStackTraceAnalysis2007} tool developed by Laurence Livermore National Laboratory.
+Please see the CSTG paper for more on the differences between the two.
 
-CSTG~\cite{humphreySystematicDebuggingMethods2014}.
-What else???
+\subsection{Floating-point exceptions on the GPU}
 
-GPU-FPX state of art, companion project~\cite{llsflg-hpdc-2023}.
-Latest in line of work.
-BinFPE exn checking, slower than GPU-FPX, 
-\cite{llg-soap-2022}.
-FPChecker, early pioneer, fpx in gpus, llvm-level instrumentation
-if compiled with Clang, but most gpus use closed NVCC compiler~\cite{l-ase-2019}.
+\GPUFPX{}…
+
+% Error analysis, typical answer.
+% Difficult.
+% CITE
+
+% % Herbie, improving accuracy automatically.
+% % ParetoHerbie
+
+% %% http://eprints.maths.manchester.ac.uk/2873/1/fami22.pdf
+% %% https://drive.google.com/drive/folders/1HkgMpxi6hsqLSZj9tSnuCxnFExPPjmvx?usp=sharing
+% %% https://docs.google.com/presentation/d/1ueqzb5QY9YHmtY2DfY_hpN9oufSOcXErLWsJOZQxgzo/edit?usp=sharing
+% %% https://ieeexplore.ieee.org/document/8916392
+% %% https://link.springer.com/chapter/10.1007/978-3-319-73721-8_24
+
+% FPChecker, " considers CPU OpenMP and MPI codes, again using LLVM
+% instrumentation"~\cite{ltlg-iiswc-2022}
+% % FPChecker has a nice table of tools; some already mentioned here
+
+% 6-8,24,25
+% Arnab Das scalable yet rigorous
+% Marc Daumas Certification of Bounds Expres
+% David Delmas Towards an Industrial Use of
+% Alexey Solovyev FPTaylor Results table
+% Titolo absint
+
+% GPU-FPX state of art, companion project~\cite{llsflg-hpdc-2023}.
+% Latest in line of work.
+% BinFPE exn checking, slower than GPU-FPX, 
+% \cite{llg-soap-2022}.
+% FPChecker, early pioneer, fpx in gpus, llvm-level instrumentation
+% if compiled with Clang, but most gpus use closed NVCC compiler~\cite{l-ase-2019}.
 
@@
 % Local Variables:
-% jinx-local-words: "CFL CSTG's CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq RxInfer ShallowWaters Sherlogs TrackedFloat isnan"
+% jinx-local-words: "CFL CSTG's CSTGs Courant Demmel Friedrichs Herbie JuliaHub Klöwer Lewy OrdinaryDiffEq RxInfer ShallowWaters Sherlogs TrackedFloat isnan"
 % citar-bibliography: ("./ref.bib")

### 2023-08-09 paper.tex
Improve intro to case studies

@@
 
-We have used \FlowFPX{} to debug and harden various programs against exceptional values, which we now present in the following case studies.
+In this section we'll discuss some case studies where tools from \FlowFPX{} aided in finding or fixing \fp{} issues.
+Our case studies are as follows:
 
 \begin{itemize}
-\item \Cref{s:sw} uses the \FT{} and CSTG to uncover the source of \NaN{} and \Inf{} in a simulation.
-\item \Cref{s:ode} uses the fuzzing abilities of \FT{} to harden an existing library against unexpected exceptional values.
-\item \Cref{s:finch} is another case where fuzzing uncovered a hard-to-find bug.
-\item \Cref{s:ocean} showcases using \GPUFPX{} in a simulation.
+\item \Cref{s:sw} covers a simulation that produces \NaN{}s in the output---a common and frustrating occurrence for many a scientist.
+\item \Cref{s:ode} puts a widely-used mathematics library under a fuzzer to harden it against bad behavior in the presence of exceptional \fp{} values.
+\item \Cref{s:finch} puts another program under the fuzzer.  % <- ugh, rewrite
+\item \Cref{s:ocean} is a simulation that runs on the GPU---this puts it out of reach of \FT{} but not for other components of the \FlowFPX{} tool set.
 \end{itemize}
@@
 % TODO: make sure that what I said about what FPSpy can track is true.
-FPSpy~\cite{dindaSpyingFloatingPoint2020} is an \texttt{LD_PRELOAD} shared library that works on unmodified x86 binaries.
+FPSpy~\cite{dindaSpyingFloatingPoint2020} is an \texttt{LD\_PRELOAD} shared library that works on unmodified x86 binaries.
 It can monitor a program during execution for any \fp{} arithmetic issues, such as division by zero, denormalization, underflow, and overflow.
@@
 \subsection{Performance}
-\ref{s:discussion-performance}
+\label{s:discussion-performance}
 

### 2023-08-10 paper.tex
Fix ref
More related work

@@
 \subsection{A Bug-Hunting Safari}
-\ref{s:safari}
+\label{s:safari}
 
@@
 Demmel et al.\cite{ddghlllprr-correctness-2022} examine \fp{} exceptional value handling in the BLAS and LAPACK libraries and propose solutions to inconsistencies.
-Their solution includes a characterization equivalent to our gen-prop-kill framework. % pg. 18---good way to cite this page specifically?
+They describe a proposal to pack numeric codes into an \texttt{INFO\_ARRAY} that mirrors our gen-prop-kill framework.
 They also suggest a fuzzing technique that we realize in \FT{}:
@@
 
-\GPUFPX{}…
+FPChecker~\cite{l-ase-2019} is a tool to report \fp{} exceptions occurring on the GPU.
+FPChecker relies on LLVM-level instrumentation of GPU kernels, and so cannot run on the plethora of closed-source GPU kernels in usage today.
+
+BinFPE~\cite{llg-soap-2022} is another tool in the same space;
+BinFPE performs SASS-level analysis of GPU kernels, but is limited in that it is slow and does not catch errors that alter control flow.
+The latter deficiency is particularly worrisome, as we have seen, silent \NaN{} kills can invalidate results without the user noticing.
+
+\GPUFPX{}~\cite{llsflg-hpdc-2023} improves on the work of FPChecker and BinFPE by being more performant and catching a wider set of errors, including those that alter control flow.
+\GPUFPX{} is a shared library that, like FPSpy uses \texttt{LD\_PRELOAD} to work on unmodified binaries.
+\GPUFPX{} runs on CUDA cores from NVIDIA and reports total numbers of exceptional values.
+Like \FT{}, \GPUFPX{} can catch \NaN{} gens and kills, but it does this on the GPU where \FT{} doesn't apply.
+Despite these improvements, \GPUFPX{} is limited by the closed-source nature of common GPU cores, and cannot report at the rich level of source detail that \FT{} can.
+
+Work is ongoing to use \GPUFPX{} and \FT{} in tandem.
 
@@
 % Local Variables:
-% jinx-local-words: "CFL CSTG's CSTGs Courant Demmel Friedrichs Herbie JuliaHub Klöwer Lewy OrdinaryDiffEq RxInfer ShallowWaters Sherlogs TrackedFloat isnan"
+% jinx-local-words: "BinFPE CFL CSTG's CSTGs Courant Demmel FPChecker Friedrichs Herbie JuliaHub Klöwer Lewy OrdinaryDiffEq RxInfer ShallowWaters Sherlogs TrackedFloat isnan"
 % citar-bibliography: ("./ref.bib")

### 2023-08-11 paper.tex
Major rework of section 2

@@
 
-%% TODO use the GPU-FPX paper as a guide for the background. Give a little more detail than that, but not much!
+\Fp{} provides a finite approximation of the real numbers in a format that is efficient for calculation on a computer.
+IEEE 754~\cite{IEEEStandardBinary1985} standardizes \fp{} representation and operations---these days, most machines and languages follow the IEEE 754 standard, and we make no effort to generalize our work beyond what IEEE 754 specifies.
 
-% TODO: talk about how floating point != real numbers → errors appear and build up
-% (see other tutorials; Art of Computer Programming, Ganesh, slides in Google Drive from guy in England)
-% 2 paragraphs or so: floating point primer
+\Fp{} is essentially scientific notation---\fp{} numbers are represented by packing a sign bit, an exponent, and an fraction part (also called the ``significand'' or ``mantissa'') into contiguous bit fields.
+The advantages of this representation are that allows for compact representation of both very small and very large numbers in a narrow range of bits---a 64-bit float can represent numbers as big as $2^{1024}$ and as small as $2^{-1074}$.
+This range is sufficient to represent both Planck's constant as well as Avogadro's constant with reasonable accuracy, making 
+The flip side is that not every real number within that wide range is representable as a \fp{} value---there is almost always some gap between the real number that we wish to represent and the \fp{} number that can represent.
+Whenever a value we have doesn't have a precise representation, we must \emph{round} that real value to the nearest \fp{} value.
+We call this gap error, and error, along with the finite nature of \fp{} values, means that we can run into some funny situations.
 
-\Fp{} numbers are a way of representing real numbers in a computer.
-However, since we limit our precision to a fixed, finite set of bits, \fp{} values are \emph{inherently approximate} and necessarily incur some error during computation.
-This is not bad; it is a trade-off: we get fast and compact representations that are good enough for most use cases.
-However, the downside is that \fp{} numbers \emph{do not} behave like real numbers.
+For example, if two numbers' exponents differ by too large an amount, adding or subtracting the smaller from the bigger can result in just the bigger number.
+For example, $6.022 \times 10^{23} + 6.626 \times 10^{-34}$ (Avogadro's constant plus Planck's constant) will leave with with just Avogadro's constant---there is no way to pack all the significant figures from true value that differs minutely from Avogadro's number into a 64-bit \fp{} number.
+This is a fairly extreme example; even when the numbers are closer in magnitude, most operations on \fp{} numbers induce loss of accuracy.
 
-% FIXME: more resources!
-There are many excellent resources\cite{knuthArtComputerProgramming1997,torontoPracticallyAccurateFloatingPoint2014} on understanding the theory behind \fp{};
-we will explain the most essential aspects here briefly.
+\subsection{Exceptions and exceptional values}
 
-A \fp value approximates real number $r \in \mathbb{R}$ by expressing $r$ in scientific notation $r = s \times f \times b^{e-q}$,
-for some base $b$,
-an integer exponent $e \in \mathbb{N}$,
-a fixed value called an \emph{excess} $q$,\footnote{The purpose of the excess $q$ is so that with value of $e > q$ we can represent positive exponents, and $e < q$ we can represent negative exponents.}
-a \emph{sign} $s \in \{-1, 1\}$,
-and a \emph{fraction} $f$ such that $0 \leq f < b$.\footnote{Another name for the fraction part commonly seen in the literature is ``mantissa'' or ``significand''.
-  You might occasionally also see ``characteristic'' referring to the exponent.
-  We will stick to ``fraction'' and ``exponent'' however.}
+IEEE 754 defines \emph{\fp{} exceptions}, which are events that occur when something goes wrong in a calculation.
+William Kahan says,
 
-We pack $s$, $e$, and $f$ into bit fields.
-This gives us a compact representation for both very small and very large values: because the exponent is variable, we can represent numbers as big as Avogadro's number or as small as Planck's constant with the same number of bits.
-However, since we have only a finite number of bits $p$, we have to round $f$ and the range of $e$ is limited.
-Moreover, there is some error between the real number $r$ and the value we can express with a float.
-This is why $0.1 + 0.2 = 0.30000000000000004$---$0.3$ is not precisely representable.
+\begin{quote}
+An arithmetic exception arises when [there is] no result that would be acceptable universally.
+\end{quote}
 
-To make matters worse, things get tricky when we do math.
-Adding e.g. Avogadro's number and Planck's constant together is beyond 16-, 32-, and 64-bit floats to represent.
-In fact, whenever the exponents differ by more than $p+2$, the result of the addition will just be the same as the larger number.
-In the case of Avogadro's number, adding any value less than $FIXME$ will just result in Avogadro's number again.
+For example, dividing by zero is an exception, as is exceeding your \fp{} format's range.
+These are exceptions because it is impossible to say what the right answer \emph{should} have been: perhaps that zero was meant to be a value that got really \emph{close} to zero, and suddenly rounded towards zero before division.
+Or maybe that value that got really big \emph{does} need to get that big, but you need to start using a larger format for your \fp{} values.
+Whatever, the case, there isn't a universally acceptable answer, so a \fp{} exception gets raised.
 
-The gradual accumulation of these errors and other representation issues leads to surprising results!
-% FIXME: more examples of a NaN popping up out of nowhere, e.g. with bad addition
+IEEE 754 doesn't always obligate the programmer to handle these exceptions, however, so we end up with \emph{exceptional values}, which are the results of exception-raising operations.
+The two values we work with in this paper are \Inf{} and \NaN{}, which result from overflow and nonsensical operations respectively.\footnote{The third kind of exceptional value is called a \emph{subnormal} value.
+Please see~\cite{torontoPracticallyAccurateFloatingPoint2014} for a good introduction to subnormal values, as well as other details of \fp{} representation.}
 
-In our work we focused on two of the most common problematic values in the \fp{} space: \NaN{} and \Inf{}.
-We leave the possibility of tracking other kinds of troublesome values (e.g. subnormal numbers) open, but for our work we restrict ourselves to just these values.
-
-\subsection{Where do \NaN{}s come from?}
-
-IEEE 754~\cite{IEEEStandardBinary1985} lists the following operations (among a few others) where \NaN{}s result:
+IEEE 754 lists the following operations (among a few others) where \NaN{}s result:
 
@@
 
-When one of these events happens, we call it a \NaN{} \emph{generation}.
-
-Most mathematical operations preserve \NaN{}s.
-We call these instances \NaN{} \emph{propagation}, as the \NaN{} is propagated from the source in the arguments to the result.
-
-There are, however, some instances where the IEEE standard requires that \NaN{}s to \emph{not} propagate from arguments to the result; we call these \NaN{} \emph{kills}.
-Sometimes this can be good---an \texttt{isnan} check should clearly return a boolean and not a \NaN{}---but kills are often harmful in two ways:
-
-\begin{enumerate}
-  \item Users no longer see that there was some errant operation: \NaN{}s are an opportunity to closely examine data for problems and code for unsoundness.
-  \item \NaN{} kills can influence the control of a program leading to incorrect results, as we saw in~\cref{s:to-kill-a-fp}.
-\end{enumerate}
-
-Here are cases where, according to the IEEE standard, a \NaN{} can be killed:
-
-\begin{align*}
-x \; \texttt{cmp} \; y &\rightarrow false &\text{where \texttt{cmp} is \texttt{<}, \texttt{>}, \texttt{<=}, etc.} \\
-1^{\text{\texttt{\NaN{}}}} &\rightarrow 1 \\
-\text{\texttt{\NaN{}}}^{0} &\rightarrow 1
-\end{align*}
-
-FloatTracker monitors these and all other operations for interesting events.
-
-\subsubsection{TODO good quote, try to work this in}
-
-From IEEE 754. Section 6.2 here: \url{http://www.dsc.ufcg.edu.br/~cnum/modulos/Modulo2/IEEE754\_2008.pdf}:
-
-\begin{quote}
-  Quiet NaNs should, by means left to the implementer’s discretion, afford
-  retrospective diagnostic information inherited from invalid or unavailable
-  data and results.
-\end{quote}
-
-Found via \url{https://anniecherkaev.com/the-secret-life-of-nan}
-
-TODO what other tools are there for retrospective diagnosis?
-
-\subsection{Infinity}
-
-%%ben: shouldn't this come before NaN b/c infs lead to NaNs?
-
-There are two infinities, $+\infty$ and $-\infty$, but they behave the same for our purposes, so we will refer to them just as ``\Inf{}''.
+There are actually two infinities, $+\infty$ and $-\infty$, but they behave the same for our purposes, so we will refer to them just as ``\Inf{}''.
 Like \NaN{}s, \Inf{} can crop up when bad math happens: $\frac{x}{0} \rightarrow \infty$ when $x \ne 0$.
 \Inf{} can also arise when numbers get too big or too small for the choice of representation.
+
+There are many excellent resources\cite{knuthArtComputerProgramming1997,torontoPracticallyAccurateFloatingPoint2014} on understanding the theory behind \fp{} as well as how error is measured and accumulates.
+Please see those works for more details.
+What we have described above should suffice for the remainder of this paper.
 
@@
 % If anyone has a better "kill" pun/reference than "To Kill a Mockingbird", I'm all ears.
+
+Exceptional values have a life cycle: they are born, live, and die, as seen in \cref{f:gpk}.
 
@@
 
+When a \NaN{} or an \Inf{} comes into being as the result of a bad math operation, we call this a generation, or \emph{gen} for short.
+
 The IEEE 754 specification for \fp{} numbers governs how \NaN{}s are handled as arguments to arithmetic functions~\cite{IEEEStandardBinary1985}.
 Most of the time, when any or all of the arguments are \NaN{}, the result is also a \NaN{}.
+We call this a propagation, or \emph{prop} for short.
+
 This is good and desirable, because it means that an arithmetic error---even if buried deep in an expression---will still be observable, allowing a user to find and debug it.
 
-However, there are a few places where \NaN{}s flow into an expression, but not out.
-We call these instances ``\NaN{}-kills'', and these can lead to subtle logic bugs.
-Consider the following example:
+However, there are a few places where exceptional values flow into an expression, but not out.
+We call these instances \emph{kills}, and these can lead to subtle logic bugs.
+For example, \emph{any} comparison ($<$, $>$, $=$, $\neq$, etc.) with a \NaN{} returns \texttt{false}.
+
+Consider the following two functions to find the maximum value in a list:
 

### 2023-08-11 paper.tex
Finish sentence

@@
 The advantages of this representation are that allows for compact representation of both very small and very large numbers in a narrow range of bits---a 64-bit float can represent numbers as big as $2^{1024}$ and as small as $2^{-1074}$.
-This range is sufficient to represent both Planck's constant as well as Avogadro's constant with reasonable accuracy, making 
+
+This range is sufficient to represent both Planck's constant as well as Avogadro's constant with reasonable accuracy, making IEEE 754 \fp{} adequate for a wide range of numerical computations.
 The flip side is that not every real number within that wide range is representable as a \fp{} value---there is almost always some gap between the real number that we wish to represent and the \fp{} number that can represent.

### 2023-08-15 paper.tex
Use Julia for examples of floating-point inaccuracy

@@
 %% - NaN vs missing https://discourse.julialang.org/t/is-there-any-reason-to-use-nan-instead-of-missing/84396
-
-% If using Emacs, put cursor on last closing paren and call "eval-last-sexp" (usually `C-x C-e') to regenerate
-% (save-excursion
-%  (let* ((toc (funcall imenu-create-index-function))
-%         (headers (mapcar (pcase-lambda (`(,str . ,loc)) str)
-%                          toc)))
-%    (beginning-of-buffer)
-%    (search-forward (concat "%% " "--- autogenerated TOC ---"))
-%    (next-line)
-%    (beginning-of-line)
-%    (let ((start (point)))
-%      (search-forward (concat "%% " "--- end autogenerated TOC ---"))
-%      (beginning-of-line)
-%      (delete-region start (point))
-%      (goto-char start))
-%    (cl-loop for header in headers
-%             do (insert (format "%% %s\n" header)))
-%    (next-line)))
-
-%% --- autogenerated TOC ---
-% juliacon
-% document
-%   Introduction
-%   Floating Point Exception Primer
-%     Where do \NaN
-%       TODO good quote, try to work this in
-%     Infinity
-%     To Kill a Floating Point
-%   \FlowFPX
-%     \FT
-%       Tracking Excptional Values
-%       Fuzzing
-%     Visualization
-%     \GPUFPX
-%   Case Studies
-%     Surprise \NaN
-%       CSTGs help us get a bigger picture
-%     \NaN
-%     Finch
-%     A Bug-Hunting Safari
-%     \GPUFPX
-%   Related Work
-%   Discussion
-%     Tracking more than \NaN
-%     Enhanced Fuzzing
-%     Static FPX, Gradual Types
-%   Acknowledgments
-% document
-%% --- end autogenerated TOC ---
 
@@
 The advantages of this representation are that allows for compact representation of both very small and very large numbers in a narrow range of bits---a 64-bit float can represent numbers as big as $2^{1024}$ and as small as $2^{-1074}$.
+This range is sufficient to represent numbers as small as Planck's constant or as big as Avogadro's constant with reasonable accuracy.
 
-This range is sufficient to represent both Planck's constant as well as Avogadro's constant with reasonable accuracy, making IEEE 754 \fp{} adequate for a wide range of numerical computations.
 The flip side is that not every real number within that wide range is representable as a \fp{} value---there is almost always some gap between the real number that we wish to represent and the \fp{} number that can represent.
@@
 For example, if two numbers' exponents differ by too large an amount, adding or subtracting the smaller from the bigger can result in just the bigger number.
-For example, $6.022 \times 10^{23} + 6.626 \times 10^{-34}$ (Avogadro's constant plus Planck's constant) will leave with with just Avogadro's constant---there is no way to pack all the significant figures from true value that differs minutely from Avogadro's number into a 64-bit \fp{} number.
+Below we take the Planck constant and add it to Avogadro's constant and are left bit-for-bit with Avogadro's constant, as if we had never done any addition at all: there is no way to pack all the significant figures from true value that differs minutely from Avogadro's number into a 64-bit \fp{} number.
+
+\begin{lstlisting}[language = Julia]
+p = 6.62607015e-34
+a = 6.02214076e23
+
+println("planck:   $p : $(bitstring(p))")
+println("avogadro: $a  : $(bitstring(a))")
+println("p + a:    $(p + a)  : $(bitstring(p + a))")
+\end{lstlisting}
+
+This outputs:
+
+\begin{lstlisting}
+planck:   6.62607015e-34 : 0011100100001 ... 001
+avogadro: 6.02214076e23  : 0100010011011 ... 111
+p + a:    6.02214076e23  : 0100010011011 ... 111
+\end{lstlisting}
+
 This is a fairly extreme example; even when the numbers are closer in magnitude, most operations on \fp{} numbers induce loss of accuracy.

### 2023-08-15 paper.tex
Many big points addressed

@@
 \begin{document}
+
+%% TODO items
+%%  - [ ] Normalize casing in all sections, sub-sections, etc
 
@@
 
-\FT{} has two roles:
-
-\begin{enumerate}
-  \item Track exceptional values
-  \item Fuzz \fp{} code from the inside-out
-\end{enumerate}
-
-Building the mechanism to track exceptional values also gave us a way to inject exceptional values wherever \fp{} operations occur.
-We will examine each of these roles, how they are used, and how \FT{} realizes them.
-
-%% What it does, then how it works
+\FT{} is implemented as a Julia library, and is available for download on JuliaHub.\footnote{\url{https://juliahub.com/ui/Packages/FloatTracker/dBXig/1.0.0}}
 
@@
 
-% note: there's a section on performance in the "discussion" section at the end, so we can leave it out here I think
+\subsubsection{Fuzzing}
 
-While other tools like profilers might accomplish this tracking by writing extensions to the language's compiler or runtime, we were able to implement \FT{} entirely in user land.
+The second role of \FT{} is to fuzz code from the inside out.
+While it is wholly possible to fuzz a program by giving it random inputs, this might not help you find \emph{all} the places where an exceptional value could crop up during a calculation.\cite{ddghlllprr-correctness-2022}
+
+The exception tracking infrastructure that we have gives us a simple way to inject exceptional values wherever we want in our code.
+We have a global configuration system that allows us to turn on fuzzing and configure where, when, and how many \NaN{}s\footnote{At time of writing, \FT{} only injects \NaN{}, but it is trivial to extend the set of values to inject.} to inject.
+
+\FT{} exposes these configuration options:
+
+\begin{itemize}
+\item \texttt{active::Bool} Inject only if this is true.
+\item \texttt{n\_inject::Int64} Inject at most this many \NaN{}s.
+\item \texttt{odds::Int64} Inject if \texttt{rand(odds) == 1}.
+\item \texttt{functions::Array} Inject if we are within the dynamic extent of any of these functions.
+\item \texttt{libraries::Array} Inject if the top-most stack frame\footnote{After stripping off the stack frames that come from \FT{} itself.} comes from any of these libraries.
+\item \texttt{replay::String} Name of a file to read an injection recording from.
+\item \texttt{record::String} Name of a file to write injection events to for later replay.
+\end{itemize}
+
+Then, each time we intercept a \fp{} operation, we check to see if we should inject a \NaN{}.
+First we check if we're replaying a recording: if we are, just do as the recording says.
+Otherwise, we check to make sure we have \NaN{}s remaining to inject and we roll a die to see if we should inject.
+We also check the current stack trace against the lists of functions and libraries to inject in.
+If all looks good, we inject.
+
+In our experience, requesting the stack trace from Julia is the most expensive part of this operation.
+Therefore, we defer the stack checks as late as we can.
+Once the fuzzer runs out of \NaN{}s, it effectively turns off, so the rest of the fuzzing session proceeds much more quickly.
+That said, gathering stack traces still incurs a heavy overhead during logging when interesting events occur;
+we recommend limiting the number and kinds of events logged for optimal performance.
+
+\subsubsection{\FT{} internals}
+
+While other tools like profilers might accomplish the tracking that \FT{} does by writing extensions to the language's compiler or runtime, we were able to implement \FT{} entirely in user land.
 We created custom types \texttt{TrackedFloat16}, \texttt{TrackedFloat32}, and \texttt{TrackedFloat64} that wrap their corresponding \texttt{Float16}, \texttt{Float32}, and \texttt{Float64} counterparts.
@@
 The same classification rules apply to \Inf{} and would apply to any other value of interest.
-
-\subsubsection{Fuzzing}
-
-The second role of \FT{} is to fuzz code from the inside out.
-While it is wholly possible to fuzz a program by giving it random inputs, this might not help you find \emph{all} the places where an exceptional value could crop up during a calculation.\cite{ddghlllprr-correctness-2022}
-
-The exception tracking infrastructure that we have gives us a simple way to inject exceptional values wherever we want in our code.
-We have a global configuration system that allows us to turn on fuzzing and configure where, when, and how many \NaN{}s\footnote{At time of writing, \FT{} only injects \NaN{}, but it is trivial to extend the set of values to inject.} to inject.
-
-\FT{} exposes these configuration options:
-
-\begin{itemize}
-\item \texttt{active::Bool} Inject only if this is true.
-\item \texttt{n\_inject::Int64} Inject at most this many \NaN{}s.
-\item \texttt{odds::Int64} Inject if \texttt{rand(odds) == 1}.
-\item \texttt{functions::Array} Inject if we are within the dynamic extent of any of these functions.
-\item \texttt{libraries::Array} Inject if the top-most stack frame\footnote{After stripping off the stack frames that come from \FT{} itself.} comes from any of these libraries.
-\item \texttt{replay::String} Name of a file to read an injection recording from.
-\item \texttt{record::String} Name of a file to write injection events to for later replay.
-\end{itemize}
-
-Then, each time we intercept a \fp{} operation, we check to see if we should inject a \NaN{}.
-First we check if we're replaying a recording: if we are, just do as the recording says.
-Otherwise, we check to make sure we have \NaN{}s remaining to inject and we roll a die to see if we should inject.
-We also check the current stack trace against the lists of functions and libraries to inject in.
-If all looks good, we inject.
-
-In our experience, requesting the stack trace from Julia is the most expensive part of this operation.
-Therefore, we defer the stack checks as late as we can.
-Once the fuzzer runs out of \NaN{}s, it effectively turns off, so the rest of the fuzzing session proceeds much more quickly.
-That said, gathering stack traces still incurs a heavy overhead during logging when interesting events occur;
-we recommend limiting the number and kinds of events logged for optimal performance.
 
@@
 CSTG~\cite{humphreySystematicDebuggingMethods2014} is built on the work of and related to the STAT~\cite{arnoldStackTraceAnalysis2007} tool developed by Laurence Livermore National Laboratory.
-Please see the CSTG paper for more on the differences between the two.
+STAT is a tool for collecting, analyzing, and visualizing stack traces from thousands of concurrent processes, and is designed to highlight anomalies quickly.
+It produces visualizations similar to those of CSTG, thought CSTG offers more compact views.
 
@@
 \subsection{Tracking more than \NaN{}s and \Inf{}s}
+
 FloatTracker is capable of monitoring all sorts of events in a \fp{} value's lifetime---not just a value going to \NaN{} or \Inf{}.
@@
 
-\subsection{Static FPX, Gradual Types}
+\subsection{Tracking exceptions in external libraries}
 
-FILL
+\FT{}'s scope is limited to Julia code: it cannot track the lifetime of exceptional values that are born, live, or die in external libraries, such as C programs.
+One approach would be to create an API for \FT{} so that programs using these external libraries could leverage \FT{}'s tooling to monitor the interface between Julia and e.g. external C programs.
 
@@
 
+We would like to thank Alex Larsen and Rob Durst for their help with reading drafts of the paper and feedback on the presentation.
 We would also like to thank the developers of \code{RxInfer.jl} for feedback on \FlowFPX{}.
+Finally, many thanks to Milan Klöwer, the author of the \texttt{Sherlogs} library, for inspiration on the architecture of \FT{}.
 
@@
 
-% Below is some Emacs-specific stuff. If you're using Emacs, you should try
-% using the new jinx package for spell-checking!
+% Below is some Emacs-specific stuff to help me (Ashton) write this paper. If
+% you're using Emacs, you should try using the new jinx package for
+% spell-checking!
 

### 2023-08-15 paper.tex
Finish first column of local feedback

@@
   Reliable numerical computations are central to high-performance computing and machine learning.
+  However, floating-point numbers exhibit non-intuitive behaviors that make writing correct numerical computation difficult.
+  In particular, \emph{floating-point exceptions} occur when no mathematically sensible result is available, and it is difficult to find \emph{where} these exceptions occur.
+  Unhandled exceptions can cause programs crash, exhibit bad behavior, or worse, silently produce good-looking but incorrect data.
   We present FlowFPX: a tool set for debugging floating-point code.
-  FlowFPX is composed of two primary parts: FloatTracker, a Julia-based tool for tracking the onset and flow of floating-point exceptions; and CSTG, a stack trace visualizer.
-  FlowTracker's design exploits Julia's operator overloading to trace exception flows and even inject exceptions to accelerate testing.
-  We present intuitive visualizations of summarized exception flows including how they are generated, propagated and killed, thus helping with debugging and repair.
+  FlowFPX can track floating-point exceptions occurring in Julia code to their sources.
+  FlowFPX also detects otherwise silent exceptions which can foul program outputs.
+  FlowFPX produces intuitive visualizations of summarized exception flows including how they are generated, propagated and killed, thus helping with debugging.
 \end{abstract}
@@
 unreliable~(\cref{s:background}).
-Exceptional values, namely Not a Number (\Nan{}) and infinity (\Inf{}),
+Exceptional values, such as Not a Number (\Nan{}) and infinity (\Inf{}),
 can and often do arise thanks to culprits such as roundoff error,
@@
 There is little tool support to assist in this difficult task,
-thus (unsurpringly) a quick GitHub search reports over 4,000 open issues
+thus (unsurprisingly) a quick GitHub search reports over 4,000 open issues
 related to \NaN{} or \Inf{} exceptions~\cite{github-issues}.
@@
 We introduce \FlowFPX{}, a toolkit for debugging
-\fp{} exceptions in an easy and intuitive way~(\cref{s:flowfpx}):
+\fp{} exceptions~(\cref{s:flowfpx}):
 \begin{itemize}
@@
 %
-Together, the \FlowFPX{} tools have helped debug exceptions in a variety
+\FlowFPX{} has helped debug exceptions in a variety
 of applications: from ocean simulations to heat models~(\cref{s:casestudies}).
-That said, there is much more to explore.
-\FlowFPX{} is a growing project and we welcome feedback from domain experts
-on how to best suit their workflow.
-
 This paper concludes with related work~(\cref{s:related}) and a
@@
 
-\subsection{Enhanced Fuzzing}
+\subsection{Enhanced fuzzing}
 
@@
 
+\subsection{Future work}
+
+\FlowFPX{} is a growing project and we welcome feedback from domain experts
+on how to best suit their workflow.
+If you try any of our tools, please send us comments, suggestions, and bug reports via the tools' repositories on GitHub.
+
 \section{Acknowledgments}

### 2023-08-15 paper.tex
Work from second column

@@
 \Fp{} provides a finite approximation of the real numbers in a format that is efficient for calculation on a computer.
-IEEE 754~\cite{IEEEStandardBinary1985} standardizes \fp{} representation and operations---these days, most machines and languages follow the IEEE 754 standard, and we make no effort to generalize our work beyond what IEEE 754 specifies.
-
-\Fp{} is essentially scientific notation---\fp{} numbers are represented by packing a sign bit, an exponent, and an fraction part (also called the ``significand'' or ``mantissa'') into contiguous bit fields.
+\Fp{} is essentially scientific notation---\fp{} numbers are represented by packing a sign bit, an exponent, and a fraction part (also called the ``significand'' or ``mantissa'') into contiguous bit fields.
 The advantages of this representation are that allows for compact representation of both very small and very large numbers in a narrow range of bits---a 64-bit float can represent numbers as big as $2^{1024}$ and as small as $2^{-1074}$.
@@
 This is a fairly extreme example; even when the numbers are closer in magnitude, most operations on \fp{} numbers induce loss of accuracy.
+
+IEEE 754~\cite{IEEEStandardBinary1985} standardizes \fp{} representation and operations---these days, most machines and languages follow the IEEE 754 standard, and we make no effort to generalize our work beyond what IEEE 754 specifies.
 

### 2023-08-16 paper.tex
Fix casing in section titles
Lots more edit work

@@
 \begin{document}
-
-%% TODO items
-%%  - [ ] Normalize casing in all sections, sub-sections, etc
 
@@
 
-\section{Floating Point Exception Primer}
+\section{Floating-Point Exception Primer}
 \label{s:background}
 
-\Fp{} provides a finite approximation of the real numbers in a format that is efficient for calculation on a computer.
 \Fp{} is essentially scientific notation---\fp{} numbers are represented by packing a sign bit, an exponent, and a fraction part (also called the ``significand'' or ``mantissa'') into contiguous bit fields.
@@
 
-The flip side is that not every real number within that wide range is representable as a \fp{} value---there is almost always some gap between the real number that we wish to represent and the \fp{} number that can represent.
-Whenever a value we have doesn't have a precise representation, we must \emph{round} that real value to the nearest \fp{} value.
-We call this gap error, and error, along with the finite nature of \fp{} values, means that we can run into some funny situations.
+\begin{figure}[h]
+  \label{fig:real_vs_float}
+  \includegraphics[width=\columnwidth]{ fig/real_vs_fp.png }
+  \caption{The real number line vs the points representable with \fp{}.}
+\end{figure}
 
-For example, if two numbers' exponents differ by too large an amount, adding or subtracting the smaller from the bigger can result in just the bigger number.
+The flip side is that not every real number within that wide range is representable as a \fp{} value---there is almost always some gap between the real number that we wish to represent and the \fp{} number that approximates that value.
+Whenever a value we have doesn't have a precise representation, we must \emph{round} that real value to the nearest \fp{} value, which introduces some error in our calculations.
+In addition to this error, the finite nature of \fp{} values means that we can run into some situations that wouldn't arise when working with real numbers.
+
+For example, if two numbers' exponents differ by too large an amount, adding or subtracting the smaller from the bigger will leave us with just the bigger number.
 Below we take the Planck constant and add it to Avogadro's constant and are left bit-for-bit with Avogadro's constant, as if we had never done any addition at all: there is no way to pack all the significant figures from true value that differs minutely from Avogadro's number into a 64-bit \fp{} number.
@@
 
-This outputs:
+You can see in the output that \texttt{p + a} is bit-for-bit equivalent to \texttt{a}:
 
@@
 
-IEEE 754~\cite{IEEEStandardBinary1985} standardizes \fp{} representation and operations---these days, most machines and languages follow the IEEE 754 standard, and we make no effort to generalize our work beyond what IEEE 754 specifies.
+\subsection{Floating-Point Standards}
 
-\subsection{Exceptions and exceptional values}
+Prior to 1985, incomparable techniques for representing \fp{} values proliferated as mainframes and later personal computers emerged.
+This made writing portable \fp{} code challenging.~\cite{houghIEEEStandard7542019}
+IEEE 754~\cite{IEEEStandardBinary1985} standardizes \fp{} representation and operations across different architectures---these days, most machines and languages follow the IEEE 754 standard, and we make no effort to generalize our work beyond what IEEE 754 specifies.
+
+\subsection{Exceptions and Exceptional Values}
 
@@
 
-For example, dividing by zero is an exception, as is exceeding your \fp{} format's range.
-These are exceptions because it is impossible to say what the right answer \emph{should} have been: perhaps that zero was meant to be a value that got really \emph{close} to zero, and suddenly rounded towards zero before division.
-Or maybe that value that got really big \emph{does} need to get that big, but you need to start using a larger format for your \fp{} values.
+For example, dividing by zero is an exception, as is exceeding a \fp{} format's range.
+These are exceptions because it is impossible to say what the right answer \emph{should} have been: perhaps that zero was meant to be a value that got really \emph{close} to zero,\footnote{The third kind of exceptional value is called a \emph{subnormal} value.} and suddenly rounded towards zero before division.
+Or maybe that value that got really big \emph{does} need to get that big, but you need to start using a larger bit width for your \fp{} values.
 Whatever, the case, there isn't a universally acceptable answer, so a \fp{} exception gets raised.
 
-IEEE 754 doesn't always obligate the programmer to handle these exceptions, however, so we end up with \emph{exceptional values}, which are the results of exception-raising operations.
-The two values we work with in this paper are \Inf{} and \NaN{}, which result from overflow and nonsensical operations respectively.\footnote{The third kind of exceptional value is called a \emph{subnormal} value.
-Please see~\cite{torontoPracticallyAccurateFloatingPoint2014} for a good introduction to subnormal values, as well as other details of \fp{} representation.}
+IEEE 754 doesn't force the programmer to always handle these exceptions, however, so we end up with \emph{exceptional values}, which are the results of exception-raising operations.
+The two values we work with in this paper are \Inf{}\footnote{There are actually two infinities: $+\infty$ and $-\infty$; they behave the same for our purposes, so we will refer to them just as ``\Inf{}''.} and \NaN{}, which result from overflow and nonsensical operations respectively.
+\Cref{fig:nan-gens} lists some of the situations that lead to exceptional values.
 
-IEEE 754 lists the following operations (among a few others) where \NaN{}s result:
-
-\begin{figure}[h]
+\begin{figure}[ht]
+  \centering
   \label{fig:nan-gens}
-  \begin{itemize}
-  \item $\frac{0}{0}$ or $\frac{\infty}{\infty}$
-  \item $0 * \infty$
-  \item $x \% y$ when $x = \infty$ or $y = 0$
-  \item $+\infty + -\infty$ or $\infty - \infty$
-  \item The square root or logarithm of a negative number
-  \end{itemize}
-  \caption{\NaN{}-producing operations}
+  % \begin{tabular}{c | p{0.5 \columnwidth}}
+  \begin{tabular}{c | c}
+    \NaN{} & \Inf{} \\
+    \hline
+    $\tfrac{0}{0}$ & $\tfrac{x}{0},\, x \neq 0$ \\
+    $0 * \infty$ & $\log{(0)}$ \\
+    $\text{\texttt{NaN}} + x$ & $\infty + x,\, x \neq \text{\texttt{NaN}}$ \\
+    $\infty - \infty$ & Any result exceeding \\
+    & representation capacity
+  \end{tabular}
+  \caption{Examples of operations leading to exceptional values}
 \end{figure}
 
-There are actually two infinities, $+\infty$ and $-\infty$, but they behave the same for our purposes, so we will refer to them just as ``\Inf{}''.
-Like \NaN{}s, \Inf{} can crop up when bad math happens: $\frac{x}{0} \rightarrow \infty$ when $x \ne 0$.
-\Inf{} can also arise when numbers get too big or too small for the choice of representation.
+For more details on exceptional values, see~\cite{knuthArtComputerProgramming1997,torontoPracticallyAccurateFloatingPoint2014} as well as the \textit{Handbook of Floating-Point Arithmetic}.~\cite{mullerHandbookFloatingPointArithmetic2018}
 
-There are many excellent resources\cite{knuthArtComputerProgramming1997,torontoPracticallyAccurateFloatingPoint2014} on understanding the theory behind \fp{} as well as how error is measured and accumulates.
-Please see those works for more details.
-What we have described above should suffice for the remainder of this paper.
+\subsection{Lifetimes of Unhandled Exceptions}
+\label{s:to-kill-a-fp}
 
-% Here is a short, illustrative example about the problem.
-\subsection{To Kill a Floating Point}
-\label{s:to-kill-a-fp}
-% If anyone has a better "kill" pun/reference than "To Kill a Mockingbird", I'm all ears.
+Exceptional values have a life cycle: instead a program-halting exception, an exceptional value comes into being at the point of an undefined operation and propagates through the rest of the computation.
+Often they go on to manifest in the output of the code, but sometimes they can ``die'' by disappearing silently.
+We characterize this life cycle in \cref{fig:gpk}.
 
-Exceptional values have a life cycle: they are born, live, and die, as seen in \cref{f:gpk}.
-
-\begin{figure}
+\begin{figure}[ht]
   \includegraphics[width=\columnwidth]{fig/genpropkill-outline.png}
@@
 
-When a \NaN{} or an \Inf{} comes into being as the result of a bad math operation, we call this a generation, or \emph{gen} for short.
+When a \NaN{} or an \Inf{} comes into being as the result of an undefined operation, we call this a generation, or \emph{gen} for short.
 
-The IEEE 754 specification for \fp{} numbers governs how \NaN{}s are handled as arguments to arithmetic functions~\cite{IEEEStandardBinary1985}.
 Most of the time, when any or all of the arguments are \NaN{}, the result is also a \NaN{}.
 We call this a propagation, or \emph{prop} for short.
-
 This is good and desirable, because it means that an arithmetic error---even if buried deep in an expression---will still be observable, allowing a user to find and debug it.
@@
 Both \texttt{bad\_max} and \texttt{good\_max} compute the maximum in a list.
-However, the \texttt{<=} operator used in the \texttt{bad\_max} implementation does \emph{not} propagate \NaN{}s, so you might not get the result you expected:
+However, the \texttt{<=} operator used in the \texttt{bad\_max} implementation does \emph{not} propagate \NaN{}s: \texttt{NaN <= 4} returns \texttt{false}, not \NaN{}.
+Because of this, you might not get the result you expected:
 
@@
 The built-in \texttt{max} function correctly propagates \NaN{}s, so we can see from its return that we got a \NaN{} somewhere.
-It's not entirely obvious \emph{where} that \NaN{} came from---but more on that later.   % this promise answered in "NaN propagation monitoring"
+It's not entirely obvious \emph{where} that \NaN{} came from.
 
@@
 
-\FlowFPX{} is a growing set of tools to aid in tracking down \fp{} exceptions.
+\FlowFPX{} is a set of tools to aid in tracking down \fp{} exceptions.
 The two primary tools we cover in this paper are \FT{} and CSTG:
@@
 \begin{itemize}
-\item \FT{}'s role is twofold: to trace the lifetime of exceptional \fp{} values and to enable inside-out fuzzing.
-\item CSTG's role is to visualize the tracing output from \FT{} in compact form.
+\item \FT{} traces the lifetime of exceptional \fp{} values and enables inside-out fuzzing.
+\item CSTG visualizes the tracing output from \FT{} in compact form.
 \end{itemize}
 
-\GPUFPX\cite{llsflg-hpdc-2023} is another component of \FlowFPX{} that can track \fp{} exceptions inside a GPU kernel.
-Please see the cited paper for more details on \GPUFPX.
-Work to use \GPUFPX in tandem with \FT{} is ongoing.
+\GPUFPX~\cite{llsflg-hpdc-2023} is another component of \FlowFPX{} that can track \fp{} exceptions inside a GPU kernel.
+See the cited paper for more details on \GPUFPX.
 
@@
 
-\subsubsection{\FT{} internals}
+\subsubsection{\FT{} Internals}
 
@@
 
-\subsubsection{CSTGs help us get a bigger picture}
+\subsubsection{CSTGs Help Us Get a Bigger Picture}
 
@@
 
-\subsection{A Bug-Hunting Safari}
+\subsection{RxInfer: No Help Needed}
 \label{s:safari}
@@
 
-\subsection{Floating-point error analysis}
+\subsection{Floating-Point Error Analysis}
 
@@
 
-\subsection{Stack trace graphs}
+\subsection{Stack Trace Graphs}
 
@@
 
-\subsection{Floating-point exceptions on the GPU}
+\subsection{Floating-Point Exceptions on the GPU}
 
@@
 
-\subsection{Tracking more than \NaN{}s and \Inf{}s}
+\subsection{Tracking More than \NaN{}s and \Inf{}s}
 
@@
 
-\subsection{Enhanced fuzzing}
+\subsection{Enhanced Fuzzing}
 
@@
 
-\subsection{Tracking exceptions in external libraries}
+\subsection{Tracking Exceptions in External Libraries}
 
@@
 
-\subsection{Future work}
+\subsection{Future Work}
 

### 2023-08-16 paper.tex
Big rewrite of §3.1

@@
 
-\FT{}'s first purpose is to monitor \fp{} operations for exceptional values.
-\FT{} interposes on every arithmetic operation and standard library call that involves a \fp{} and logs a stack trace to that point whenever an exception is generated, propagated, or killed.
+\FT{}'s first purpose is to monitor \fp{} operations for exceptional value lifetime events.
+\FT{} interposes on arithmetic operations and standard library calls that involve a \fp{} value and logs a stack trace to that point whenever an exception is generated, propagated, or killed.
+\FT{} can also log the arguments to the suspect call, so you can get \emph{normal, concrete} values to help discover what is going wrong.
 A programmer can then use these stack traces to pinpoint where an exceptional value like \NaN{} came into being or where a guard against a \NaN{} kill is needed.
-\FT{} can also log the arguments to the suspect call, so you can get \emph{normal, concrete} values to help you further discover what is going wrong.
 
 All the logs that \FT{} captures are straight from Julia's \texttt{stacktrace()} function.
-\FT{} writes these logs to file; one file for each kind of event.
+\FT{} writes these logs to three files: one file for each kind of event in a \fp{} exception's lifetime.
 This way, if a user of \FT{} is attempting to determine where a \NaN{} or an \Inf{} came from, they can look in the file for all the gen events.
@@
 The second role of \FT{} is to fuzz code from the inside out.
-While it is wholly possible to fuzz a program by giving it random inputs, this might not help you find \emph{all} the places where an exceptional value could crop up during a calculation.\cite{ddghlllprr-correctness-2022}
+While can fuzz a program by giving it random inputs, this might not help you find \emph{all} the places where an exceptional value could crop up during a calculation.
+Demmel et. al suggest fuzzing by injecting exceptional values at intermediate calculations;~\cite{ddghlllprr-correctness-2022} this is the approach that we enable with \FT{}.
 
-The exception tracking infrastructure that we have gives us a simple way to inject exceptional values wherever we want in our code.
-We have a global configuration system that allows us to turn on fuzzing and configure where, when, and how many \NaN{}s\footnote{At time of writing, \FT{} only injects \NaN{}, but it is trivial to extend the set of values to inject.} to inject.
+\FT{}'s exception tracking infrastructure offers us a simple way to inject exceptional values wherever we want in our code: we are already interposing on \fp{} operations to monitor for exceptional lifetime values, so we can inject a \NaN{} at any of these points.
 
-\FT{} exposes these configuration options:
+Since the injection happens at random, \FT{} also provides a way to save \NaN{} injections so that injections that exercise bugs can be replayed to examine the bug repeatedly and to verify that hardening efforts have indeed eliminated the vulnerability.
+
+We have a global configuration system that allows us to turn on fuzzing and configure where, when, and how many \NaN{}s to inject, as well as whether or not to record or replay a recording:
 
 \begin{itemize}
+\item \texttt{replay::String} Name of a file to read an injection recording from. If set, this option makes all others irrelevant.
+\item \texttt{record::String} Name of a file to write injection events to for later replay.
 \item \texttt{active::Bool} Inject only if this is true.
@@
 \item \texttt{libraries::Array} Inject if the top-most stack frame\footnote{After stripping off the stack frames that come from \FT{} itself.} comes from any of these libraries.
-\item \texttt{replay::String} Name of a file to read an injection recording from.
-\item \texttt{record::String} Name of a file to write injection events to for later replay.
 \end{itemize}
@@
 First we check if we're replaying a recording: if we are, just do as the recording says.
-Otherwise, we check to make sure we have \NaN{}s remaining to inject and we roll a die to see if we should inject.
-We also check the current stack trace against the lists of functions and libraries to inject in.
+Otherwise, we check that the injector is \texttt{active}, we have \NaN{}s remaining to inject (\texttt{n\_inject}), and we roll an \texttt{odds}-sided die to see if we should inject.
+Finally, check the current stack trace against the lists of functions and libraries to inject in.
 If all looks good, we inject.
-
-In our experience, requesting the stack trace from Julia is the most expensive part of this operation.
-Therefore, we defer the stack checks as late as we can.
-Once the fuzzer runs out of \NaN{}s, it effectively turns off, so the rest of the fuzzing session proceeds much more quickly.
-That said, gathering stack traces still incurs a heavy overhead during logging when interesting events occur;
-we recommend limiting the number and kinds of events logged for optimal performance.
 
@@
 
-While other tools like profilers might accomplish the tracking that \FT{} does by writing extensions to the language's compiler or runtime, we were able to implement \FT{} entirely in user land.
-We created custom types \texttt{TrackedFloat16}, \texttt{TrackedFloat32}, and \texttt{TrackedFloat64} that wrap their corresponding \texttt{Float16}, \texttt{Float32}, and \texttt{Float64} counterparts.
-We then defined methods for every function in the \texttt{Base} module, including new implementations for operators like \texttt{+} and \texttt{*}.
-This lets us intercept all \fp{} operations working on our \texttt{TrackedFloat} types.
+While other tools like profilers might accomplish the tracking that \FT{} does by writing extensions to the language's compiler or runtime, \FT{} works entirely in user land.
+\FT{} provides custom types---\texttt{TrackedFloat16}, \texttt{TrackedFloat32}, and \texttt{TrackedFloat64}---that wrap their corresponding \texttt{Float16}, \texttt{Float32}, and \texttt{Float64} counterparts.
+\FT{} also defines methods for every function in the \texttt{Base} module.
+This lets \FT{} intercept all \fp{} operations involving \texttt{TrackedFloat} types.
 
@@
 function +(x::TrackedFloat64, y::TrackedFloat64)
-  result = x.val + y.val
+  result = run_or_inject(+, x.val, y.val)
   check_error(+, result, x.val, y.val)
@@
 \end{lstlisting}
+
+This new method does a few things:
+
+\begin{enumerate}
+\item First it checks if it should inject a \NaN{} here.
+\item If no \NaN{} is to be injected, it unwraps the \texttt{TrackedFloat}s to get to the primitive \texttt{Float} value, and computes the result normally.
+\item It examines the arguments as well as the return value, and then uses \cref{fig:lifetime-class} to determine the type of event and logs it if it is of interest.
+\item It wraps the result in a new \texttt{TrackedFloat} value and returns it.
+\end{enumerate}
+
+\begin{figure}[h]
+  \begin{tabular}{ccc}
+    Arguments & Result & Event Type \\
+    \hline
+    No \NaN{} & \NaN{} & Gen \\
+    \NaN{} & \NaN{} & Prop \\
+    \NaN{} & No \NaN{} & Kill \\
+  \end{tabular}
+  \label{fig:lifetime-class}
+  \caption{Exceptional value event classification rules}
+\end{figure}
+
+The same classification rules apply to \Inf{} and would apply to any other value of interest.
 
@@
 This approach was inspired by \texttt{Sherlogs.jl}.\cite{kMilanklSherlogsJl2021}
-
-With the overrides in place, we can examine the arguments to a \fp{} operation to determine if an interesting event occurred.
-Here is how we classify events around a \NaN{}:
-
-\begin{center}
-  \begin{tabular}{ccc}
-    Arguments & Result & Event Type \\
-    \hline
-    No \NaN{} & \NaN{} & Gen \\
-    \NaN{} & \NaN{} & Prop \\
-    \NaN{} & No \NaN{} & Kill \\
-  \end{tabular}
-\end{center}
-
-The same classification rules apply to \Inf{} and would apply to any other value of interest.
 
@@
 
-\begin{figure}
+\begin{figure}[ht]
   \centering
@@
 
+When injection is turned on we defer the stack checks as late as we can.
+Once the fuzzer runs out of \NaN{}s, it effectively turns off, so the rest of the fuzzing session proceeds much more quickly.
+That said, gathering stack traces still incurs a heavy overhead during logging when interesting events occur;
+we recommend limiting the number and kinds of events logged for optimal performance.
+
 \subsection{Tracking More than \NaN{}s and \Inf{}s}

### 2023-08-16 paper.tex
Finish §3.2 edits

@@
 \FT{} can produce copious amounts of log files which can be challenging to sift through manually.
-Coalesced Stack Trace Graphs, or CSTGs\cite{humphreySystematicDebuggingMethods2014} provide a way to visualize large amounts of stack traces in a compact form.
+Coalesced Stack Trace Graphs, or CSTGs~\cite{humphreySystematicDebuggingMethods2014} provide a way to visualize large amounts of stack traces in a compact form.
 CSTG is a stand-alone tool and was not designed with \FT{} specifically in mind.
-However, it pairs well with \FT{} and makes coming through the log files easier.
+However, it pairs well with \FT{} and makes analyzing the log files easier.
 
@@
 
-CSTG's output looks like the output seen in \cref{fig:sw_nan_cstg}, which was generated by running \FT{} on a library called ShallowWaters.
-We will go into more detail about ShallowWaters in \cref{s:sw}.
-This graph has been edited for clarity and to fit into the size constraints of this paper.
-
 These graphs are meant to be read bottom-up: the bottom nodes are the roots where exceptional values manifested.
@@
 \label{s:gpufpx}
-
-% Acknowledge that apps get offloaded to GPU; FT doesn't do that, but there's this companion tool…
-% You run it on a kernel; catches things not caught by GPUs; see Oceananigans as an example below
-% TODO: add note in future work/discussion about this as well; other languages too—not just GPU
-% FloatTracker could supply an API that wraps calls to a C library—lib is still a black box, but FT gives you a big picture
 

### 2023-08-17 paper.tex
More edits

@@
 
+% HELP: @ben, I'm not sure what paper you mean by "Julia CPU"
 Many programs offload work onto GPUs, which are no less susceptible to \fp{} exceptions than CPUs.
@@
 Thus, tools to catch \fp{} exceptions on the GPU are essential.
-\FT{} is constrained to CPU workloads; however, there is a companion tool called \GPUFPX{}\cite{llsflg-hpdc-2023} that can help diagnose \fp{} exceptions on the GPU.
+\FT{} is constrained to CPU workloads; however, there is a companion tool called \GPUFPX{}~\cite{llsflg-hpdc-2023} that can help diagnose \fp{} exceptions on the GPU.
 \GPUFPX{} can detect floating-point exceptions on the GPU that GPU code cannot do on its own.
@@
 
-See \cref{s:ocean} for an example of \GPUFPX{} in action, and \ref{f:gpufpx} for an example of its output.
-
-% * what about GPUs? companion project at Utah GPU-FPX
-% * architecture diagram, CPU + GPU, show Ocean source code
-% * replace CPU() with GPU() in code, get output like below & attached
-
-%% Ganesh
-% 3) These are conceptually usable together (Julia calling GPU kernels below).
-% The usage is simple: Start the Julia application that calls a GPU kernel under
-% LD_PRELOAD as with GPU-FPX launch. When this launch transfers control to a GPU
-% kernel, the GPU-FPX instrumentation starts notifying the exceptions on the
-% console of the CPU. When the GPU launch returns, the exceptions go in the
-% return call path of the Julia execution. Ashton and Xinyi plz validate this
-% scenario
-
 \section{Case Studies}
@@
 In this section we'll discuss some case studies where tools from \FlowFPX{} aided in finding or fixing \fp{} issues.
-Our case studies are as follows:
-
-\begin{itemize}
-\item \Cref{s:sw} covers a simulation that produces \NaN{}s in the output---a common and frustrating occurrence for many a scientist.
-\item \Cref{s:ode} puts a widely-used mathematics library under a fuzzer to harden it against bad behavior in the presence of exceptional \fp{} values.
-\item \Cref{s:finch} puts another program under the fuzzer.  % <- ugh, rewrite
-\item \Cref{s:ocean} is a simulation that runs on the GPU---this puts it out of reach of \FT{} but not for other components of the \FlowFPX{} tool set.
-\end{itemize}
-
-Besides case studies that we ourselves did, \cref{s:safari} is an ``in the wild'' usage of \FT{} that helped researchers uncover the source of a \NaN{} in their own code.
+\Cref{s:sw} covers an ocean simulation that produces \NaN{}s in the output---a common and frustrating occurrence for many a scientist.
+\Cref{s:ode} puts the popular \texttt{OrdinaryDiffEq} solver library under a fuzzer to harden it against bad behavior in the presence of exceptional \fp{} values.
+\Cref{s:finch} fuzzes another library---this time a library for specifying partial differential equations (PDEs).
+Finally, in \cref{s:ocean}, we examine another ocean simulation library---this time the library runs on the GPU.
+This puts it out of reach of \FT{}, but \GPUFPX{} can handle it.
+In addition to the case studies that we did ourselves, \cref{s:safari} is an ``in the wild'' usage of \FT{} that helped researchers uncover the source of a \NaN{} in their own code without help from us.
 
@@
 
-For our first case study, we looked at the \texttt{ShallowWaters} library~\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019,klowerLowprecisionClimateComputing2021}, which runs a simulates the flow of water over a seabed.
+\texttt{ShallowWaters}~\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019,klowerLowprecisionClimateComputing2021} simulates the flow of water over a seabed.
 This is not a general shallow waters simulation library---it is designed primarily to showcase numerical methods involving 16-bit floats.
-That said, it works for our purposes as a proving ground for FloatTracker.
 
-Many physics simulations of this sort include a parameter called the Courant-Friedrichs-Lewy (``CFL'') parameter, which roughly describes the size of the time step to take in running the simulation.
+The Courant-Friedrichs-Lewy (``CFL'') parameter roughly describes the size of the time step to take in running the simulation.
 Normally this value sits between 0 and 1---smaller values mean the simulation runs slower but with higher accuracy, whereas higher values mean the simulation takes bigger time steps and consequently runs faster, but results might be inaccurate.
@@
 
-The holes in the plots are regions where \NaN{}s crept into the computation and destroyed the results.
+The white regions in the right graph of \cref{fig:sw_nans} are regions where \NaN{}s crept into the computation and destroyed the results.
 In one regard, this is not as bad as a \NaN{} kill---knowing that there are \NaN{}s in the computation means that we have a chance to fix it, whereas a \NaN{} kill might silently give us the wrong answer.
-Incidentally, we know that there were no \NaN{} kills in this test because our kill logs are empty.
-But finding the sources of \NaN{}s can be tricky.
+Incidentally, we know that there were no \NaN{} kills in this test all the events in our kill logs are from calling the \texttt{isnan} function---a clear sign that the program is deliberately taking action on the presence of a \NaN{}.
+
+Finding the sources of \NaN{}s can be tricky.
 We expected \NaN{}s in this case because we set the CFL parameter up so high, but tracking down the sources of \NaN{}s generally is a difficult and time-consuming problem.
 
-Fortunately in this case, one of the key features \texttt{ShallowWaters} is its \emph{type flexibility}:
-\texttt{ShallowWaters} was designed to showcase physics simulations using variable levels of \fp{} precision.
-Upon model construction, \texttt{ShallowWaters} allows you to specify a type for floats in the model.
+\FT{} solves this problem for us by automatically tracking all sources of potential \NaN{} generation.
+And in this case \texttt{ShallowWaters} makes this doubly easy:
+\texttt{ShallowWaters} was designed to showcase physics simulations using variable levels of \fp{} precision, and allows you to specify a \fp{} type to use in the computation.
 We simply asked it to use our \texttt{TrackedFloat32} type, and then the entire simulation ran under the supervision of FloatTracker.
 
-We ran ShallowWaters with a CFL value known to make bad outputs and recorded \NaN{} \texttt{gen} events:
-
 \begin{lstlisting}[language = Julia]
-set_logger(buffersize=20, cstg=true, cstgArgs=true,
-           cstgLineNum=true)
-
-P = run_model(T=TrackedFloat32,
+run_model(T=TrackedFloat32,
           cfl=10, Ndays=100, nx=100, L_ratio=1,
@@
 
-This produced the following in the log file: (formatted for clarity)
+This produced almost 17000 lines of logs in the gens file. Below is a sample stack trace from these logs: (formatted for clarity)
 
 \begin{verbatim}
--                 FT/TrackedFloat.jl:106
+-([-Inf, -Inf])   FT/TrackedFloat.jl:106
 momentum_u!       ShallowWaters/rhs.jl:246
@@
 
-We can see that the \NaN{} appeared from subtraction (first line) and that it was used in the \texttt{momentum\_u!} function on line 246 of the \texttt{rhs.jl} file (second line).
-In figure~\ref{fig:nan-gens} we can tell that this must be the result of $\infty - \infty$.
+We can see that the \NaN{} appeared as the result of $-\infty - -\infty$ (first line) and that subtraction happened in the \texttt{momentum\_u!} function on line 246 of the \texttt{rhs.jl} file (second line).
 Thus, with little effort we were able to track down where the \NaN{}s were originating.
+
 Now the question is: where did the $\infty$ come from?
-
-FloatTracker is not limited to \NaN{}s---we can track any kind of value that we want.
-We have built-in support for tracking \NaN{} and $\infty$; we can look at the file of the $\infty$ gens where we see the following: (formatted for clarity)
+We can look at the file of the $\infty$ gens where we see the following: (formatted for clarity)
 
 \begin{verbatim}
-check_error([-1.515f31, 2])  FT/TrackedFloat.jl:10
-^(::TrackedFloat32, ::Int64) FT/TrackedFloat.jl:138
+^([-1.515f31, 2])            FT/TrackedFloat.jl:138
 …
@@
 We can tell that because the first line shows us the arguments\footnote{There is a configuration option that allows us to control this behavior.} and the second line shows us what function got called.
-Moreover, we get the top-level line: we can see that it came from line 14 in our ShallowWaters file.
 
-Tracking down why that value got so large is beyond the scope of FloatTracker right now.
-Most likely we will need some domain-expertise to find a solution to this problem.
-The point is, we were able to quickly and automatically arrive at this excellent starting point with FloatTracker.
+Now we have some concrete values to work with: instead of wondering where an opaque \Inf{} or a \NaN{} came from, we have \emph{specific} values that a domain expert can use to diagnose the problem.
 

### 2023-08-17 paper.tex
Fix some refs; finish edits through §4.2

@@
 Often they go on to manifest in the output of the code, but sometimes they can ``die'' by disappearing silently.
-We characterize this life cycle in \cref{fig:gpk}.
+We characterize this life cycle in \cref{f:gpk}.
 
@@
 \item It examines the arguments as well as the return value, and then uses \cref{fig:lifetime-class} to determine the type of event and logs it if it is of interest.
-\item It wraps the result in a new \texttt{TrackedFloat} value and returns it.
+\item It wraps the result in a new \texttt{TrackedFloat} value.
 \end{enumerate}
@@
 \begin{figure}[h]
+  \centering
   \begin{tabular}{ccc}
@@
 ^([-1.515f31, 2])          FT/TrackedFloat.jl:138
+literal_pow()              intfuncs.jl:325
 …
 materialize(^)             broadcast.jl:860
-top-level getproperty(P, :u) examples/sw_nan_tf.jl:14
+top-level getproperty(…)  examples/sw_nan_tf.jl:14
 …
@@
 It looks like the $-\infty$ came from $-1.515e31^2$---no wonder a Float32 type couldn't handle something that large.
-We can tell that because the first line shows us the arguments\footnote{There is a configuration option that allows us to control this behavior.} and the second line shows us what function got called.
-
 Now we have some concrete values to work with: instead of wondering where an opaque \Inf{} or a \NaN{} came from, we have \emph{specific} values that a domain expert can use to diagnose the problem.
@@
 To get a quick overview of the most common paths to problems, we can use CSTG to get an overview of what stack frames appear the most often.
+
+We can produce a CSTG of both the \Inf{} as well as the \NaN{} logs.
+\Cref{fig:sw_nan_cstg} and \cref{fig:sw_inf_cstg} show what the two graphs look like.\footnote{As with the call stack logs, these have been edited for clarity.
+We only remove a few boxes in between—mostly due to space constraints in this paper.}
 
@@
 
-We can produce a CSTG of both the \Inf{} as well as the \NaN{} logs.
-Here is what the two graphs look like:\footnote{As with the call stack logs, these have been edited for clarity. We only remove a few boxes in between—mostly due to space constraints in this paper.}
-
 \begin{figure}[ht]
@@
 
+We can see in \cref{fig:sw_nan_cstg} that a thick line goes from the top-level call down to a call to \texttt{+} goes down the right side of the graph.
+This indicates that the majority of the flows took this path.
+That means we might be able to fix the majority of our problems by inspecting that particular flow.
+\Cref{fig:sw_inf_cstg} shows us that all but two flows started with a call to \texttt{^}---a promising place to start looking for improvements.
+
+\subsubsection{CSTG Diffing}
+
 Another one of CSTGs abilities is \emph{graph diffing}.
 We took the \FT{} logs from running ShallowWaters and split the logs at the 10\% mark.
-We used CSTG to diff the flow graphs between the first 10\% of the run of the program and the remaining 90\%.
+We used CSTG to diff the flow graphs between the first 10\% of the logs and the remaining 90\%.
 The diff can be seen in \cref{fig:cstg_diff_demo}.
 
-\begin{figure}
+\begin{figure}[ht]
   \centering
@@
 
-The green lines indicate flows that were \emph{new} in the latter 90\% of the program's run.
-The red lines indicate flows that were \emph{only in} the first 10\% of the program's run.
+The green lines indicate flows that were \emph{new} in the latter 90\% of the logs.
+The red lines indicate flows that were \emph{only in} the first 10\% of the logs.
 What this means depends on the particular program; a domain expert can use these clues to e.g.
 find where instability started in the first part of the program, and where that instability got reinforced later on.
+In this case, we can see that \texttt{momentum_u!} only showed up in the beginning of the logs---we might look there for ways to prevent instability from beginning.
 
@@
 
-We configured \FT{} to inject a single \NaN{} somewhere within the scope of the \texttt{NBodySimulator} and \texttt{OrdinaryDiffEq} libraries.
+We configured \FT{} to inject a single \NaN{} randomly within the scope of the \texttt{NBodySimulator} and \texttt{OrdinaryDiffEq} libraries.
 We found a run where something curious happened: the library reported that it had detected a \NaN{} and would then exit.
@@
 
-% FIXME: should I show that CSTG here in the paper?
-We used CSTGs to get an idea of what the important flows were in the run of this errant behavior.
-With an idea of what was going on, we took a look at the logs and found a \NaN{} kill inside the \texttt{solve.jl} file in the \texttt{OrdinaryDiffEq} library.
+Using CSTGs to guide our search, we found a \NaN{} kill that manifested repeatedly in the logs.
+The stack traces for that kill originated from inside the \texttt{solve.jl} file in the \texttt{OrdinaryDiffEq} library.
 
 \begin{verbatim}
-…
-<       at FloatTracker/src/TrackedFloat.jl:193
-solve!  at OrdinaryDiffEq/src/solve.jl:515
+<([NaN, 3.0e6])  at FT/TrackedFloat.jl:193
+solve!           at OrdinaryDiffEq/solve.jl:515
 …
@@
 
-Our kill logs led us right to this line; this is a perfect example of how \NaN{} kills can influence control flow to go awry.
+Our kill logs led us right to this line and let us know which of the arguments to \texttt{<} was a \NaN{}.
+This is a perfect example of how \NaN{} kills can influence control flow to go awry.
 In this case, the problem was apparent besides what our logs told us and it manifested as an infinite loop.
@@
 
-We were able to use the injection recording and the gen logs to track down where the first \NaN{} was injected to cause the infinite loop.
-This made it simple for us to create an example to reproduce the behavior, which we included in the GitHub issue that we opened.\footnote{\url{https://github.com/SciML/OrdinaryDiffEq.jl/issues/1939}}
+We were able to use the injection recording to reproduce the bug in the GitHub issue that we opened.\footnote{\url{https://github.com/SciML/OrdinaryDiffEq.jl/issues/1939}}
 

### 2023-08-17 paper.tex
Last edits to paper (this round)

@@
 -([-Inf, -Inf])   FT/TrackedFloat.jl:106
-momentum_u!       ShallowWaters/rhs.jl:246
-rhs_nonlinear!    ShallowWaters/rhs.jl:50
-rhs!              ShallowWaters/rhs.jl:14 [inlined]
-time_integration  ShallowWaters/time_integration.jl:77
-run_model         ShallowWaters/run_model.jl:37
+momentum_u!       SW/rhs.jl:246
+rhs_nonlinear!    SW/rhs.jl:50
+rhs!              SW/rhs.jl:14 [inlined]
+time_integration  SW/time_integration.jl:77
+run_model         SW/run_model.jl:37
 top-level
@@
 materialize(^)             broadcast.jl:860
-top-level getproperty(…)  examples/sw_nan_tf.jl:14
+top-level getproperty(...) examples/sw_nan_tf.jl:14
 …
@@
 That means we might be able to fix the majority of our problems by inspecting that particular flow.
-\Cref{fig:sw_inf_cstg} shows us that all but two flows started with a call to \texttt{^}---a promising place to start looking for improvements.
+\Cref{fig:sw_inf_cstg} shows us that all but two flows started with a call to \texttt{\^}---a promising place to start looking for improvements.
 
@@
 find where instability started in the first part of the program, and where that instability got reinforced later on.
-In this case, we can see that \texttt{momentum_u!} only showed up in the beginning of the logs---we might look there for ways to prevent instability from beginning.
+In this case, we can see that \texttt{momentum\_u!} only showed up in the beginning of the logs---we might look there for ways to prevent instability from beginning.
 
@@
 <([NaN, 3.0e6])  at FT/TrackedFloat.jl:193
-solve!           at OrdinaryDiffEq/solve.jl:515
+solve!           at ODE/solve.jl:515
 …
@@
 A \NaN{} injected in the mesh led to a crash further on:
+
 \begin{verbatim}
@@
 \end{verbatim}
+
 Defensive programming solves the issue.
@@
 Even though we were unable to provide hands-on guidance as the code in question was proprietary, the authors of \code{RxInfer.jl} were able use \FT{} themselves, and within a day they tracked down the location of a \NaN{} gen.
-The \code{RxInfer.jl} authors provided us with valuable feedback on \FT{}'s API.
 
@@
 
-Toronto and McCarthy\cite{torontoPracticallyAccurateFloatingPoint2014} developed a methodology for assessing where arithmetic operations are liable to magnify \fp{} error.
-Their work includes error visualization and work on how to avoid error-magnifying ``badlands''.
-They focus on rewriting arithmetic expressions to avoid error which could lead to exceptional values, rather than on exceptional values themselves.
+Toronto and McCarthy\cite{torontoPracticallyAccurateFloatingPoint2014} develop a methodology for assessing where arithmetic operations are liable to magnify \fp{} error.
+They include error visualization and work on how to avoid error-magnifying ``badlands''.
+They focus on rewriting arithmetic expressions to avoid error which could lead to exceptional values, rather than on the exceptional values themselves.
 
@@
 
-The Julia library \texttt{Sherlogs}\footnote{\url{https://github.com/milankl/Sherlogs.jl}} by Milan Klöwer inspired the use of a custom number type to intercept operations on a number.
+The Julia library \texttt{Sherlogs}~\cite{kMilanklSherlogsJl2021} inspired the use of a custom number type to intercept operations on a number.
 In contrast to \FT{} which monitors for exceptional values and logs stack traces at interesting points in their lifetime, \texttt{Sherlogs} tracks and reports the range of values seen over the course of a computation.
@@
 FloatTracker is capable of monitoring all sorts of events in a \fp{} value's lifetime---not just a value going to \NaN{} or \Inf{}.
+For example, \FT{} could be made to monitor extreme values: in some circumstances it would be useful to notice when a value started climbing above or falling below a certain threshold.
+
 We spent most of the time engineering and testing \NaN{} tracking and injection.
@@
 after the fact.
-Program slicing is especially relevant to the latter point (and effectively
-what we did by hand when fuzzing Finch~\cref{s:finch}).
+Program slicing is especially relevant to the latter point and effectively
+what we did by hand when fuzzing Finch (\cref{s:finch}).
 For each operation, an expert needs to study the values that feed into it to

### 2023-08-17 header.tex
checkpoint: 1,2

@@
 
-\keywords{Julia, Floating point, Numerical computing}
+\keywords{Julia, floating-point, debugging}
 
@@
 pdfauthor = {Taylor Allred, Xinyi Li, Ashton Wiersdorf, Ben Greenman, Ganesh Gopalakrishnan},
-pdfkeywords = {Julia, Floating point, Numerical computing},
+pdfkeywords = {Julia, floating-point, debugging},
 }

### 2023-08-17 paper.tex
fix includegraphics, move figs up
[ht] is not the same as [th]

top figures are best to avoid confusion about boundaries
between text, figure, and captions
checkpoint: 1,2

@@
 % Misc typographical tweaks.
-\setlength{\parindent}{1.2em}
+\setlength{\parindent}{10pt}
 % \setmonofont{JuliaMono}[Extension = .ttf, Path = ./, UprightFont = *-Regular, ItalicFont = *-Italic]
@@
 \begin{abstract}
-  Reliable numerical computations are central to high-performance computing and machine learning.
-  However, floating-point numbers exhibit non-intuitive behaviors that make writing correct numerical computation difficult.
-  In particular, \emph{floating-point exceptions} occur when no mathematically sensible result is available, and it is difficult to find \emph{where} these exceptions occur.
-  Unhandled exceptions can cause programs crash, exhibit bad behavior, or worse, silently produce good-looking but incorrect data.
-  We present FlowFPX: a tool set for debugging floating-point code.
-  FlowFPX can track floating-point exceptions occurring in Julia code to their sources.
-  FlowFPX also detects otherwise silent exceptions which can foul program outputs.
-  FlowFPX produces intuitive visualizations of summarized exception flows including how they are generated, propagated and killed, thus helping with debugging.
+  Reliable numerical computations are central to scientific computing,
+  but the floating-point arithmetic that enables large-scale models
+  is fundamentally unreliable.
+  Numeric exceptions are a common occurrence and can propagate through
+  code, leading to flawed results.
+  This paper presents FlowFPX, a toolkit for systematically debugging
+  floating-point exceptions by recording their flow,
+  coalescing exception contexts, and fuzzing in select locations.
+  These tools help scientists discover when exceptions happen
+  and track down their origin, smoothing the way to a reliable codebase.
 \end{abstract}
@@
 machine learning, and scientific applications.
-Yet, the \fp{} arithmetic that powers these computations is fundamentally
+Yet the \fp{} arithmetic that powers these computations is fundamentally
 unreliable~(\cref{s:background}).
@@
 derivatives~\cite{sdjmrstp-pc-2022,ddghlllprr-correctness-2022,gllprt-correctness-2021,fpchecker-reports,llg-soap-2022,bllmg-xloop-2022}.
-When exceptions happen, domain experts must take time away from their main
-focus and trek down a rabbit hole to find the source of the error.
-There is little tool support to assist in this difficult task,
+Developers are responsible for guarding against exceptions, but this task
+is difficult because many operations can generate and propagate exceptions.
+Worst of all, operations such as less-than (\code{<}) can kill an exceptional value,
+leaving no trace of the problem.
+There is little tool support to assist in exception debugging,
 thus (unsurprisingly) a quick GitHub search reports over 4,000 open issues
@@
 
-We introduce \FlowFPX{}, a toolkit for debugging
-\fp{} exceptions~(\cref{s:flowfpx}):
+This paper introduces \FlowFPX{}, a toolkit for debugging
+\fp{} exceptions~(\cref{s:flowfpx})
+that hase helped improve a variety of applications,
+from ocean simulations to heat modes~(\cref{s:casestudies}):
 \begin{itemize}
   \item
-    The centerpiece of \FlowFPX{} is \FT{}, a dynamic analysis tool that uses
-    Julia's operator overloading to selectively monitor for exceptions and to
-    fuzz code for vulnerabilities.
+    The centerpiece of \FlowFPX{} is \FT{}, a dynamic analysis tool that
+    selectively monitors for exceptions and fuzzes code for vulnerabilities.
   \item
-    To visualize results, \FT{} adapts coalesced stack-trace graphs 
-    (CSTGs, or \CSTG{}s)~\cite{humphreySystematicDebuggingMethods2014} for a whole-program view on the generation,
-    propagation, and death of exceptional values.
-  \item A companion tool \GPUFPX{}~\cite{llsflg-hpdc-2023} provides fine-grained insights for GPU kernels.
+    To visualize results, \FT{} adapts coalesced stack-trace graphs (CSTGs, or
+    \CSTG{}s)~\cite{humphreySystematicDebuggingMethods2014}
+    to summarize the program contexts that handled exceptional values.
+  \item
+    A companion tool \GPUFPX{}~\cite{llsflg-hpdc-2023} provides fine-grained insights for GPU kernels.
 \end{itemize}
-%
-\FlowFPX{} has helped debug exceptions in a variety
-of applications: from ocean simulations to heat models~(\cref{s:casestudies}).
-This paper concludes with related work~(\cref{s:related}) and a
-brief discussion~(\cref{s:discussion}).
 
@@
 
-\Fp{} is essentially scientific notation---\fp{} numbers are represented by packing a sign bit, an exponent, and a fraction part (also called the ``significand'' or ``mantissa'') into contiguous bit fields.
-The advantages of this representation are that allows for compact representation of both very small and very large numbers in a narrow range of bits---a 64-bit float can represent numbers as big as $2^{1024}$ and as small as $2^{-1074}$.
-This range is sufficient to represent numbers as small as Planck's constant or as big as Avogadro's constant with reasonable accuracy.
-
-\begin{figure}[h]
-  \label{fig:real_vs_float}
-  \includegraphics[width=\columnwidth]{ fig/real_vs_fp.png }
-  \caption{The real number line vs the points representable with \fp{}.}
+\begin{figure}[t]\centering
+  \includegraphics[trim=0 0 0 60,clip,width=0.95\columnwidth]{fig/real_vs_fp.png}
+  \caption{Floats are spread across the real number line.}
+  \label{f:real-vs-fp}
 \end{figure}
 
-The flip side is that not every real number within that wide range is representable as a \fp{} value---there is almost always some gap between the real number that we wish to represent and the \fp{} number that approximates that value.
-Whenever a value we have doesn't have a precise representation, we must \emph{round} that real value to the nearest \fp{} value, which introduces some error in our calculations.
-In addition to this error, the finite nature of \fp{} values means that we can run into some situations that wouldn't arise when working with real numbers.
-
-For example, if two numbers' exponents differ by too large an amount, adding or subtracting the smaller from the bigger will leave us with just the bigger number.
-Below we take the Planck constant and add it to Avogadro's constant and are left bit-for-bit with Avogadro's constant, as if we had never done any addition at all: there is no way to pack all the significant figures from true value that differs minutely from Avogadro's number into a 64-bit \fp{} number.
+\Fp{} numbers use a finite number of bits to represent a spectrum of points along the real number line~(\cref{f:real-vs-fp}).
+The implementation strategy is essentially scientific notation.
+A \fp{} number packs a sign bit, an exponent, and a fraction part (also called the ``significand'' or ``mantissa'') into a bitstring (typically 64 or 32 bits, but 16-bit and 8-bit formats are on the rise~\cite{klowerLowprecisionClimateComputing2021,fp8}).
+This representation supports very small and very large numbers in a narrow range of bits.
 
 \begin{lstlisting}[language = Julia]
-p = 6.62607015e-34
-a = 6.02214076e23
-
-println("planck:   $p : $(bitstring(p))")
-println("avogadro: $a  : $(bitstring(a))")
-println("p + a:    $(p + a)  : $(bitstring(p + a))")
+julia> prevfloat(typemax(Float64))
+1.7976931348623157e308
+julia> nextfloat(typemin(Float64))
+-1.7976931348623157e308
 \end{lstlisting}
 
-You can see in the output that \texttt{p + a} is bit-for-bit equivalent to \texttt{a}:
+The flip side is that most real numbers fall into the gaps between \fp{} numbers and must be rounded, which introduces error
+and can lead to surprising results.
+For example, adding the tiny Planck constant to the large Avogadro number results
+in Avogadro's number after rounding:
 
-\begin{lstlisting}
-planck:   6.62607015e-34 : 0011100100001 ... 001
-avogadro: 6.02214076e23  : 0100010011011 ... 111
-p + a:    6.02214076e23  : 0100010011011 ... 111
+\begin{lstlisting}[language = Julia]
+julia> planck = 6.62607015e-34
+6.62607015e-34
+julia> avogadro = 6.02214076e23
+6.02214076e23
+julia> avogadro + planck == avogadro
+true
 \end{lstlisting}
 
-This is a fairly extreme example; even when the numbers are closer in magnitude, most operations on \fp{} numbers induce loss of accuracy.
-
-\subsection{Floating-Point Standards}
-
-Prior to 1985, incomparable techniques for representing \fp{} values proliferated as mainframes and later personal computers emerged.
-This made writing portable \fp{} code challenging.~\cite{houghIEEEStandard7542019}
-IEEE 754~\cite{IEEEStandardBinary1985} standardizes \fp{} representation and operations across different architectures---these days, most machines and languages follow the IEEE 754 standard, and we make no effort to generalize our work beyond what IEEE 754 specifies.
+This is an extreme example, but many operations on \fp{} numbers closer in magnitude induce a loss of accuracy.
+Refer to the literature for more details, e.g.,~\cite{knuthArtComputerProgramming1997,torontoPracticallyAccurateFloatingPoint2014,mullerHandbookFloatingPointArithmetic2018}.
 
 \subsection{Exceptions and Exceptional Values}
+\label{s:exnvalue}
 
-IEEE 754 defines \emph{\fp{} exceptions}, which are events that occur when something goes wrong in a calculation.
-William Kahan says,
+The IEEE~754 \fp{} standard~\cite{IEEEStandardBinary1985}
+defines exceptions and exceptional values as the outcome of
+operations that have ``no single universally acceptable result''~\cite{p-draft-1997}.
+For example, dividing by zero and exceeding the \code{Float64} range
+both lead to exceptions:
 
-\begin{quote}
-An arithmetic exception arises when [there is] no result that would be acceptable universally.
-\end{quote}
+% julia> 2 / 0
+% Inf
+\begin{lstlisting}[language = Julia]
+julia> 0 / 0
+NaN
+julia> avogadro^avogadro
+Inf
+julia> log(0)
+-Inf
+julia> Inf + NaN
+NaN
+\end{lstlisting}
 
-For example, dividing by zero is an exception, as is exceeding a \fp{} format's range.
-These are exceptions because it is impossible to say what the right answer \emph{should} have been: perhaps that zero was meant to be a value that got really \emph{close} to zero,\footnote{The third kind of exceptional value is called a \emph{subnormal} value.} and suddenly rounded towards zero before division.
-Or maybe that value that got really big \emph{does} need to get that big, but you need to start using a larger bit width for your \fp{} values.
-Whatever, the case, there isn't a universally acceptable answer, so a \fp{} exception gets raised.
+IEEE~754 defers the question of how to handle such exceptions to application code.
+It is up to developers to watch for \Nan{} and \Inf{} (and also subnormal numbers)
+and implement an appropriate repair.
+Without guards, bugs happen.
 
-IEEE 754 doesn't force the programmer to always handle these exceptions, however, so we end up with \emph{exceptional values}, which are the results of exception-raising operations.
-The two values we work with in this paper are \Inf{}\footnote{There are actually two infinities: $+\infty$ and $-\infty$; they behave the same for our purposes, so we will refer to them just as ``\Inf{}''.} and \NaN{}, which result from overflow and nonsensical operations respectively.
-\Cref{fig:nan-gens} lists some of the situations that lead to exceptional values.
-
-\begin{figure}[ht]
-  \centering
-  \label{fig:nan-gens}
-  % \begin{tabular}{c | p{0.5 \columnwidth}}
-  \begin{tabular}{c | c}
-    \NaN{} & \Inf{} \\
-    \hline
-    $\tfrac{0}{0}$ & $\tfrac{x}{0},\, x \neq 0$ \\
-    $0 * \infty$ & $\log{(0)}$ \\
-    $\text{\texttt{NaN}} + x$ & $\infty + x,\, x \neq \text{\texttt{NaN}}$ \\
-    $\infty - \infty$ & Any result exceeding \\
-    & representation capacity
-  \end{tabular}
-  \caption{Examples of operations leading to exceptional values}
-\end{figure}
-
-For more details on exceptional values, see~\cite{knuthArtComputerProgramming1997,torontoPracticallyAccurateFloatingPoint2014} as well as the \textit{Handbook of Floating-Point Arithmetic}.~\cite{mullerHandbookFloatingPointArithmetic2018}
-
-\subsection{Lifetimes of Unhandled Exceptions}
+\subsection{Lifetime of an Unhandled Exceptions}
 \label{s:to-kill-a-fp}
 
-Exceptional values have a life cycle: instead a program-halting exception, an exceptional value comes into being at the point of an undefined operation and propagates through the rest of the computation.
-Often they go on to manifest in the output of the code, but sometimes they can ``die'' by disappearing silently.
-We characterize this life cycle in \cref{f:gpk}.
-
-\begin{figure}[ht]
+\begin{figure}[t]
   \includegraphics[width=\columnwidth]{fig/genpropkill-outline.png}
@@
 
-When a \NaN{} or an \Inf{} comes into being as the result of an undefined operation, we call this a generation, or \emph{gen} for short.
+Exceptional values have a lifetime: they are born, or \emph{generated}, by some operation; they \emph{propagate} through other operations; and they either appear in the program output, go out of scope, or get \emph{killed} by a numeric operation.
+\Cref{f:gpk} summarizes this \emph{gen}-\emph{prop}-\emph{kill} process.
+The gens and props are straightforward; see above for examples~(\cref{s:exnvalue}).
+The kills often arise from numeric comparisons (\code{<}, \code{=}, etc.),
+but exponents (\code{1\^NaN}) and over-eager matrix optimizations can kill
+exceptions~\cite{ddghlllprr-correctness-2022}.
 
-Most of the time, when any or all of the arguments are \NaN{}, the result is also a \NaN{}.
-We call this a propagation, or \emph{prop} for short.
-This is good and desirable, because it means that an arithmetic error---even if buried deep in an expression---will still be observable, allowing a user to find and debug it.
-
-However, there are a few places where exceptional values flow into an expression, but not out.
-We call these instances \emph{kills}, and these can lead to subtle logic bugs.
-For example, \emph{any} comparison ($<$, $>$, $=$, $\neq$, etc.) with a \NaN{} returns \texttt{false}.
+To illustrate TODO
 
@@
 \begin{lstlisting}[language = Julia]
-function bad_max(lst)
+function max1(lst)
   max_seen = 0.0
   for x in lst
+    # swap if x is not too small
     if ! (x <= max_seen)
-      # swap if new val greater
       max_seen = x
@@
 
-function good_max(lst)
+function max2(lst)
   foldl(max, lst)
@@
 
-Both \texttt{bad\_max} and \texttt{good\_max} compute the maximum in a list.
-However, the \texttt{<=} operator used in the \texttt{bad\_max} implementation does \emph{not} propagate \NaN{}s: \texttt{NaN <= 4} returns \texttt{false}, not \NaN{}.
+Both \code{max1} and \code{max2} compute the maximum in a list.
+However, the \texttt{<=} operator used in the \code{max1} implementation does \emph{not} propagate \NaN{}s: \texttt{NaN <= 4} returns \texttt{false}, not \NaN{}.
 Because of this, you might not get the result you expected:
@@
 \begin{lstlisting}[language = Julia]
-bad_max([1, 5, NaN, 4])     # returns 4
-good_max([1, 5, NaN, 4])    # returns NaN
+max1([1, 5, NaN, 4])     # returns 4
+max2([1, 5, NaN, 4])    # returns NaN
 \end{lstlisting}
 
-Not only is the result from \texttt{bad\_max} problematic for obscuring the fact that there was a \NaN{} in the list, the result is also wrong!
+Not only is the result from \texttt{max1} problematic for obscuring the fact that there was a \NaN{} in the list, the result is also wrong!
 The built-in \texttt{max} function correctly propagates \NaN{}s, so we can see from its return that we got a \NaN{} somewhere.
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
   \centering
@@
 \FT{} tracks down exceptional \fp{} values; once the location of a gen is found, Herbie could be employed to rewrite the arithmetic expression, if \fp{} error is the cause of the exceptional value.
-Odyssey\cite{misbackOdysseyInteractiveWorkbench2023} provides a helpful interactive interface to the tooling that Herbie provides.
+The Odyssey\cite{misbackOdysseyInteractiveWorkbench2023} workbench provides an interactive interface to Herbie.
 

### 2023-08-18 paper.tex
abstract, sec 1, sec 2

@@
 The implementation strategy is essentially scientific notation.
-A \fp{} number packs a sign bit, an exponent, and a fraction part (also called the ``significand'' or ``mantissa'') into a bitstring (typically 64 or 32 bits, but 16-bit and 8-bit formats are on the rise~\cite{klowerLowprecisionClimateComputing2021,fp8}).
-This representation supports very small and very large numbers in a narrow range of bits.
+A \fp{} number packs a sign bit, an exponent, and a fraction part (also called the ``significand'' or ``mantissa'') into a bitstring.
+Typical strings are 64 or 32 bits long, but 16-bit and 8-bit formats are on the rise~\cite{klowerLowprecisionClimateComputing2021,fp8}.
+This representation supports very small and very large numbers in a narrow range of bits:
 
@@
 IEEE~754 defers the question of how to handle such exceptions to application code.
-It is up to developers to watch for \Nan{} and \Inf{} (and also subnormal numbers)
+It is up to developers to watch for \Nan{}, \Inf{}, and subnormal numbers (underflow)
 and implement an appropriate repair.
-Without guards, bugs happen.
+This is no easy task, however, because of the approximations inherent to
+floating point.
+Even an apparently-safe division could result in a \Nan{} if its
+denominator gets truncated to zero.
+Some \Nan{}s might be spurious, others might be fatal, but in any event
+anticipating the various exceptions is a burden.
+All too often, exceptional values go unhandled and flow through the code.
 
@@
 
-Exceptional values have a lifetime: they are born, or \emph{generated}, by some operation; they \emph{propagate} through other operations; and they either appear in the program output, go out of scope, or get \emph{killed} by a numeric operation.
+Unhandled exceptional values have a lifetime: they are born, or \emph{generated}, by some operation; they \emph{propagate} through other operations; and they either appear in the program output, go out of scope, or get \emph{killed} by a numeric operation.
 \Cref{f:gpk} summarizes this \emph{gen}-\emph{prop}-\emph{kill} process.
@@
 The kills often arise from numeric comparisons (\code{<}, \code{=}, etc.),
-but exponents (\code{1\^NaN}) and over-eager matrix optimizations can kill
-exceptions~\cite{ddghlllprr-correctness-2022}.
+but exponents (\code{1{\string^}NaN}) and over-eager matrix
+optimizations~\cite{ddghlllprr-correctness-2022} can kill exceptions as well.
 
-To illustrate TODO
-
-Consider the following two functions to find the maximum value in a list:
+To illustrate the perils of killed exceptions, consider the following two ways
+of finding the maximum value in a list.
+The first compares numbers with \code{<=}
+while the second uses the built-in \code{max} function:
 
@@
 
-Both \code{max1} and \code{max2} compute the maximum in a list.
-However, the \texttt{<=} operator used in the \code{max1} implementation does \emph{not} propagate \NaN{}s: \texttt{NaN <= 4} returns \texttt{false}, not \NaN{}.
-Because of this, you might not get the result you expected:
+For lists with a \Nan{} inside, the functions can give different
+results because \code{<=} kills \Nan{}s whereas \code{max}
+propagates them:
 
 \begin{lstlisting}[language = Julia]
-max1([1, 5, NaN, 4])     # returns 4
-max2([1, 5, NaN, 4])    # returns NaN
+julia> max1([1, 5, NaN, 4]) 
+4.0
+julia> max2([1, 5, NaN, 4]) 
+NaN
 \end{lstlisting}
 
-Not only is the result from \texttt{max1} problematic for obscuring the fact that there was a \NaN{} in the list, the result is also wrong!
-The built-in \texttt{max} function correctly propagates \NaN{}s, so we can see from its return that we got a \NaN{} somewhere.
-It's not entirely obvious \emph{where} that \NaN{} came from.
+Not only is the result from \code{max1} problematic for obscuring the fact that there was a \NaN{} in the list, the result is arguably wrong!
+The result from \code{max2} at least shows that a \Nan{} was in the works, though in a realistic setting it may not be clear where the \Nan{} came from.
+Both versions would thus benefit from tools that track exceptions across their lifetime.
+FlowFPX can help.
 
@@
 
-\FlowFPX{} is a set of tools to aid in tracking down \fp{} exceptions.
-The two primary tools we cover in this paper are \FT{} and CSTG:
-
-\begin{itemize}
-\item \FT{} traces the lifetime of exceptional \fp{} values and enables inside-out fuzzing.
-\item CSTG visualizes the tracing output from \FT{} in compact form.
-\end{itemize}
-
-\GPUFPX~\cite{llsflg-hpdc-2023} is another component of \FlowFPX{} that can track \fp{} exceptions inside a GPU kernel.
-See the cited paper for more details on \GPUFPX.
-
-%% TODO julia proposed consistent NaN handling,
-%%  do we propagate the exact NaN
-%%  https://github.com/JuliaLang/julia/issues/48523
-%%
-%% Ashton: @Ben we simply propagate whatever we got from the operation. If
-%% Julia's behavior around an operator change, that should be preserved even
-%% when using FloatTracker
+\FlowFPX{} is a toolkit for tracking down \fp{} exceptions.
+The two primary tools we cover in this paper are \FT{}, which records lifetimes
+and enables fuzzing, and \CSTG{}s (more precisely, CSTGs), which visualize
+the flow of exceptions.
+\GPUFPX{} is a third component of \FlowFPX{} that tracks \fp{} exceptions inside GPU kernels~\cite{llsflg-hpdc-2023}.
 
@@
 
+FILL Consistent exception handling.
+
+https://github.com/JuliaLang/julia/issues/48523
+
+Ashton: @Ben we simply propagate whatever we got from the operation. If
+Julia's behavior around an operator change, that should be preserved even
+when using FloatTracker
+
+Ben: @Ashton tell the JuliaCon readers not just me!
+
 \subsection{Performance}

### 2023-08-18 paper.tex
sec 3
delete GPU figure because we have no implementation for the arrow

@@
 \usepackage{microtype}
+\usepackage{pifont}
 
@@
 \newcommand{\zerocode}[1]{\zerowidth{\code{#1}}}
+\newcommand{\genpropkill}{\emph{gen}-\emph{prop}-\emph{kill}}
+\newcommand{\bigcheckmark}{\ding{51}}
+\newcommand{\bigxmark}{\ding{53}}
+\newcommand{\tblnext}{\(\Rightarrow\)}
+\newcommand{\tblY}{\mbox{\bigcheckmark}}
+\newcommand{\tblN}{\scalebox{0.8}{\bigxmark}}
 
@@
   \includegraphics[trim=0 0 0 60,clip,width=0.95\columnwidth]{fig/real_vs_fp.png}
-  \caption{Floats are spread across the real number line.}
+  \caption{Floats are spread across the real number line}
   \label{f:real-vs-fp}
@@
   \includegraphics[width=\columnwidth]{fig/genpropkill-outline.png}
-  \caption{Gen, Prop, Kill: Lifetime of a \NaN{} value}
+  \caption{Gen, Prop, Kill: Lifetime of an exceptional value}
   \label{f:gpk}
@@
 Unhandled exceptional values have a lifetime: they are born, or \emph{generated}, by some operation; they \emph{propagate} through other operations; and they either appear in the program output, go out of scope, or get \emph{killed} by a numeric operation.
-\Cref{f:gpk} summarizes this \emph{gen}-\emph{prop}-\emph{kill} process.
+\Cref{f:gpk} summarizes this \genpropkill{} process.
 The gens and props are straightforward; see above for examples~(\cref{s:exnvalue}).
@@
 \FlowFPX{} is a toolkit for tracking down \fp{} exceptions.
-The two primary tools we cover in this paper are \FT{}, which records lifetimes
+The primary tools in this paper are \FT{}, which records lifetimes
 and enables fuzzing, and \CSTG{}s (more precisely, CSTGs), which visualize
@@
 
-\FT{} is implemented as a Julia library, and is available for download on JuliaHub.\footnote{\url{https://juliahub.com/ui/Packages/FloatTracker/dBXig/1.0.0}}
+\FT{} tracks exceptional values across their \genpropkill{} lifetime
+and can fuzz code for vulnerabilities by injecting a \Nan{} or \Inf{}
+as the result of an operation.
+\FT{} is implemented in Julia and is available on JuliaHub:
+
+\noindent\begin{center}\noindent\!\!\url{https://juliahub.com/ui/Packages/FloatTracker/dBXig}\!\!
+\end{center}
 
@@
 
-\FT{}'s first purpose is to monitor \fp{} operations for exceptional value lifetime events.
-\FT{} interposes on arithmetic operations and standard library calls that involve a \fp{} value and logs a stack trace to that point whenever an exception is generated, propagated, or killed.
-\FT{} can also log the arguments to the suspect call, so you can get \emph{normal, concrete} values to help discover what is going wrong.
-A programmer can then use these stack traces to pinpoint where an exceptional value like \NaN{} came into being or where a guard against a \NaN{} kill is needed.
+\begin{figure}[t]\centering
+  \begin{tabular}{ccccc}
+    Exn. Input? & \tblnext & Exn. Output? & = & Event       \\ \hline
+    \tblN       & \tblnext &  \tblY{}     & = & {gen}  \\
+    \tblY       & \tblnext &  \tblY{}     & = & {prop} \\
+    \tblY       & \tblnext &  \tblN{}     & = & {kill} \\
+  \end{tabular}
+  \caption{How to classify operations that see exceptions}
+  \label{fig:lifetime-class}
+\end{figure}
 
-All the logs that \FT{} captures are straight from Julia's \texttt{stacktrace()} function.
-\FT{} writes these logs to three files: one file for each kind of event in a \fp{} exception's lifetime.
-This way, if a user of \FT{} is attempting to determine where a \NaN{} or an \Inf{} came from, they can look in the file for all the gen events.
-Likewise, if the user wants to find where \NaN{}s get killed, then they can look in the kills file.
+\FT{} monitors \Nan{} and \Inf{} exceptions by overloading arithmetic
+operations and logging key events.
+When the input to an operation is exception-free but the output is
+exceptional, the operation is a {gen} event~(\cref{fig:lifetime-class}).
+When the input and output contain exceptions, the operation is a {prop}
+event.
+And when the input contains an exception but the output does not,
+the operation is a {kill} event.
+Put together, the log of all {gen}s, {prop}s, and {kill}s
+sheds light of how various exceptions traveled across the program.
+The logs also record the call context (stack trace) and the arguments to the operation as a starting point for debugging efforts.
+
+\FT{} writes logs to three files: one for {gen} events, one for {prop}s, and one for {kill}s.
+This way, discovering where a \Nan{} came from is a matter of sifting through the {gen} file.
+Users can also lower the overhead of logging by turning it off for prop events.
+
+The instrumentation works through custom \fp{} types: \code{TrackedFloat64},
+\code{TrackedFloat32}, and \code{TrackedFloat16}.
+Developers must opt in to \FT{} by wrapping numbers in a custom type.
+From then on, tracking is automatic and extends transitively to all outputs
+of tracked operations.
+Multiplying a \code{Float64} value with a \code{TrackedFloat64}, for example, yields
+a \code{TrackedFloat64} to continue the logging trail.
 
@@
 
-The second role of \FT{} is to fuzz code from the inside out.
-While can fuzz a program by giving it random inputs, this might not help you find \emph{all} the places where an exceptional value could crop up during a calculation.
-Demmel et. al suggest fuzzing by injecting exceptional values at intermediate calculations;~\cite{ddghlllprr-correctness-2022} this is the approach that we enable with \FT{}.
+Operator overloading gives \FT{} a powerful way to fuzz code from the inside out.
+Each overloaded function serves as a hook where \FT{} can decide whether to observe
+the operation or step in, discard the original result, and return an exception instead.
+Injecting faults in a random way~\cite{hamlet1994random}, also known as fuzzing,
+is a useful way to discover vulnerabilities in a large codebase.
+Demmel et~al. propose essentially the same idea for BLAS and
+LAPACK~\cite{ddghlllprr-correctness-2022}.
 
-\FT{}'s exception tracking infrastructure offers us a simple way to inject exceptional values wherever we want in our code: we are already interposing on \fp{} operations to monitor for exceptional lifetime values, so we can inject a \NaN{} at any of these points.
-
-Since the injection happens at random, \FT{} also provides a way to save \NaN{} injections so that injections that exercise bugs can be replayed to examine the bug repeatedly and to verify that hardening efforts have indeed eliminated the vulnerability.
-
-We have a global configuration system that allows us to turn on fuzzing and configure where, when, and how many \NaN{}s to inject, as well as whether or not to record or replay a recording:
+\FT{} exposes several parameters to let developers control the fuzzer:
 
 \begin{itemize}
-\item \texttt{replay::String} Name of a file to read an injection recording from. If set, this option makes all others irrelevant.
-\item \texttt{record::String} Name of a file to write injection events to for later replay.
-\item \texttt{active::Bool} Inject only if this is true.
-\item \texttt{n\_inject::Int64} Inject at most this many \NaN{}s.
-\item \texttt{odds::Int64} Inject if \texttt{rand(odds) == 1}.
-\item \texttt{functions::Array} Inject if we are within the dynamic extent of any of these functions.
-\item \texttt{libraries::Array} Inject if the top-most stack frame\footnote{After stripping off the stack frames that come from \FT{} itself.} comes from any of these libraries.
+\item \texttt{odds::Int64} inject if \texttt{rand(odds) == 1}.
+\item \texttt{n\_inject::Int64} upper bound on the number of \Nan{}s to inject.
+\item \texttt{active::Bool} fuzz only when set to true.
+\item \texttt{functions::Array{String}} limit fuzzing to the dynamic extent of the listed functions.
+\item \texttt{libraries::Array{String}} limit fuzzing to functions from the following libraries
 \end{itemize}
 
-Then, each time we intercept a \fp{} operation, we check to see if we should inject a \NaN{}.
-First we check if we're replaying a recording: if we are, just do as the recording says.
-Otherwise, we check that the injector is \texttt{active}, we have \NaN{}s remaining to inject (\texttt{n\_inject}), and we roll an \texttt{odds}-sided die to see if we should inject.
-Finally, check the current stack trace against the lists of functions and libraries to inject in.
-If all looks good, we inject.
+When fuzzing reveals an error, the next step is to craft a regression test to
+guide repairs.
+\FT{} therefore records the sequence of injections that it makes during a fuzzing
+run and enables a replay of any recording after the fact.
+Replay runs proceed deterministically so that developers can harden
+their code and check that the fixes remove the error.
 
@@
 
-While other tools like profilers might accomplish the tracking that \FT{} does by writing extensions to the language's compiler or runtime, \FT{} works entirely in user land.
-\FT{} provides custom types---\texttt{TrackedFloat16}, \texttt{TrackedFloat32}, and \texttt{TrackedFloat64}---that wrap their corresponding \texttt{Float16}, \texttt{Float32}, and \texttt{Float64} counterparts.
-\FT{} also defines methods for every function in the \texttt{Base} module.
+\FT{} takes advantage of Julia's operator overloading to track exceptions and
+fuzz for vulnerabilities.
+For example, below is a variant of \code{+} that is overloaded
+for \code{TrackedFloat64}.
+It calls the basic \code{+} (or injects a \Nan{}) and checks for exceptions
+before returning:
 This lets \FT{} intercept all \fp{} operations involving \texttt{TrackedFloat} types.
@@
 
-This new method does a few things:
+Every arithmetic operation requires a similar overloading.
+Thus, the implementation of \FT{} uses a metaprogramming technique
+adapted from \texttt{Sherlogs.jl}~\cite{kMilanklSherlogsJl2021}
+to abstract over the common patterns.
+For every binary operation, the library creates an overloading similar
+to the one for \code{+} above.
+Unary operations and others work analogously.
+This approach saved thousands of lines of code.
+The implementation weighs in at 218 lines and defines 645 overloaded function;
+assuming 5 lines of code per function, a handwritten version would require over 3,000 lines.
 
-\begin{enumerate}
-\item First it checks if it should inject a \NaN{} here.
-\item If no \NaN{} is to be injected, it unwraps the \texttt{TrackedFloat}s to get to the primitive \texttt{Float} value, and computes the result normally.
-\item It examines the arguments as well as the return value, and then uses \cref{fig:lifetime-class} to determine the type of event and logs it if it is of interest.
-\item It wraps the result in a new \texttt{TrackedFloat} value.
-\end{enumerate}
-
-\begin{figure}[h]
-  \centering
-  \begin{tabular}{ccc}
-    Arguments & Result & Event Type \\
-    \hline
-    No \NaN{} & \NaN{} & Gen \\
-    \NaN{} & \NaN{} & Prop \\
-    \NaN{} & No \NaN{} & Kill \\
-  \end{tabular}
-  \label{fig:lifetime-class}
-  \caption{Exceptional value event classification rules}
-\end{figure}
-
-The same classification rules apply to \Inf{} and would apply to any other value of interest.
-
-These new methods had to be defined for every combination of tracked/untracked arguments;
-fortunately Julia's metaprogramming abilities made it simple for us to scale without the code becoming unwieldy.
-We made a macro that first loops through all the \texttt{TrackedFloat} types that we wished to implement, and then looped through all of the operators that we wished to support.
-This approached saved us over 10x the lines a non-macro approach would have required.\footnote{We implemented overrides for 645 function/type combinations.
-At roughly 5 lines per variant, this is 3225 lines of code.
-Our implementation at time of writing weighs in at 218 lines of code, which is \emph{including} whitespace and comments.}
-This approach was inspired by \texttt{Sherlogs.jl}.\cite{kMilanklSherlogsJl2021}
-
-\subsection{CSTG}
+\subsection{Stack Graphs}
 \label{s:cstg}
@@
 \FT{} can produce copious amounts of log files which can be challenging to sift through manually.
-Coalesced Stack Trace Graphs, or CSTGs~\cite{humphreySystematicDebuggingMethods2014} provide a way to visualize large amounts of stack traces in a compact form.
-CSTG is a stand-alone tool and was not designed with \FT{} specifically in mind.
-However, it pairs well with \FT{} and makes analyzing the log files easier.
+Coalesced stack trace graphs (CSTGs or \CSTG{}s for short) provide a way to visualize large amounts of stack traces in a compact form~\cite{humphreySystematicDebuggingMethods2014}.
+This techinque pairs well with \FT{} and makes analyzing the log files easier.
 
-CSTG takes a set of stack traces and makes a graph: for each stack trace, it creates a node for every frame, and then creates or increments an edge between two adjacent frames, as seen in \cref{fig:cstg_demo}.
+\Cref{fig:cstg_demo} illustrates the construction of a \CSTG{} from a collection of stack traces.
+Each trace on the left contributes nodes and edges to the graph on the right.
+Repeated edges get emphasized with darker lines and larger counts.
 
@@
   \centering
-  \includegraphics[width=3in]{./fig/cstg_static_diagram.png}
-  \caption{CSTG makes a graph by connecting adjacent stack frames}
+  \includegraphics[width=0.9\columnwidth]{./fig/cstg_static_diagram.png}
+  \caption{From stack traces (left) to \CSTG{} (right)}
   \label{fig:cstg_demo}
@@
 
-These graphs are meant to be read bottom-up: the bottom nodes are the roots where exceptional values manifested.
-Higher nodes in the graph show function calls progressively closer to the top-level call that spawned them.
-Thick lines and high numbers indicate flows with the highest volume leading to errors---though thin areas are still worth investigating.
-CSTG provides a map to aid in contextualizing a stack trace in a log file.
+Reading bottom-up, a \CSTG{} based on the \emph{gen} events in a program
+highlights the contexts that frequently produced exceptional values.
+Using~\cref{fig:cstg_demo} as an example, \code{Frame D} produced
+three exceptions, two of which arose under \code{Frame C}.
+In a large program with many exceptions, \CSTG{}s offer a way to prioritize
+debugging efforts: go for the heavily-trodden paths first.
 
@@
 
-% HELP: @ben, I'm not sure what paper you mean by "Julia CPU"
 Many programs offload work onto GPUs, which are no less susceptible to \fp{} exceptions than CPUs.
-In some regards they are \emph{more} susceptible because GPUs do not have the same exception-handling mechanisms that CPUs do.
-Thus, tools to catch \fp{} exceptions on the GPU are essential.
-\FT{} is constrained to CPU workloads; however, there is a companion tool called \GPUFPX{}~\cite{llsflg-hpdc-2023} that can help diagnose \fp{} exceptions on the GPU.
-\GPUFPX{} can detect floating-point exceptions on the GPU that GPU code cannot do on its own.
-\GPUFPX{} reports these exceptions back to the CPU for further analysis.
-
-\FT{} and \GPUFPX{} can cover the entire computational stack, as seen in \cref{fig:ft_and_gpu_fpx}.
-Work remains to tighten the integration between \FT{} and \GPUFPX{} to reduce manual work to employ both tools for heterogeneous computation.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=3in]{./fig/float_tracker_context_boxes.png}
-  \caption{FloatTracker and GPU-FPX compliment each other}
-  \label{fig:ft_and_gpu_fpx}
-\end{figure}
+In fact, GPU programs are worse off because they lack exception-handling mechanisms from the CPU world~\cite{llg-soap-2022}.
+Since \FT{} instruments Julia programs, it cannot help directly;
+however, the companion tool \GPUFPX{} instruments GPU kernels to detect and report
+\fp{} exceptions~\cite{llsflg-hpdc-2023}.
+Together, \FT{} and \GPUFPX{} provide insights for accelerated programs.
 
@@
 
-In this section we'll discuss some case studies where tools from \FlowFPX{} aided in finding or fixing \fp{} issues.
-\Cref{s:sw} covers an ocean simulation that produces \NaN{}s in the output---a common and frustrating occurrence for many a scientist.
-\Cref{s:ode} puts the popular \texttt{OrdinaryDiffEq} solver library under a fuzzer to harden it against bad behavior in the presence of exceptional \fp{} values.
-\Cref{s:finch} fuzzes another library---this time a library for specifying partial differential equations (PDEs).
-Finally, in \cref{s:ocean}, we examine another ocean simulation library---this time the library runs on the GPU.
-This puts it out of reach of \FT{}, but \GPUFPX{} can handle it.
-In addition to the case studies that we did ourselves, \cref{s:safari} is an ``in the wild'' usage of \FT{} that helped researchers uncover the source of a \NaN{} in their own code without help from us.
+\FlowFPX{} has helped to debug exceptions and fuzz for issues in a variety
+of settings, some synthetic and some realistic.
+The case studies include a shallow water simulation, the
+\texttt{OrdinaryDiffEq} solver, and a Bayesian inference library.
 
@@
 
-\texttt{ShallowWaters}~\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019,klowerLowprecisionClimateComputing2021} simulates the flow of water over a seabed.
-This is not a general shallow waters simulation library---it is designed primarily to showcase numerical methods involving 16-bit floats.
+\texttt{ShallowWaters.jl} simulates the flow of water over a
+seabed~\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019}.
+The library has dozens of parameters that a scientist can experiment with.
+One notable parameter is the Courant-Friedrichs-Lewy (CFL) number, which roughly describes the size of the time step to take in running the simulation.
+A small CFL number makes the simulation run slowly, but accurately; a large number speeds it up but loses precision
+because the system does not get enough time to propagate information.
 
-The Courant-Friedrichs-Lewy (``CFL'') parameter roughly describes the size of the time step to take in running the simulation.
-Normally this value sits between 0 and 1---smaller values mean the simulation runs slower but with higher accuracy, whereas higher values mean the simulation takes bigger time steps and consequently runs faster, but results might be inaccurate.
-The primary reason for problems arising comes from information not having enough time to propagate through the simulation before the next time step is taken.
+Normal CFL values range between zero and one.
+X
 
@@
   \centering
-  \includegraphics[width=3in]{fig/sw_nan_cstg_clean.png}
+  \includegraphics[width=0.96\columnwidth]{fig/sw_nan_cstg_clean.png}
   \caption{CSTG of \NaN{} generation logs from ShallowWaters.}
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=3in]{fig/sw_inf_cstg_clean.png}
-  \caption{CSTG of \Inf{} generation logs from the same run as~\cref{fig:sw_nan_cstg}.}
-  \label{fig:sw_inf_cstg}
-\end{figure}
+% \begin{figure}[t]
+%   \centering
+%   \includegraphics[width=3in]{fig/sw_inf_cstg_clean.png}
+%   \caption{CSTG of \Inf{} generation logs from the same run as~\cref{fig:sw_nan_cstg}.}
+%   \label{fig:sw_inf_cstg}
+% \end{figure}
 

### 2023-08-18 paper.tex
sec 4

@@
 \begin{figure}[t]\centering
-  \includegraphics[trim=0 0 0 60,clip,width=0.95\columnwidth]{fig/real_vs_fp.png}
+  \includegraphics[trim=10 0 10 0,clip,width=0.95\columnwidth]{fig/real_vs_fp.pdf}
   \caption{Floats are spread across the real number line}
@@
 
-Normal CFL values range between zero and one.
-X
+Normal values for the CFL number range between zero and one.
+With a CFL of 0.9, the shallow water simulation produces the graph on the left half
+of~\cref{fig:sw_nans} showing current speed in a rectangular basin.
+Raising the CFL too high, to 1.6, causes trouble
+(\cref{fig:sw_nans}, right) with large, white regions where the
+current speed is \NaN{} instead of a normal value.
 
-It is not uncommon to encounter cases where setting the CFL parameter too high results in \NaN{}s in the result.
-Driving the CFL parameter low enough to excise the \NaN{}s can make the simulation too long and unwieldy to be comfortable to use.
-Identifying where \NaN{}s begin to arise can help point the way to where some mathematical refinement can be used to make the simulation both fast and accurate enough for practical use.
-
-We dialed the CFL parameter up to extraordinarily high values and we started seeing \NaN{}s in the result.
-\code{ShallowWaters} detected the instability, but not before producing some bad output.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=3in]{./fig/shallow_waters_cfl_diff.png}
-  \caption{Effect of high CFL parameters: on the left, normal output. On the right: \NaN{}s cause white holes and gaps in the result.}
-  \label{fig:sw_nans}
-\end{figure}
-
-The white regions in the right graph of \cref{fig:sw_nans} are regions where \NaN{}s crept into the computation and destroyed the results.
-In one regard, this is not as bad as a \NaN{} kill---knowing that there are \NaN{}s in the computation means that we have a chance to fix it, whereas a \NaN{} kill might silently give us the wrong answer.
-Incidentally, we know that there were no \NaN{} kills in this test all the events in our kill logs are from calling the \texttt{isnan} function---a clear sign that the program is deliberately taking action on the presence of a \NaN{}.
-
-Finding the sources of \NaN{}s can be tricky.
-We expected \NaN{}s in this case because we set the CFL parameter up so high, but tracking down the sources of \NaN{}s generally is a difficult and time-consuming problem.
-
-\FT{} solves this problem for us by automatically tracking all sources of potential \NaN{} generation.
-And in this case \texttt{ShallowWaters} makes this doubly easy:
-\texttt{ShallowWaters} was designed to showcase physics simulations using variable levels of \fp{} precision, and allows you to specify a \fp{} type to use in the computation.
-We simply asked it to use our \texttt{TrackedFloat32} type, and then the entire simulation ran under the supervision of FloatTracker.
+Running \FT{} on the high-CFL simulation reveals where the simulation drifted
+from numbers to \NaN{} exceptions.
+The first step in applying \FT{} is to convert relevant floats to tracked
+floats, e.g., \code{Float32} to \code{TrackedFloat32}.
+For shallow waters, this step is easy because the simulation is parameterized by
+a \fp{} type for internal use.
+Swapping in a tracked type is enough:
 
@@
 run_model(T=TrackedFloat32,
-          cfl=10, Ndays=100, nx=100, L_ratio=1,
+          cfl=10,
+          nx=100,
+          Ndays=100,
+          L_ratio=1,
           bc="nonperiodic",
@@
 
-This produced almost 17000 lines of logs in the gens file. Below is a sample stack trace from these logs: (formatted for clarity)
+With tracking, the simulation runs as before, producing an ugly graph.
+It additionally outputs logs for all gen, prop, and kill events
+as they happen.
+Below is one event from the gen logs:
 
-\begin{verbatim}
+\begin{lstlisting}
 -([-Inf, -Inf])   FT/TrackedFloat.jl:106
@@
 top-level
-\end{verbatim}
+\end{lstlisting}
 
-We can see that the \NaN{} appeared as the result of $-\infty - -\infty$ (first line) and that subtraction happened in the \texttt{momentum\_u!} function on line 246 of the \texttt{rhs.jl} file (second line).
-Thus, with little effort we were able to track down where the \NaN{}s were originating.
+We can see that the \NaN{} appeared as the result of subtracting two infinities (\code{-Inf - -Inf}).
+The trace further shows that this subtraction happens inside the function \texttt{momentum\_u!} on line 246 of the \texttt{rhs.jl} file.
 
-Now the question is: where did the $\infty$ come from?
-We can look at the file of the $\infty$ gens where we see the following: (formatted for clarity)
+This solves the mystery of where one \NaN{} came from, but raises a new question
+about the source of the \Inf{} value.
+The logs for \Inf{} gens have an answer:
 
-\begin{verbatim}
+\begin{lstlisting}
 ^([-1.515f31, 2])          FT/TrackedFloat.jl:138
 literal_pow()              intfuncs.jl:325
-…
+...
 materialize(^)             broadcast.jl:860
 top-level getproperty(...) examples/sw_nan_tf.jl:14
-…
-\end{verbatim}
+..
+\end{lstlisting}
+%
+This \Inf{} came from an exponent that overflowed the float type (\code{-1.515e31\string^2}).
+\FT{} has shown exactly which numeric values in which operation
+caused exceptions to occur.
 
-It looks like the $-\infty$ came from $-1.515e31^2$---no wonder a Float32 type couldn't handle something that large.
-Now we have some concrete values to work with: instead of wondering where an opaque \Inf{} or a \NaN{} came from, we have \emph{specific} values that a domain expert can use to diagnose the problem.
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=3in]{./fig/shallow_waters_cfl_diff.png}
+  \caption{Raising the CFL number creates white gaps due to \NaN{}s}
+  \label{fig:sw_nans}
+\end{figure}
 
-\subsubsection{CSTGs Help Us Get a Bigger Picture}
+\subsubsection{Stack Graphs for a Bigger Picture}
 
-Running FloatTracker can produce a lot of logs.
-To get a quick overview of the most common paths to problems, we can use CSTG to get an overview of what stack frames appear the most often.
+While the logs for shallow waters contain useful information,
+there is an overwhelming amount of it.
+There are over ten thousand lines in the gen file alone.
+Converting these logs to a stack graph gives a quick overview
+of the most common paths to exceptional values.
 
-We can produce a CSTG of both the \Inf{} as well as the \NaN{} logs.
-\Cref{fig:sw_nan_cstg} and \cref{fig:sw_inf_cstg} show what the two graphs look like.\footnote{As with the call stack logs, these have been edited for clarity.
-We only remove a few boxes in between—mostly due to space constraints in this paper.}
+\Cref{fig:sw_nan_cstg} presents the stack graph for \NaN{} gen events
+in shallow waters with the high CFL number.
+Reading bottom-up, every \NaN{} came from calls to the \code{-} and \code{+} operations.
+Calls to \code{+} account for most of the \NaN{}s.
+These \NaN{}s arose in two different contexts: a small number (30)
+occurred within a momentum calculation, while the rest (162)
+occurred within a continuity step.
+Moving to the top of the graph, it shows that the function \code{run\_model} drove
+the entire simulation.
+With this overview of the program, a promising next step is to guard against \NaN{}s
+in the momentum and continuity functions.
 
@@
   \includegraphics[width=0.96\columnwidth]{fig/sw_nan_cstg_clean.png}
-  \caption{CSTG of \NaN{} generation logs from ShallowWaters.}
+  \caption{Stack graph for \NaN{} gens in shallow waters}
   \label{fig:sw_nan_cstg}
@@
 
-We can see in \cref{fig:sw_nan_cstg} that a thick line goes from the top-level call down to a call to \texttt{+} goes down the right side of the graph.
-This indicates that the majority of the flows took this path.
-That means we might be able to fix the majority of our problems by inspecting that particular flow.
-\Cref{fig:sw_inf_cstg} shows us that all but two flows started with a call to \texttt{\^}---a promising place to start looking for improvements.
+\subsubsection{Stack Graph Differences}
 
-\subsubsection{CSTG Diffing}
+Graph diffing works well for \CSTG{}s; it shows how flows in the program evolved
+from one stage to another.
+\Cref{fig:cstg_diff_demo} illustrates one diff in the context of \NaN{} gens for
+shallow waters.
+It compares the first 10\% of gen events to the latter 90\% of gens.
+The positive numbers and green lines indicate flows that are new in the latter part.
+The negative numbers and red lines show flows that disappeared in the latter part.
 
@@
 The diff can be seen in \cref{fig:cstg_diff_demo}.
+A domain expert might use these clues to find where an instability started in
+the first part of the program.
+In this case, the function \texttt{momentum\_u!} showed up only in the
+beginning of the logs; it may be an effective point to check for \Nan{}s.
 
@@
   \centering
-  \includegraphics[width=3in]{./fig/cstg_diff_pretty.png}
-  \caption{Diff between the early and latter stages of ShallowWaters}
+  \includegraphics[width=0.96\columnwidth]{./fig/cstg_diff_pretty.png}
+  \caption{Stack graph diff}
   \label{fig:cstg_diff_demo}
 \end{figure}
-
-The green lines indicate flows that were \emph{new} in the latter 90\% of the logs.
-The red lines indicate flows that were \emph{only in} the first 10\% of the logs.
-What this means depends on the particular program; a domain expert can use these clues to e.g.
-find where instability started in the first part of the program, and where that instability got reinforced later on.
-In this case, we can see that \texttt{momentum\_u!} only showed up in the beginning of the logs---we might look there for ways to prevent instability from beginning.
 
@@
 
-To exercise the fuzzing abilities of \FT{}, we tried fuzzing \texttt{NBodySimulator}\footnote{\url{https://github.com/SciML/NBodySimulator.jl}} from the SciML team, which simulates the gravitational interactions between two or more massive bodies.
-However, when we used \FT{}, we got no \fp{} operations.
-We discovered that \texttt{NBodySimulator} merely sets up the system of equations to solve, then hands all the work of running the simulation off to the \texttt{OrdinaryDiffEq}\footnote{\url{https://github.com/SciML/OrdinaryDiffEq.jl}} library from the same team.
+With the targeted fuzzing abilities of \FT{}, we set focus on the
+\texttt{NBodySimulator} package from the SciML
+team.\footnote{\url{https://github.com/SciML/NBodySimulator.jl}}
+Fuzzing discovered zero bugs and zero exceptional events.
+In fact, fuzzing injected zero exceptional values whatsoever, even with maximum odds.
+The reason for these null results, it turns out, is that the nbody
+simulation merely creates a problem for the \code{OrdinaryDiffEq}
+solver.\footnote{\url{https://github.com/SciML/OrdinaryDiffEq.jl}}
+All the \fp{} operations were in another library.
 
-We configured \FT{} to inject a single \NaN{} randomly within the scope of the \texttt{NBodySimulator} and \texttt{OrdinaryDiffEq} libraries.
-We found a run where something curious happened: the library reported that it had detected a \NaN{} and would then exit.
+Fuzzing on \code{OrdinaryDiffEq} led to a curious situation.
+The library itself reported a \NaN{} and printed a message that it would exit.
 However, after printing that message, the program went into an infinite loop.
+Using \CSTG{}s to guide our search, we found a \NaN{} kill that manifested repeatedly in the logs.
+The stack traces for that kill originated from inside the file \texttt{solve.jl}:
 
-Using CSTGs to guide our search, we found a \NaN{} kill that manifested repeatedly in the logs.
-The stack traces for that kill originated from inside the \texttt{solve.jl} file in the \texttt{OrdinaryDiffEq} library.
-
-\begin{verbatim}
+\begin{lstlisting}
 <([NaN, 3.0e6])  at FT/TrackedFloat.jl:193
 solve!           at ODE/solve.jl:515
-…
-\end{verbatim}
+...
+\end{lstlisting}
 
-The relevant part of \texttt{solve.jl} looks like this:
+The relevant part of \texttt{solve.jl} contains a pair of loops.
+With injection, the variable \code{tdir} holds a \NaN{},
+stopping all productivity:
 
@@
 
-The problem here is on line 2 when \texttt{tdir} is \NaN{}: multiplication propagates \NaN{}s, and comparison kills it, so the condition on the inner \texttt{while} loop is \emph{always} false.
-No work got done in the inner loop, and so the size of the \texttt{time\_stops} vector never decreases.
+In more detail, the \NaN{} for \code{tdir} propagates though the multiplication
+and gets killed by the \code{<} comparison.
+Hence the condition for inner \code{while} loop is always false,
+which means the outer loop never ends up with an empty list.
 
-Our kill logs led us right to this line and let us know which of the arguments to \texttt{<} was a \NaN{}.
-This is a perfect example of how \NaN{} kills can influence control flow to go awry.
-In this case, the problem was apparent besides what our logs told us and it manifested as an infinite loop.
-More dangerous cases can occur when the influence of a bad comparison is not as readily observable.
-In any case, \FT{} can help monitor for kills and help fuzz to find kills before they arise in production.
-
-We were able to use the injection recording to reproduce the bug in the GitHub issue that we opened.\footnote{\url{https://github.com/SciML/OrdinaryDiffEq.jl/issues/1939}}
+This is a real-world example of how \NaN{} kills can affect control flow, and we filed
+an issue for it.\footnote{\url{https://github.com/SciML/OrdinaryDiffEq.jl/issues/1939}}
+Fortunately, the problem in this case is benign as the code was already trying to halt.
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}
   BoundsError: attempt to access 1-element
    Vector{Int64} at index [2]
-\end{verbatim}
+\end{lstlisting}
 
-Defensive programming solves the issue.
+Before accessing the input, Finch needed to check for \NaN{}s.
 The second place was in setting bounds for the
@@
 Additionally, \FT{} and \CSTG{}s have been useful for identifying \NaN{}s that
-appear in unstable systems.
+appear in unstable heat simulations written in Finch.
 %% advection2d, high cfl and high T (time) led to NaN during simulation
 
-\subsection{\GPUFPX{}: Oceananigans}
+\subsection{Oceananigans}
 \label{s:ocean}
@@
 
-\subsection{RxInfer: No Help Needed}
+\subsection{RxInfer}
 \label{s:safari}
 
-\FT{} has proven to be useful for more than its creators---we discovered an open issue in the \code{RxInfer.jl} library with an issue revolving around \NaN{} detection.\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}}
-Even though we were unable to provide hands-on guidance as the code in question was proprietary, the authors of \code{RxInfer.jl} were able use \FT{} themselves, and within a day they tracked down the location of a \NaN{} gen.
+We discovered an open issue in the \code{RxInfer.jl} library related to \NaN{} detection
+and suggested \FT{} to the developers.\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}}
+Within a day, they were able to track down the location of an important \NaN{} gen.
+This success came with little input from our end; in fact, the application program was
+closed-source.
+We merely explained how to use \FT{} by wrapping inputs in a tracked type.
 
@@
 
-\subsection{Floating-Point Error Analysis}
+\paragraph{Error Analysis}
 
@@
 
-\subsection{Diagnosing floating-point exceptions}
+\paragraph{Diagnosing floating-point exceptions}
 
@@
 
-\subsection{Stack Trace Graphs}
+\paragraph{Stack Graphs}
 
@@
 
-\subsection{Floating-Point Exceptions on the GPU}
+\paragraph{GPU Exception Tracking}
 
@@
 
-\subsection{Future Work}
-
-\FlowFPX{} is a growing project and we welcome feedback from domain experts
-on how to best suit their workflow.
-If you try any of our tools, please send us comments, suggestions, and bug reports via the tools' repositories on GitHub.
-
 \section{Acknowledgments}

### 2023-08-18 paper.tex
sec 5-end

@@
 a \code{TrackedFloat64} to continue the logging trail.
+
+Crucially, the tracked version of an exception, say a \NaN{}, points
+to the original exceptional value.
+This enables techniques such as \NaN{}-packing, or otherwise
+using the payload in creative ways.
+Not all Julia operations currently preserve \NaN{}s
+bit-for-bit,\footnote{\url{https://github.com/JuliaLang/julia/issues/48523}}
+but \FT{} does its part.
 
@@
 
-\paragraph{Error Analysis}
+\subsection{Error Analysis}
 
-Demmel et al.\cite{ddghlllprr-correctness-2022} examine \fp{} exceptional value handling in the BLAS and LAPACK libraries and propose solutions to inconsistencies.
-They describe a proposal to pack numeric codes into an \texttt{INFO\_ARRAY} that mirrors our gen-prop-kill framework.
-They also suggest a fuzzing technique that we realize in \FT{}:
-instead of fuzzing the \emph{inputs} to a program, we fuzz by randomly injecting exceptional values at \emph{any} point where there is an arithmetic operation.
-Demmel suggests that this approach is useful since any \fp{} operation could lead to an exceptional value if the inputs are in dangerous ranges.
+Demmel et al.\cite{ddghlllprr-correctness-2022} examine \fp{} exceptional value handling in BLAS and LAPACK, identify several inconsistencies, and propose an API for debugging and adding determinism.
+The proposal includes an extension to the \texttt{INFO\_ARRAY} parameter with fields that record \genpropkill{} information.
+It also includes fuzzing, which we realize in \FT{}.
 
-Toronto and McCarthy\cite{torontoPracticallyAccurateFloatingPoint2014} develop a methodology for assessing where arithmetic operations are liable to magnify \fp{} error.
-They include error visualization and work on how to avoid error-magnifying ``badlands''.
-They focus on rewriting arithmetic expressions to avoid error which could lead to exceptional values, rather than on the exceptional values themselves.
+Toronto and McCarthy~\cite{torontoPracticallyAccurateFloatingPoint2014} propose a test-driven method for detecting numeric error: plot the results of an expression on a range of inputs and look for sharp deviations, or \emph{badlands}.
+They point out several ways to rewrite code to avoid badlands by rewriting code in a semantics-preserving way.
 
-In a similar vein, Herbie\cite{panchekhaAutomaticallyImprovingAccuracy2015} is a tool for automatically rewriting arithmetic expressions to reduce \fp{} error.
-\FT{} tracks down exceptional \fp{} values; once the location of a gen is found, Herbie could be employed to rewrite the arithmetic expression, if \fp{} error is the cause of the exceptional value.
+Herbie~\cite{panchekhaAutomaticallyImprovingAccuracy2015} automatically rewrites arithmetic expressions to reduce \fp{} error.
+This would combine well with \FT{}: first identify the location of a \NaN{}, ask Herbie to find a repair.
 The Odyssey\cite{misbackOdysseyInteractiveWorkbench2023} workbench provides an interactive interface to Herbie.
 
-\paragraph{Diagnosing floating-point exceptions}
+\subsection{Diagnosing floating-point exceptions}
 
-The Julia library \texttt{Sherlogs}~\cite{kMilanklSherlogsJl2021} inspired the use of a custom number type to intercept operations on a number.
+The Julia library \texttt{Sherlogs}~\cite{kMilanklSherlogsJl2021} inspired our use of a custom number type to intercept operations on a number.
 In contrast to \FT{} which monitors for exceptional values and logs stack traces at interesting points in their lifetime, \texttt{Sherlogs} tracks and reports the range of values seen over the course of a computation.
@@
 FPSpy~\cite{dindaSpyingFloatingPoint2020} is an \texttt{LD\_PRELOAD} shared library that works on unmodified x86 binaries.
-It can monitor a program during execution for any \fp{} arithmetic issues, such as division by zero, denormalization, underflow, and overflow.
-\FT{} in contrast focuses on the lifetime of exceptional values; \FT{} can notice when a \NaN{} gets killed, for instance, while that is not an event that FPSpy appears to be able to track.
+It monitors a program during execution for operations that generate an exception, such as division by zero, underflow, and overflow.
+By contrast to \FT{}, it does not track prop or kill events.
 FPSpy has the advantage of being lightweight enough to run on production code for certain loads.
 
-\paragraph{Stack Graphs}
+\subsection{Stack Graphs}
 
-CSTG~\cite{humphreySystematicDebuggingMethods2014} is built on the work of and related to the STAT~\cite{arnoldStackTraceAnalysis2007} tool developed by Laurence Livermore National Laboratory.
-STAT is a tool for collecting, analyzing, and visualizing stack traces from thousands of concurrent processes, and is designed to highlight anomalies quickly.
-It produces visualizations similar to those of CSTG, thought CSTG offers more compact views.
+Our stack graphs utilize the CSTG library for coalesced stack trace graphs~\cite{humphreySystematicDebuggingMethods2014}.
+In turn, CSTGs build on the STAT tool from LLNL~\cite{arnoldStackTraceAnalysis2007}.
+STAT collects, analyzes, and visualizes stack traces from concurrent processes
+to highlight anomalies.
+It produces visualizations similar to those of CSTG, thought CSTG offers more compact views and supports diffs.
 
-\paragraph{GPU Exception Tracking}
+\subsection{GPU Exception Tracking}
 
@@
 FPChecker relies on LLVM-level instrumentation of GPU kernels, and so cannot run on the plethora of closed-source GPU kernels in usage today.
-
 BinFPE~\cite{llg-soap-2022} is another tool in the same space;
@@
 The latter deficiency is particularly worrisome, as we have seen, silent \NaN{} kills can invalidate results without the user noticing.
-
 \GPUFPX{}~\cite{llsflg-hpdc-2023} improves on the work of FPChecker and BinFPE by being more performant and catching a wider set of errors, including those that alter control flow.
@@
 Despite these improvements, \GPUFPX{} is limited by the closed-source nature of common GPU cores, and cannot report at the rich level of source detail that \FT{} can.
-
-Work is ongoing to use \GPUFPX{} and \FT{} in tandem.
-
-% Error analysis, typical answer.
-% Difficult.
-% CITE
 
@@
 
-FILL Consistent exception handling.
+Lightweight tools for error analysis that can quickly identify
+\fp{} problems and suggest repairs are an important topic.
+The number of scale of scientific application has grown tremendously
+over the years.
+For small teams that cannot afford a full-time
+analyst, tools like \FlowFPX{} fill a critical role.
 
-https://github.com/JuliaLang/julia/issues/48523
-
-Ashton: @Ben we simply propagate whatever we got from the operation. If
-Julia's behavior around an operator change, that should be preserved even
-when using FloatTracker
-
-Ben: @Ashton tell the JuliaCon readers not just me!
+\FlowFPX{} it itself an evolving toolkit.
+Below we discuss some topics for future work.
 
@@
 
-% TODO: citations!!!!
-% TODO: real numbers!!!
 \FT{} incurs significant overhead on the order of 100x slower than a non-instrumented run of the same program.
+It is a debugging tool, not a production tool.
 To put this number into context, Valgrind runs with a similar level of slowdown.
 
-In addition to the cost incurred by intercepting \fp{} operations, gathering stack traces is expensive---we measured a significant slowdown on the order of 10x relative to a non-logging setup of \FT{}.
+In addition to the cost incurred by intercepting \fp{} operations, gathering stack traces is expensive.
+We observe a 10x slowdown on shallow waters with logging disabled.
 We recommend that users of \FT{} make use of the \texttt{maxLogs} and \texttt{exclude\_stacktrace} configuration options to limit the number and kind of logs gathered, and thereby reduce the number of calls to \texttt{stacktrace()}.
-
-When injection is turned on we defer the stack checks as late as we can.
-Once the fuzzer runs out of \NaN{}s, it effectively turns off, so the rest of the fuzzing session proceeds much more quickly.
-That said, gathering stack traces still incurs a heavy overhead during logging when interesting events occur;
-we recommend limiting the number and kinds of events logged for optimal performance.
-
-\subsection{Tracking More than \NaN{}s and \Inf{}s}
-
-FloatTracker is capable of monitoring all sorts of events in a \fp{} value's lifetime---not just a value going to \NaN{} or \Inf{}.
-For example, \FT{} could be made to monitor extreme values: in some circumstances it would be useful to notice when a value started climbing above or falling below a certain threshold.
-
-We spent most of the time engineering and testing \NaN{} tracking and injection.
-Adding \Inf{} tracking took a few hours.
-Adding other values of interest presents no fundamental difficulties.
+Stack traces are essential to decide where to inject a fault when fuzzing,
+but we defer them as late as possible to maximize performance.
 
@@
 While fuzzing is useful for discovering issues, its success rate
-is low because \emph{every} \fp{} operation is a candidate
+is low because {every} \fp{} operation is a candidate
 for injection.
@@
 
-\FT{}'s scope is limited to Julia code: it cannot track the lifetime of exceptional values that are born, live, or die in external libraries, such as C programs.
-One approach would be to create an API for \FT{} so that programs using these external libraries could leverage \FT{}'s tooling to monitor the interface between Julia and e.g. external C programs.
+\FT{} is limited to Julia code.
+It cannot track the lifetime of exceptional values in external libraries, such as GPU kernels or C programs.
+For GPUs it relies on \GPUFPX{}, through the connection between these tools is loose.
+In the future, \FT{} would benefit from an API to plug in tools for external libraries, gather their output,
+and present a comprehensive view of exceptions in a multi-language program.
+
+\subsection{Interface Concerns}
+
+FloatTracker is capable of monitoring all sorts of events and \fp{} values beyond exceptions.
+For example, \FT{} could monitor large or small values with guidance from the user about which values are worth logging.
+On a similar note, the interface to \FT{} is primitive: users express interest in a number by wrapping it
+in a constructor such as \code{TrackedFloat64}.
+Helper functions that track data structures and choose a default size would
+make it easier to experiment with \FT{}.
 
@@
 
-We would like to thank Alex Larsen and Rob Durst for their help with reading drafts of the paper and feedback on the presentation.
-We would also like to thank the developers of \code{RxInfer.jl} for feedback on \FlowFPX{}.
-Finally, many thanks to Milan Klöwer, the author of the \texttt{Sherlogs} library, for inspiration on the architecture of \FT{}.
+Thanks to the \code{Sherlogs.jl} developers for inspiring the architecture of \FT{}.
+Thanks to the \code{RxInfer.jl} developers for testing \FT{} and for feedback on its user interface.
+Thanks to Alex Larsen and Rob Durst for comments on an early draft.
+This work is supported by
+NSF grant \href{https://nsf.gov/awardsearch/showAward?AWD_ID=2030859&HistoricalAwards=false}{2030859}
+to the CRA for the \href{https://cifellows2020.org}{CIFellows} project.
 

### 2023-08-19 paper.tex
png → pdf

@@
 \begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/genpropkill-outline.png}
+  \includegraphics[width=\columnwidth]{fig/genpropkill-outline.pdf}
   \caption{Gen, Prop, Kill: Lifetime of an exceptional value}
@@
   \centering
-  \includegraphics[width=0.9\columnwidth]{./fig/cstg_static_diagram.png}
+  \includegraphics[width=0.9\columnwidth]{./fig/cstg_static_diagram.pdf}
   \caption{From stack traces (left) to \CSTG{} (right)}
@@
   \centering
-  \includegraphics[width=3in]{./fig/shallow_waters_cfl_diff.png}
+  \includegraphics[width=3in]{./fig/shallow_waters_cfl_diff.pdf}
   \caption{Raising the CFL number creates white gaps due to \NaN{}s}
@@
   \centering
-  \includegraphics[width=0.96\columnwidth]{fig/sw_nan_cstg_clean.png}
+  \includegraphics[width=0.96\columnwidth]{fig/sw_nan_cstg_clean.pdf}
   \caption{Stack graph for \NaN{} gens in shallow waters}
@@
   \centering
-  \includegraphics[width=0.96\columnwidth]{./fig/cstg_diff_pretty.png}
+  \includegraphics[width=0.96\columnwidth]{./fig/cstg_diff_pretty.pdf}
   \caption{Stack graph diff}

### 2023-08-21 paper.tex
Gen prop kill outline is now all vector graphics
Add some edits from Ganesh

@@
 \Fp{} numbers use a finite number of bits to represent a spectrum of points along the real number line~(\cref{f:real-vs-fp}).
-The implementation strategy is essentially scientific notation.
+The implementation strategy is essentially that followed in scientific notation.
 A \fp{} number packs a sign bit, an exponent, and a fraction part (also called the ``significand'' or ``mantissa'') into a bitstring.
@@
 \begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/genpropkill-outline.pdf}
+  \includegraphics[width=\columnwidth]{fig/genpropkill-outline_vector.pdf}
   \caption{Gen, Prop, Kill: Lifetime of an exceptional value}
@@
 in shallow waters with the high CFL number.
-Reading bottom-up, every \NaN{} came from calls to the \code{-} and \code{+} operations.
-Calls to \code{+} account for most of the \NaN{}s.
+Reading bottom-up, every \NaN{} came from calls to the `\code{-}' and `\code{+}' operations.
+Calls to `\code{+}' account for most of the \NaN{}s.
 These \NaN{}s arose in two different contexts: a small number (30)
@@
 NSF grant \href{https://nsf.gov/awardsearch/showAward?AWD_ID=2030859&HistoricalAwards=false}{2030859}
-to the CRA for the \href{https://cifellows2020.org}{CIFellows} project.
+to the CRA for the \href{https://cifellows2020.org}{CIFellows} project,
+and DOE ASCR Award Number DE-SC0022252 and NSF CISE Awards 1956106 and 21241.
 

### 2023-08-22 paper.tex
Fix typo
fundamentally unreliable → error-prone

@@
   but the floating-point arithmetic that enables large-scale models
-  is fundamentally unreliable.
+  is error-prone.
   Numeric exceptions are a common occurrence and can propagate through
@@
 \fp{} exceptions~(\cref{s:flowfpx})
-that hase helped improve a variety of applications,
+that has helped improve a variety of applications,
 from ocean simulations to heat modes~(\cref{s:casestudies}):

### 2023-08-22 paper.tex
Fix library name formatting conventions, resolve feedback from Xinyi

@@
 Thus, the implementation of \FT{} uses a metaprogramming technique
-adapted from \texttt{Sherlogs.jl}~\cite{kMilanklSherlogsJl2021}
+adapted from \texttt{Sherlogs}~\cite{kMilanklSherlogsJl2021}
 to abstract over the common patterns.
@@
 
-\texttt{ShallowWaters.jl} simulates the flow of water over a
+\texttt{ShallowWaters} simulates the flow of water over a
 seabed~\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019}.
@@
 floats, e.g., \code{Float32} to \code{TrackedFloat32}.
-For shallow waters, this step is easy because the simulation is parameterized by
+For \texttt{ShallowWaters}, this step is easy because the simulation is parameterized by
 a \fp{} type for internal use.
@@
 
-While the logs for shallow waters contain useful information,
+While the logs for \texttt{ShallowWaters} contain useful information,
 there is an overwhelming amount of it.
@@
 \Cref{fig:sw_nan_cstg} presents the stack graph for \NaN{} gen events
-in shallow waters with the high CFL number.
+in \texttt{ShallowWaters} with the high CFL number.
 Reading bottom-up, every \NaN{} came from calls to the `\code{-}' and `\code{+}' operations.
@@
   \includegraphics[width=0.96\columnwidth]{fig/sw_nan_cstg_clean.pdf}
-  \caption{Stack graph for \NaN{} gens in shallow waters}
+  \caption{Stack graph for \NaN{} gens in \texttt{ShallowWaters}}
   \label{fig:sw_nan_cstg}
@@
 \Cref{fig:cstg_diff_demo} illustrates one diff in the context of \NaN{} gens for
-shallow waters.
-It compares the first 10\% of gen events to the latter 90\% of gens.
-The positive numbers and green lines indicate flows that are new in the latter part.
-The negative numbers and red lines show flows that disappeared in the latter part.
-
-Another one of CSTGs abilities is \emph{graph diffing}.
-We took the \FT{} logs from running ShallowWaters and split the logs at the 10\% mark.
-We used CSTG to diff the flow graphs between the first 10\% of the logs and the remaining 90\%.
-The diff can be seen in \cref{fig:cstg_diff_demo}.
+\texttt{ShallowWaters}:
+it compares the first 10\% of gen events to the latter 90\% of gens.
+The positive numbers and green lines indicate flows that are new in the latter part,
+and the negative numbers and red lines show flows that disappeared in the latter part.
 A domain expert might use these clues to find where an instability started in
@@
 In this case, the function \texttt{momentum\_u!} showed up only in the
-beginning of the logs; it may be an effective point to check for \Nan{}s.
+beginning of the logs---it may be an effective point to check for \Nan{}s.
 
@@
 team.\footnote{\url{https://github.com/SciML/NBodySimulator.jl}}
-Fuzzing discovered zero bugs and zero exceptional events.
-In fact, fuzzing injected zero exceptional values whatsoever, even with maximum odds.
-The reason for these null results, it turns out, is that the nbody
+We only wanted to test the robustness of the \texttt{NBodySimulator} package, so we configured the fuzzer to inject \NaN{}s \emph{only} when inside of a function of that library.
+This way, we could avoid injecting in the standard library or other dependencies of \texttt{NBodySimulator} and concentrate on improving this one library.
+
+However, Fuzzing discovered zero bugs and zero exceptional events.
+In fact, fuzzing injected \emph{zero} exceptional values whatsoever, even with adjusting the odds to \emph{force} an injection at the first available opportunity.
+The reason for this lack of \NaN{} injection, it turns out, is that the nbody
 simulation merely creates a problem for the \code{OrdinaryDiffEq}
@@
 
-Oceananigans~\cite{OceananigansJOSS} is simulation package for incompressible
+\texttt{Oceananigans}~\cite{OceananigansJOSS} is simulation package for incompressible
 fluid dynamics that can generate code for Nvidia GPUs.
@@
 In addition to the cost incurred by intercepting \fp{} operations, gathering stack traces is expensive.
-We observe a 10x slowdown on shallow waters with logging disabled.
+We observe a 10x slowdown on \texttt{ShallowWaters} with logging disabled.
 We recommend that users of \FT{} make use of the \texttt{maxLogs} and \texttt{exclude\_stacktrace} configuration options to limit the number and kind of logs gathered, and thereby reduce the number of calls to \texttt{stacktrace()}.

### 2024-05-31 paper.tex
Clear preamble
Bump
Whedon does not like pifont it seems
cite p3109 wg report
flowfpx landing site link

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{cleveref}
 \usepackage{microtype}
-\usepackage{pifont}
+% \usepackage{pifont}
 
@@
 \newcommand{\bigcheckmark}{\ding{51}}
-\newcommand{\bigxmark}{\ding{53}}
+% \newcommand{\bigxmark}{\ding{53}}
 \newcommand{\tblnext}{\(\Rightarrow\)}
-\newcommand{\tblY}{\mbox{\bigcheckmark}}
-\newcommand{\tblN}{\scalebox{0.8}{\bigxmark}}
+% \newcommand{\tblY}{\mbox{\bigcheckmark}}
+% \newcommand{\tblN}{\scalebox{0.8}{\bigxmark}}
+\newcommand{\tblY}{\checkmark}
+\newcommand{\tblN}{\scalebox{1.2}{$\times$}}
 
@@
 
+\noindent%
+The toolkit is online: \url{https://utahplt.github.io/flowfpx}
+
 \section{Floating-Point Exception Primer}
@@
 A \fp{} number packs a sign bit, an exponent, and a fraction part (also called the ``significand'' or ``mantissa'') into a bitstring.
-Typical strings are 64 or 32 bits long, but 16-bit and 8-bit formats are on the rise~\cite{klowerLowprecisionClimateComputing2021,fp8}.
+Typical strings are 64 or 32 bits long, but 16-bit and 8-bit formats are on the rise~\cite{klowerLowprecisionClimateComputing2021,fp8,p3109}.
 This representation supports very small and very large numbers in a narrow range of bits:
@@
 both lead to exceptions:
+
+\bigskip
 

### 2024-07-25 paper.tex
Start work on fixing paper for submission
Fix floatmin/floatmax issue

@@
 \begin{lstlisting}[language = Julia]
-julia> prevfloat(typemax(Float64))
+julia> floatmax(Float64)
 1.7976931348623157e308
-julia> nextfloat(typemin(Float64))
--1.7976931348623157e308
+julia> floatmin(Float64)
+2.2250738585072014e-308
 \end{lstlisting}
@@
 
-\FlowFPX{} is a toolkit for tracking down \fp{} exceptions.
+\FlowFPX{} (FPX: Floating Point eXeption) is a toolkit for tracking down \fp{} exceptions.
 The primary tools in this paper are \FT{}, which records lifetimes
@@
 % jinx-local-words: "BinFPE CFL CSTG's CSTGs Courant Demmel FPChecker Friedrichs Herbie JuliaHub Klöwer Lewy OrdinaryDiffEq RxInfer ShallowWaters Sherlogs TrackedFloat isnan"
-% citar-bibliography: ("./ref.bib")
+% citar-bibliography: nil
 % End:

### 2024-07-25 paper.tex
Typos; attempt to explain injection better; clarify what FlowFPX is

@@
 that has helped improve a variety of applications,
-from ocean simulations to heat modes~(\cref{s:casestudies}):
-\begin{itemize}
+from ocean simulations to heat modes~(\cref{s:casestudies}).
+\FlowFPX{} consists of three tools that can work together to debug \fp{} computations:
+\begin{enumerate}
   \item
     The centerpiece of \FlowFPX{} is \FT{}, a dynamic analysis tool that
-    selectively monitors for exceptions and fuzzes code for vulnerabilities.
+    selectively monitors for exceptions and fuzzes Julia code for vulnerabilities.
+    \FT{} is a Julia library that operates on Julia code.
   \item
-    To visualize results, \FT{} adapts coalesced stack-trace graphs (CSTGs, or
+    To visualize results, \FlowFPX{} uses coalesced stack-trace graphs (CSTGs, or
     \CSTG{}s)~\cite{humphreySystematicDebuggingMethods2014}
     to summarize the program contexts that handled exceptional values.
+    \CSTG{} is a generic stack-trace graph summarization tool.
   \item
     A companion tool \GPUFPX{}~\cite{llsflg-hpdc-2023} provides fine-grained insights for GPU kernels.
-\end{itemize}
+    \GPUFPX{} is a dynamic library that programmers of GPU kernels can use.
+\end{enumerate}
 
@@
 The implementation strategy is essentially that followed in scientific notation.
-A \fp{} number packs a sign bit, an exponent, and a fraction part (also called the ``significand'' or ``mantissa'') into a bitstring.
+A \fp{} number packs a sign bit, an exponent, and a fraction part (also called the ``significand'' or ``mantissa'') into a bit string.
 Typical strings are 64 or 32 bits long, but 16-bit and 8-bit formats are on the rise~\cite{klowerLowprecisionClimateComputing2021,fp8,p3109}.
@@
 \FlowFPX{} (FPX: Floating Point eXeption) is a toolkit for tracking down \fp{} exceptions.
-The primary tools in this paper are \FT{}, which records lifetimes
+The primary tools in the \FlowFPX{} toolkit are \FT{}, which records lifetimes
 and enables fuzzing, and \CSTG{}s (more precisely, CSTGs), which visualize
@@
 \GPUFPX{} is a third component of \FlowFPX{} that tracks \fp{} exceptions inside GPU kernels~\cite{llsflg-hpdc-2023}.
+This section covers each tool in detail.
 
@@
 
-\FT{} tracks exceptional values across their \genpropkill{} lifetime
-and can fuzz code for vulnerabilities by injecting a \Nan{} or \Inf{}
-as the result of an operation.
-\FT{} is implemented in Julia and is available on JuliaHub:
+\FT{} tracks exceptional values in Julia programs across
+their \genpropkill{} lifetime and can fuzz code for vulnerabilities by
+injecting a \Nan{} or \Inf{} as the result of an operation. \FT{} is
+implemented in Julia as a library and is available on JuliaHub:
 
@@
 
+We note that \FT{} is currently not thread-safe; event logs are
+collected globally and running \FT{} in a multithreaded environment
+carries risks of corrupting the event log stack.
+Improving \FT{} to be thread-safe is an area for future work.
+
 \subsubsection{Fuzzing}
 
-Operator overloading gives \FT{} a powerful way to fuzz code from the inside out.
+In addition to enabling tracking exceptional values, operator
+overloading gives \FT{} a powerful way to fuzz code from the inside
+out.
 Each overloaded function serves as a hook where \FT{} can decide whether to observe
@@
 
-\FT{} exposes several parameters to let developers control the fuzzer:
+When fuzzing is enabled, every time \FT{} intercepts a \fp{}
+operation, it decides whether or not it should return a \Nan{} instead
+of the actual computed value.
+This decision is based on several parameters that the developer can
+tune to control fuzzing:
 
 \begin{itemize}
+\item \texttt{active::Bool} fuzz only when set to true.
 \item \texttt{odds::Int64} inject if \texttt{rand(odds) == 1}.
 \item \texttt{n\_inject::Int64} upper bound on the number of \Nan{}s to inject.
-\item \texttt{active::Bool} fuzz only when set to true.
 \item \texttt{functions::Array{String}} limit fuzzing to the dynamic extent of the listed functions.
@@
 \end{itemize}
+
+\FT{} looks at these parameters in conjunction with stack-trace at the
+point where the operation was intercepted to make the decision.
+When all the conditions are met, the overloaded operator returns
+\Nan{} instead of whatever value would have been computed normally.
+This means that \Nan{}s can be injected deep inside of an algorithm,
+instead of just at the inputs.
+This can reveal vulnerabilities that otherwise would have been
+difficult to catch.
+
+Developers can use the \texttt{functions} and \texttt{libraries}
+parameters to narrow down the region where \Nan{}s can be injected.
+It might not be useful, for instance, to fuzz test code that a
+developer has no control over; a developer may not, for example, wish
+to fuzz the standard library and instead focus on the code that they
+are developing.
 
@@
 their code and check that the fixes remove the error.
+When running a replay, each intercepted operation consults the reply
+file (instead of the stack-trace and fuzzing parameters) to decide
+whether or not to return a \Nan{} from the computation.
 
@@
 Coalesced stack trace graphs (CSTGs or \CSTG{}s for short) provide a way to visualize large amounts of stack traces in a compact form~\cite{humphreySystematicDebuggingMethods2014}.
-This techinque pairs well with \FT{} and makes analyzing the log files easier.
+This technique pairs well with \FT{} and makes analyzing the log files easier.
 
@@
 fluid dynamics that can generate code for Nvidia GPUs.
-For example, the following program (from the project readme) simulates turbulance:
+For example, the following program (from the project readme) simulates turbulence:
 
@@
 % Arnab Das scalable yet rigorous
-% Marc Daumas Certification of Bounds Expres
+% Marc Daumas Certification of Bounds Expression
 % David Delmas Towards an Industrial Use of

### 2024-07-25 paper.tex
Better description of how injection gets decided

@@
   \item
-    To visualize results, \FlowFPX{} uses coalesced stack-trace graphs (CSTGs, or
+    To visualize results, \FlowFPX{} uses coalesced stack trace graphs (CSTGs, or
     \CSTG{}s)~\cite{humphreySystematicDebuggingMethods2014}
     to summarize the program contexts that handled exceptional values.
-    \CSTG{} is a generic stack-trace graph summarization tool.
+    \CSTG{} is a generic stack trace graph summarization tool.
   \item
@@
 
-\FT{} looks at these parameters in conjunction with stack-trace at the
+\FT{} looks at these parameters in conjunction with stack trace at the
 point where the operation was intercepted to make the decision.
+When fuzzing is active (\texttt{active == true}), there are \Nan{}s
+remaining to be injected (\texttt{n\_inject > 0}), and
+\texttt{rand(odds)} returns \texttt{1}, then \FT{} inspects the
+current stack trace to see if the point is in the dynamic extend of
+the wanted \texttt{functions} or \texttt{libraries}.
 When all the conditions are met, the overloaded operator returns
 \Nan{} instead of whatever value would have been computed normally.
-This means that \Nan{}s can be injected deep inside of an algorithm,
-instead of just at the inputs.
+The value \texttt{n\_inject} is decremented before the next operation
+gets intercepted.
+
+Since \FT{} checks if it should inject a \Nan{} at \emph{every} \fp{}
+operation it intercepts, means that \Nan{}s can be injected deep
+inside of an algorithm instead of just at the inputs.
 This can reveal vulnerabilities that otherwise would have been
-difficult to catch.
+difficult to catch with input-based fuzzers.
 
@@
 When running a replay, each intercepted operation consults the reply
-file (instead of the stack-trace and fuzzing parameters) to decide
+file (instead of the stack trace and fuzzing parameters) to decide
 whether or not to return a \Nan{} from the computation.

### 2024-07-26 paper.tex
Clarify relation between GPU-FPX and FloatTracker (no formal one)

@@
 Together, \FT{} and \GPUFPX{} provide insights for accelerated programs.
+
+\FT{} and \GPUFPX{} have no formal connection---that is, one does not
+need to use \GPUFPX{} with \FT{}, nor vice-versa.
+These two tools compliment one another in debugging mixed-paradigm
+workloads, but they can be used in isolation.
+\GPUFPX{} is a library exclusively for NVIDIA CUDA code and has no
+special relation to Julia.
+Nevertheless, in cases where Julia code leverages a CUDA kernel,
+\GPUFPX and \FT{} can both aid in tracking down exceptional \fp{}
+values.
 

### 2024-08-01 paper.tex
jcon2: all pkg names --> +.jl

@@
 \newcommand{\GPUFPX}{GPU-FPX}
-\newcommand{\FloatTracker}{FloatTracker}
-\newcommand{\FT}{\FloatTracker}
-\newcommand{\Fp}{Floating-point} % hyphen or no?
-\newcommand{\fp}{floating-point} % hyphen or no?
+\newcommand{\juliapkg}[1]{\texttt{#1.jl}}
+\newcommand{\TrackedFloats}{\juliapkg{TrackedFloats}}
+\newcommand{\TF}{\TrackedFloats}
+\newcommand{\ShallowWaters}{\juliapkg{ShallowWaters}}
+\newcommand{\OrdinaryDiffEq}{\juliapkg{OrdinaryDiffEq}}
+\newcommand{\RxInfer}{\juliapkg{RxInfer}}
+\newcommand{\Sherlogs}{\juliapkg{Sherlogs}}
+\newcommand{\NBodySimulator}{\juliapkg{NBodySimulator}}
+\newcommand{\Fp}{Floating-point}
+\newcommand{\fp}{floating-point}
 \newcommand{\CSTG}{stack graph}
@@
   \item
-    The centerpiece of \FlowFPX{} is \FT{}, a dynamic analysis tool that
+    The centerpiece of \FlowFPX{} is \TF{}, a dynamic analysis tool that
     selectively monitors for exceptions and fuzzes Julia code for vulnerabilities.
-    \FT{} is a Julia library that operates on Julia code.
+    \TF{} is a Julia library that operates on Julia code.
   \item
@@
 \FlowFPX{} (FPX: Floating Point eXeption) is a toolkit for tracking down \fp{} exceptions.
-The primary tools in the \FlowFPX{} toolkit are \FT{}, which records lifetimes
+The primary tools in the \FlowFPX{} toolkit are \TF{}, which records lifetimes
 and enables fuzzing, and \CSTG{}s (more precisely, CSTGs), which visualize
@@
 
-\subsection{\FT{}}
+\subsection{\TF{}}
 \label{s:floattracker}
 
-\FT{} tracks exceptional values in Julia programs across
+\TF{} tracks exceptional values in Julia programs across
 their \genpropkill{} lifetime and can fuzz code for vulnerabilities by
-injecting a \Nan{} or \Inf{} as the result of an operation. \FT{} is
+injecting a \Nan{} or \Inf{} as the result of an operation. \TF{} is
 implemented in Julia as a library and is available on JuliaHub:
@@
 
-\FT{} monitors \Nan{} and \Inf{} exceptions by overloading arithmetic
+\TF{} monitors \Nan{} and \Inf{} exceptions by overloading arithmetic
 operations and logging key events.
@@
 
-\FT{} writes logs to three files: one for {gen} events, one for {prop}s, and one for {kill}s.
+\TF{} writes logs to three files: one for {gen} events, one for {prop}s, and one for {kill}s.
 This way, discovering where a \Nan{} came from is a matter of sifting through the {gen} file.
@@
 \code{TrackedFloat32}, and \code{TrackedFloat16}.
-Developers must opt in to \FT{} by wrapping numbers in a custom type.
+Developers must opt in to \TF{} by wrapping numbers in a custom type.
 From then on, tracking is automatic and extends transitively to all outputs
@@
 bit-for-bit,\footnote{\url{https://github.com/JuliaLang/julia/issues/48523}}
-but \FT{} does its part.
+but \TF{} does its part.
 
-We note that \FT{} is currently not thread-safe; event logs are
-collected globally and running \FT{} in a multithreaded environment
+We note that \TF{} is currently not thread-safe; event logs are
+collected globally and running \TF{} in a multithreaded environment
 carries risks of corrupting the event log stack.
-Improving \FT{} to be thread-safe is an area for future work.
+Improving \TF{} to be thread-safe is an area for future work.
 
@@
 In addition to enabling tracking exceptional values, operator
-overloading gives \FT{} a powerful way to fuzz code from the inside
+overloading gives \TF{} a powerful way to fuzz code from the inside
 out.
-Each overloaded function serves as a hook where \FT{} can decide whether to observe
+Each overloaded function serves as a hook where \TF{} can decide whether to observe
 the operation or step in, discard the original result, and return an exception instead.
@@
 
-When fuzzing is enabled, every time \FT{} intercepts a \fp{}
+When fuzzing is enabled, every time \TF{} intercepts a \fp{}
 operation, it decides whether or not it should return a \Nan{} instead
@@
 
-\FT{} looks at these parameters in conjunction with stack trace at the
+\TF{} looks at these parameters in conjunction with stack trace at the
 point where the operation was intercepted to make the decision.
@@
 remaining to be injected (\texttt{n\_inject > 0}), and
-\texttt{rand(odds)} returns \texttt{1}, then \FT{} inspects the
+\texttt{rand(odds)} returns \texttt{1}, then \TF{} inspects the
 current stack trace to see if the point is in the dynamic extend of
@@
 
-Since \FT{} checks if it should inject a \Nan{} at \emph{every} \fp{}
+Since \TF{} checks if it should inject a \Nan{} at \emph{every} \fp{}
 operation it intercepts, means that \Nan{}s can be injected deep
@@
 guide repairs.
-\FT{} therefore records the sequence of injections that it makes during a fuzzing
+\TF{} therefore records the sequence of injections that it makes during a fuzzing
 run and enables a replay of any recording after the fact.
@@
 
-\subsubsection{\FT{} Internals}
+\subsubsection{\TF{} Internals}
 
-\FT{} takes advantage of Julia's operator overloading to track exceptions and
+\TF{} takes advantage of Julia's operator overloading to track exceptions and
 fuzz for vulnerabilities.
@@
 before returning:
-This lets \FT{} intercept all \fp{} operations involving \texttt{TrackedFloat} types.
+This lets \TF{} intercept all \fp{} operations involving \texttt{TrackedFloat} types.
 
@@
 Every arithmetic operation requires a similar overloading.
-Thus, the implementation of \FT{} uses a metaprogramming technique
+Thus, the implementation of \TF{} uses a metaprogramming technique
 adapted from \texttt{Sherlogs}~\cite{kMilanklSherlogsJl2021}
@@
 
-\FT{} can produce copious amounts of log files which can be challenging to sift through manually.
+\TF{} can produce copious amounts of log files which can be challenging to sift through manually.
 Coalesced stack trace graphs (CSTGs or \CSTG{}s for short) provide a way to visualize large amounts of stack traces in a compact form~\cite{humphreySystematicDebuggingMethods2014}.
-This technique pairs well with \FT{} and makes analyzing the log files easier.
+This technique pairs well with \TF{} and makes analyzing the log files easier.
 
@@
 In fact, GPU programs are worse off because they lack exception-handling mechanisms from the CPU world~\cite{llg-soap-2022}.
-Since \FT{} instruments Julia programs, it cannot help directly;
+Since \TF{} instruments Julia programs, it cannot help directly;
 however, the companion tool \GPUFPX{} instruments GPU kernels to detect and report
 \fp{} exceptions~\cite{llsflg-hpdc-2023}.
-Together, \FT{} and \GPUFPX{} provide insights for accelerated programs.
+Together, \TF{} and \GPUFPX{} provide insights for accelerated programs.
 
-\FT{} and \GPUFPX{} have no formal connection---that is, one does not
-need to use \GPUFPX{} with \FT{}, nor vice-versa.
+\TF{} and \GPUFPX{} have no formal connection---that is, one does not
+need to use \GPUFPX{} with \TF{}, nor vice-versa.
 These two tools compliment one another in debugging mixed-paradigm
@@
 Nevertheless, in cases where Julia code leverages a CUDA kernel,
-\GPUFPX and \FT{} can both aid in tracking down exceptional \fp{}
+\GPUFPX and \TF{} can both aid in tracking down exceptional \fp{}
 values.
@@
 The case studies include a shallow water simulation, the
-\texttt{OrdinaryDiffEq} solver, and a Bayesian inference library.
+\OrdinaryDiffEq{} solver, and a Bayesian inference library.
 
-\subsection{ShallowWaters}
+\subsection{\ShallowWaters{}}
 \label{s:sw}
 
-\texttt{ShallowWaters} simulates the flow of water over a
+\ShallowWaters{} simulates the flow of water over a
 seabed~\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019}.
@@
 
-Running \FT{} on the high-CFL simulation reveals where the simulation drifted
+Running \TF{} on the high-CFL simulation reveals where the simulation drifted
 from numbers to \NaN{} exceptions.
-The first step in applying \FT{} is to convert relevant floats to tracked
+The first step in applying \TF{} is to convert relevant floats to tracked
 floats, e.g., \code{Float32} to \code{TrackedFloat32}.
-For \texttt{ShallowWaters}, this step is easy because the simulation is parameterized by
+For \ShallowWaters{}, this step is easy because the simulation is parameterized by
 a \fp{} type for internal use.
@@
 \begin{lstlisting}
--([-Inf, -Inf])   FT/TrackedFloat.jl:106
+-([-Inf, -Inf])   TF/TrackedFloat.jl:106
 momentum_u!       SW/rhs.jl:246
@@
 \begin{lstlisting}
-^([-1.515f31, 2])          FT/TrackedFloat.jl:138
+^([-1.515f31, 2])          TF/TrackedFloat.jl:138
 literal_pow()              intfuncs.jl:325
@@
 This \Inf{} came from an exponent that overflowed the float type (\code{-1.515e31\string^2}).
-\FT{} has shown exactly which numeric values in which operation
+\TF{} has shown exactly which numeric values in which operation
 caused exceptions to occur.
@@
 
-While the logs for \texttt{ShallowWaters} contain useful information,
+While the logs for \ShallowWaters{} contain useful information,
 there is an overwhelming amount of it.
@@
 \Cref{fig:sw_nan_cstg} presents the stack graph for \NaN{} gen events
-in \texttt{ShallowWaters} with the high CFL number.
+in \ShallowWaters{} with the high CFL number.
 Reading bottom-up, every \NaN{} came from calls to the `\code{-}' and `\code{+}' operations.
@@
   \includegraphics[width=0.96\columnwidth]{fig/sw_nan_cstg_clean.pdf}
-  \caption{Stack graph for \NaN{} gens in \texttt{ShallowWaters}}
+  \caption{Stack graph for \NaN{} gens in \ShallowWaters{}}
   \label{fig:sw_nan_cstg}
@@
 \Cref{fig:cstg_diff_demo} illustrates one diff in the context of \NaN{} gens for
-\texttt{ShallowWaters}:
+\ShallowWaters{}:
 it compares the first 10\% of gen events to the latter 90\% of gens.
@@
 
-\subsection{OrdinaryDiffEq}
+\subsection{\OrdinaryDiffEq{}}
 \label{s:ode}
 
-With the targeted fuzzing abilities of \FT{}, we set focus on the
-\texttt{NBodySimulator} package from the SciML
+TODO awk.
+With the targeted fuzzing abilities of \TF{}, we set focus on the
+\NBodySimulator{} package from the SciML
 team.\footnote{\url{https://github.com/SciML/NBodySimulator.jl}}
-We only wanted to test the robustness of the \texttt{NBodySimulator} package, so we configured the fuzzer to inject \NaN{}s \emph{only} when inside of a function of that library.
-This way, we could avoid injecting in the standard library or other dependencies of \texttt{NBodySimulator} and concentrate on improving this one library.
+We only wanted to test the robustness of the \NBodySimulator{} package, so we configured the fuzzer to inject \NaN{}s \emph{only} when inside of a function of that library.
+This way, we could avoid injecting in the standard library or other dependencies of \NBodySimulator{} and concentrate on improving this one library.
 
@@
 The reason for this lack of \NaN{} injection, it turns out, is that the nbody
-simulation merely creates a problem for the \code{OrdinaryDiffEq}
+simulation merely creates a problem for the \OrdinaryDiffEq{}
 solver.\footnote{\url{https://github.com/SciML/OrdinaryDiffEq.jl}}
@@
 
-Fuzzing on \code{OrdinaryDiffEq} led to a curious situation.
+Fuzzing on \OrdinaryDiffEq{} led to a curious situation.
 The library itself reported a \NaN{} and printed a message that it would exit.
@@
 \begin{lstlisting}
-<([NaN, 3.0e6])  at FT/TrackedFloat.jl:193
+<([NaN, 3.0e6])  at TF/TrackedFloat.jl:193
 solve!           at ODE/solve.jl:515
@@
 
-Fuzzing with \FT{} revealed two places where Finch needed
+Fuzzing with \TF{} revealed two places where Finch needed
 protection against user input.
@@
 Here, a \NaN{} could leave bounds uninitialized, leading to a bounds error.
-Additionally, \FT{} and \CSTG{}s have been useful for identifying \NaN{}s that
+Additionally, \TF{} and \CSTG{}s have been useful for identifying \NaN{}s that
 appear in unstable heat simulations written in Finch.
@@
 
-\subsection{RxInfer}
+\subsection{\RxInfer{}}
 \label{s:safari}
 
-We discovered an open issue in the \code{RxInfer.jl} library related to \NaN{} detection
-and suggested \FT{} to the developers.\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}}
+We discovered an open issue in the \RxInfer{} library related to \NaN{} detection
+and suggested \TF{} to the developers.\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}}
 Within a day, they were able to track down the location of an important \NaN{} gen.
@@
 closed-source.
-We merely explained how to use \FT{} by wrapping inputs in a tracked type.
+We merely explained how to use \TF{} by wrapping inputs in a tracked type.
 
@@
 The proposal includes an extension to the \texttt{INFO\_ARRAY} parameter with fields that record \genpropkill{} information.
-It also includes fuzzing, which we realize in \FT{}.
+It also includes fuzzing, which we realize in \TF{}.
 
@@
 Herbie~\cite{panchekhaAutomaticallyImprovingAccuracy2015} automatically rewrites arithmetic expressions to reduce \fp{} error.
-This would combine well with \FT{}: first identify the location of a \NaN{}, ask Herbie to find a repair.
+This would combine well with \TF{}: first identify the location of a \NaN{}, ask Herbie to find a repair.
 The Odyssey\cite{misbackOdysseyInteractiveWorkbench2023} workbench provides an interactive interface to Herbie.
@@
 
-The Julia library \texttt{Sherlogs}~\cite{kMilanklSherlogsJl2021} inspired our use of a custom number type to intercept operations on a number.
-In contrast to \FT{} which monitors for exceptional values and logs stack traces at interesting points in their lifetime, \texttt{Sherlogs} tracks and reports the range of values seen over the course of a computation.
+The Julia library \Sherlogs{}~\cite{kMilanklSherlogsJl2021} inspired our use of a custom number type to intercept operations on a number.
+In contrast to \TF{} which monitors for exceptional values and logs stack traces at interesting points in their lifetime, \Sherlogs{} tracks and reports the range of values seen over the course of a computation.
 This is intended to provide insight into whether or not a library could tolerate a lower-precision \fp{} format.
@@
 It monitors a program during execution for operations that generate an exception, such as division by zero, underflow, and overflow.
-By contrast to \FT{}, it does not track prop or kill events.
+By contrast to \TF{}, it does not track prop or kill events.
 FPSpy has the advantage of being lightweight enough to run on production code for certain loads.
@@
 \GPUFPX{} runs on CUDA cores from NVIDIA and reports total numbers of exceptional values.
-Like \FT{}, \GPUFPX{} can catch \NaN{} gens and kills, but it does this on the GPU where \FT{} doesn't apply.
-Despite these improvements, \GPUFPX{} is limited by the closed-source nature of common GPU cores, and cannot report at the rich level of source detail that \FT{} can.
+Like \TF{}, \GPUFPX{} can catch \NaN{} gens and kills, but it does this on the GPU where \TF{} doesn't apply.
+Despite these improvements, \GPUFPX{} is limited by the closed-source nature of common GPU cores, and cannot report at the rich level of source detail that \TF{} can.
 
@@
 
-\FT{} incurs significant overhead on the order of 100x slower than a non-instrumented run of the same program.
+\TF{} incurs significant overhead on the order of 100x slower than a non-instrumented run of the same program.
 It is a debugging tool, not a production tool.
@@
 In addition to the cost incurred by intercepting \fp{} operations, gathering stack traces is expensive.
-We observe a 10x slowdown on \texttt{ShallowWaters} with logging disabled.
-We recommend that users of \FT{} make use of the \texttt{maxLogs} and \texttt{exclude\_stacktrace} configuration options to limit the number and kind of logs gathered, and thereby reduce the number of calls to \texttt{stacktrace()}.
+We observe a 10x slowdown on \ShallowWaters{} with logging disabled.
+We recommend that users of \TF{} make use of the \texttt{maxLogs} and \texttt{exclude\_stacktrace} configuration options to limit the number and kind of logs gathered, and thereby reduce the number of calls to \texttt{stacktrace()}.
 Stack traces are essential to decide where to inject a fault when fuzzing,
@@
 Even operations that are already well-defended against \NaN{}s are candidates.
-\FT{} could use two sorts of tools for improving injection.
+\TF{} could use two sorts of tools for improving injection.
 First, fine-grained control to let users decide where not to inject.
@@
 
-\FT{} is limited to Julia code.
+\TF{} is limited to Julia code.
 It cannot track the lifetime of exceptional values in external libraries, such as GPU kernels or C programs.
 For GPUs it relies on \GPUFPX{}, through the connection between these tools is loose.
-In the future, \FT{} would benefit from an API to plug in tools for external libraries, gather their output,
+In the future, \TF{} would benefit from an API to plug in tools for external libraries, gather their output,
 and present a comprehensive view of exceptions in a multi-language program.
@@
 
-FloatTracker is capable of monitoring all sorts of events and \fp{} values beyond exceptions.
-For example, \FT{} could monitor large or small values with guidance from the user about which values are worth logging.
-On a similar note, the interface to \FT{} is primitive: users express interest in a number by wrapping it
+\TF{} is capable of monitoring all sorts of events and \fp{} values beyond exceptions.
+For example, \TF{} could monitor large or small values with guidance from the user about which values are worth logging.
+On a similar note, the interface to \TF{} is primitive: users express interest in a number by wrapping it
 in a constructor such as \code{TrackedFloat64}.
 Helper functions that track data structures and choose a default size would
-make it easier to experiment with \FT{}.
+make it easier to experiment with \TF{}.
 
@@
 
-Thanks to the \code{Sherlogs.jl} developers for inspiring the architecture of \FT{}.
-Thanks to the \code{RxInfer.jl} developers for testing \FT{} and for feedback on its user interface.
+Thanks to the \Sherlogs{} developers for inspiring the architecture of \TF{}.
+Thanks to the \RxInfer{} developers for testing \TF{} and for feedback on its user interface.
 Thanks to Alex Larsen and Rob Durst for comments on an early draft.

### 2024-08-01 paper.tex
jcon2: move future-work threads issue to the future work section

@@
 \newcommand{\NBodySimulator}{\juliapkg{NBodySimulator}}
+\newcommand{\Oceananigans}{\juliapkg{Oceananigans}}
 \newcommand{\Fp}{Floating-point}
@@
 but \TF{} does its part.
-
-We note that \TF{} is currently not thread-safe; event logs are
-collected globally and running \TF{} in a multithreaded environment
-carries risks of corrupting the event log stack.
-Improving \TF{} to be thread-safe is an area for future work.
 
@@
 
-\subsection{\OrdinaryDiffEq{}}
+\subsection{Fuzzing from NBody to ODE}
 \label{s:ode}
 
-TODO awk.
-With the targeted fuzzing abilities of \TF{}, we set focus on the
-\NBodySimulator{} package from the SciML
-team.\footnote{\url{https://github.com/SciML/NBodySimulator.jl}}
-We only wanted to test the robustness of the \NBodySimulator{} package, so we configured the fuzzer to inject \NaN{}s \emph{only} when inside of a function of that library.
-This way, we could avoid injecting in the standard library or other dependencies of \NBodySimulator{} and concentrate on improving this one library.
-
-However, Fuzzing discovered zero bugs and zero exceptional events.
-In fact, fuzzing injected \emph{zero} exceptional values whatsoever, even with adjusting the odds to \emph{force} an injection at the first available opportunity.
-The reason for this lack of \NaN{} injection, it turns out, is that the nbody
-simulation merely creates a problem for the \OrdinaryDiffEq{}
+We applied the fuzzing abilities of \TF{} to the
+\NBodySimulator{} package.\footnote{\url{https://github.com/SciML/NBodySimulator.jl}}
+To avoid injecting in the standard library or other dependencies,
+we initially configured the fuzzer to inject \NaN{}s {only} when inside of a
+function from the simulator.
+Surprisingly, fuzzing injected \emph{zero} exceptional values even
+when the fuzzer's odds were configured to force injection at the
+first available opportunity.
+It turns out that all the \fp{} operations for the simulation happened
+within the \OrdinaryDiffEq{}
 solver.\footnote{\url{https://github.com/SciML/OrdinaryDiffEq.jl}}
-All the \fp{} operations were in another library.
 
 Fuzzing on \OrdinaryDiffEq{} led to a curious situation.
-The library itself reported a \NaN{} and printed a message that it would exit.
+The library itself reported a \NaN{} and printed a message stating that it would exit.
 However, after printing that message, the program went into an infinite loop.
@@
 PDEs~\cite{heislerFinchDomainSpecific2022}.\footnote{Not to be confused
-with the Finch loop optimizer~\cite{adka-cgo-2023}.}
+with the loop optimizer \juliapkg{Finch}~\cite{adka-cgo-2023}.}
 In the spirit of FEniCS~\cite{fenics} and related
@@
 
-\subsection{Oceananigans}
+\subsection{\Oceananigans{}}
 \label{s:ocean}
 
-\texttt{Oceananigans}~\cite{OceananigansJOSS} is simulation package for incompressible
+\Oceananigans{}~\cite{OceananigansJOSS} is simulation package for incompressible
 fluid dynamics that can generate code for Nvidia GPUs.
@@
 
-\subsection{Interface Concerns}
+\subsection{Interface Concerns, Multi-Threading}
 
-\TF{} is capable of monitoring all sorts of events and \fp{} values beyond exceptions.
-For example, \TF{} could monitor large or small values with guidance from the user about which values are worth logging.
+\TF{} could be extended to monitor events and values that go beyond exceptions,
+such as very-large or very-small numbers according to bounds supplied by the user.
 On a similar note, the interface to \TF{} is primitive: users express interest in a number by wrapping it
 in a constructor such as \code{TrackedFloat64}.
-Helper functions that track data structures and choose a default size would
-make it easier to experiment with \TF{}.
+Helper functions that track data structures and provide a default float size would
+make experimentation easier.
+
+\TF{} stores its configuration and event logs in global data structures,
+making it unsafe for a multithreaded environment.
+Adding locks would restore safety, but perhaps a richer interface is in order.
+Individual threads may wish to configure local logging for better
+data organization and performance.
 

### 2024-08-01 paper.tex
jcon2: editos

@@
 catastrophic cancellation, singular matrices, and vanishing
-derivatives~\cite{sdjmrstp-pc-2022,ddghlllprr-correctness-2022,gllprt-correctness-2021,fpchecker-reports,llg-soap-2022,bllmg-xloop-2022}.
+derivatives~\cite{bllmg-xloop-2022,ddghlllprr-correctness-2022,fpchecker-reports,gllprt-correctness-2021,llg-soap-2022,sdjmrstp-pc-2022}.
 Developers are responsible for guarding against exceptions, but this task
@@
   \item
-    The centerpiece of \FlowFPX{} is \TF{}, a dynamic analysis tool that
-    selectively monitors for exceptions and fuzzes Julia code for vulnerabilities.
-    \TF{} is a Julia library that operates on Julia code.
+    The centerpiece of \FlowFPX{} is \TF{}, a dynamic analysis tool written in Julia
+    for Julia code that monitors exceptions and fuzzes for vulnerabilities.
   \item
     To visualize results, \FlowFPX{} uses coalesced stack trace graphs (CSTGs, or
-    \CSTG{}s)~\cite{humphreySystematicDebuggingMethods2014}
-    to summarize the program contexts that handled exceptional values.
-    \CSTG{} is a generic stack trace graph summarization tool.
+    \CSTG{}s)~\cite{humphreySystematicDebuggingMethods2014}.
+    Our CSTG library is written in \CPP{} and accepts text-format stack traces as input.
   \item
-    A companion tool \GPUFPX{}~\cite{llsflg-hpdc-2023} provides fine-grained insights for GPU kernels.
-    \GPUFPX{} is a dynamic library that programmers of GPU kernels can use.
+    For programs that offload work to CUDA GPUs,
+    the binary instrumentation tool \GPUFPX{}~\cite{llsflg-hpdc-2023}
+    adds logging to kernels.
 \end{enumerate}
@@
 and can lead to surprising results.
-For example, adding the tiny Planck constant to the large Avogadro number results
-in Avogadro's number after rounding:
+For example, adding the tiny Planck constant to the large Avogadro number results in
+the Avogadro number after rounding:
 
@@
 and implement an appropriate repair.
-This is no easy task, however, because of the approximations inherent to
+This is a difficult task because of the approximations inherent to
 floating point.
@@
 denominator gets truncated to zero.
-Some \Nan{}s might be spurious, others might be fatal, but in any event
+Some \Nan{}s might be spurious while others might be fatal, but in any event
 anticipating the various exceptions is a burden.
@@
 
-\FlowFPX{} (FPX: Floating Point eXeption) is a toolkit for tracking down \fp{} exceptions.
+\FlowFPX{} (FPX: Floating Point eXception) is a toolkit for tracking down \fp{} exceptions.
 The primary tools in the \FlowFPX{} toolkit are \TF{}, which records lifetimes
@@
 injecting a \Nan{} or \Inf{} as the result of an operation. \TF{} is
-implemented in Julia as a library and is available on JuliaHub:
+implemented in Julia as a library and is available on JuliaHub (TODO update link):
 
@@
 
-Crucially, the tracked version of an exception, say a \NaN{}, points
-to the original exceptional value.
-This enables techniques such as \NaN{}-packing, or otherwise
-using the payload in creative ways.
-Not all Julia operations currently preserve \NaN{}s
-bit-for-bit,\footnote{\url{https://github.com/JuliaLang/julia/issues/48523}}
-but \TF{} does its part.
+Crucially,
+\TF{} does nothing to modify the bit-level representation of floats or exceptions.
+The tracked version of, say, a \NaN{}, points to the original exceptional
+value.
+Thus \NaN{}-packing and other creative uses of the payload continue to
+work---provided the code does not use some other Julia operation
+that does not preserve \NaN{}s bit-for-bit.\footnote{\url{https://github.com/JuliaLang/julia/issues/48523}}
 
@@
 
-In addition to enabling tracking exceptional values, operator
-overloading gives \TF{} a powerful way to fuzz code from the inside
+Julia's operator overloading lets us fuzz code from the inside
 out.
 Each overloaded function serves as a hook where \TF{} can decide whether to observe
-the operation or step in, discard the original result, and return an exception instead.
+the operation or replace the original result with an exceptional value.
 Injecting faults in a random way~\cite{hamlet1994random}, also known as fuzzing,
-is a useful way to discover vulnerabilities in a large codebase.
+can discover vulnerabilities in a large codebase.
 Demmel et~al. propose essentially the same idea for BLAS and
@@
 of the actual computed value.
-This decision is based on several parameters that the developer can
-tune to control fuzzing:
+This decision is based on several parameters:
 
 \begin{itemize}
-\item \texttt{active::Bool} fuzz only when set to true.
-\item \texttt{odds::Int64} inject if \texttt{rand(odds) == 1}.
-\item \texttt{n\_inject::Int64} upper bound on the number of \Nan{}s to inject.
-\item \texttt{functions::Array{String}} limit fuzzing to the dynamic extent of the listed functions.
-\item \texttt{libraries::Array{String}} limit fuzzing to functions from the following libraries
+\item $\,$ \texttt{active::Bool} fuzz only when set to true.
+\item $\,$ \texttt{odds::Int64} inject at each candidate location with probability $1 / \texttt{odds}$.
+\item $\,$ \texttt{n\_inject::Int64} upper bound on the number of \Nan{}s to inject.
+\item $\,$ \texttt{functions::Array{String}} limit fuzzing to the dynamic extent of the listed functions.
+\item $\,$ \texttt{libraries::Array{String}} limit fuzzing to functions from the following libraries
 \end{itemize}
@@
 point where the operation was intercepted to make the decision.
-When fuzzing is active (\texttt{active == true}), there are \Nan{}s
-remaining to be injected (\texttt{n\_inject > 0}), and
-\texttt{rand(odds)} returns \texttt{1}, then \TF{} inspects the
-current stack trace to see if the point is in the dynamic extend of
-the wanted \texttt{functions} or \texttt{libraries}.
-When all the conditions are met, the overloaded operator returns
-\Nan{} instead of whatever value would have been computed normally.
+When fuzzing is active (\texttt{active == true})
+and there are \Nan{}s remaining to be injected (\texttt{n\_inject > 0})
+and \texttt{rand(1:odds)} returns \texttt{1}, then \TF{} inspects the
+current stack trace to see if the point is in the dynamic extent of
+the declared \texttt{functions} or \texttt{libraries} (if any).
+When all these conditions are met,
+\TF{} returns a \NaN{} instead of calling the overloaded operation.
 The value \texttt{n\_inject} is decremented before the next operation
@@
 
-Since \TF{} checks if it should inject a \Nan{} at \emph{every} \fp{}
-operation it intercepts, means that \Nan{}s can be injected deep
-inside of an algorithm instead of just at the inputs.
-This can reveal vulnerabilities that otherwise would have been
-difficult to catch with input-based fuzzers.
-
-Developers can use the \texttt{functions} and \texttt{libraries}
-parameters to narrow down the region where \Nan{}s can be injected.
-It might not be useful, for instance, to fuzz test code that a
-developer has no control over; a developer may not, for example, wish
-to fuzz the standard library and instead focus on the code that they
-are developing.
+Every intercepted operation is a candidate for injection,
+which means that injected \Nan{}s can appear deep
+inside of an algorithm instead of merely at its toplevel inputs.
+This can reveal vulnerabilities that are difficult to catch
+with input-based testing.
+The parameters \texttt{functions} and \texttt{libraries} let
+users control how deep to fuzz; for example, it is often helpful
+to test specific libraries and avoid trusted standard ones.
 
@@
 
-\subsubsection{\TF{} Internals}
+\subsubsection{Internals}
 
@@
 Every arithmetic operation requires a similar overloading.
-Thus, the implementation of \TF{} uses a metaprogramming technique
+The implementation of \TF{} uses a metaprogramming technique
 adapted from \texttt{Sherlogs}~\cite{kMilanklSherlogsJl2021}
@@
 Coalesced stack trace graphs (CSTGs or \CSTG{}s for short) provide a way to visualize large amounts of stack traces in a compact form~\cite{humphreySystematicDebuggingMethods2014}.
-This technique pairs well with \TF{} and makes analyzing the log files easier.
+This technique pairs well with \TF{} and eases the task of analyzing log files.
 
@@
 Since \TF{} instruments Julia programs, it cannot help directly;
-however, the companion tool \GPUFPX{} instruments GPU kernels to detect and report
+however, the companion tool \GPUFPX{} instruments GPU kernels
+(specifically CUDA) to detect and report
 \fp{} exceptions~\cite{llsflg-hpdc-2023}.
-Together, \TF{} and \GPUFPX{} provide insights for accelerated programs.
-
-\TF{} and \GPUFPX{} have no formal connection---that is, one does not
-need to use \GPUFPX{} with \TF{}, nor vice-versa.
-These two tools compliment one another in debugging mixed-paradigm
-workloads, but they can be used in isolation.
-\GPUFPX{} is a library exclusively for NVIDIA CUDA code and has no
-special relation to Julia.
-Nevertheless, in cases where Julia code leverages a CUDA kernel,
-\GPUFPX and \TF{} can both aid in tracking down exceptional \fp{}
-values.
+There is no formal connection between \TF{} and \GPUFPX{}.
+Developers who choose to apply both tools to the respective Julia and CUDA
+components of their code can gain insights for accelerated programs.
 
@@
 The trace further shows that this subtraction happens inside the function \texttt{momentum\_u!} on line 246 of the \texttt{rhs.jl} file.
-
 This solves the mystery of where one \NaN{} came from, but raises a new question
@@
 top-level getproperty(...) examples/sw_nan_tf.jl:14
-..
 \end{lstlisting}
@@
 Calls to `\code{+}' account for most of the \NaN{}s.
-These \NaN{}s arose in two different contexts: a small number (30)
-occurred within a momentum calculation, while the rest (162)
-occurred within a continuity step.
+These \NaN{}s arose in two different contexts:
+a momentum calculation (30~\NaN{}s) and
+a continuity step (162~\NaN{}s).
 Moving to the top of the graph, it shows that the function \code{run\_model} drove
@@
 
-Demmel et al.\cite{ddghlllprr-correctness-2022} examine \fp{} exceptional value handling in BLAS and LAPACK, identify several inconsistencies, and propose an API for debugging and adding determinism.
+Demmel et~al.~\cite{ddghlllprr-correctness-2022} examine \fp{} exceptional value handling in BLAS and LAPACK, identify several inconsistencies, and propose an API for debugging and adding determinism.
 The proposal includes an extension to the \texttt{INFO\_ARRAY} parameter with fields that record \genpropkill{} information.
@@
 This would combine well with \TF{}: first identify the location of a \NaN{}, ask Herbie to find a repair.
-The Odyssey\cite{misbackOdysseyInteractiveWorkbench2023} workbench provides an interactive interface to Herbie.
+The Odyssey~\cite{misbackOdysseyInteractiveWorkbench2023} workbench provides an interactive interface to Herbie.
 

### 2024-08-08 paper.tex
jcon2: number → constant
jcon2: update JuliaHub URL

@@
 and can lead to surprising results.
-For example, adding the tiny Planck constant to the large Avogadro number results in
-the Avogadro number after rounding:
+For example, adding the tiny Planck constant to the large Avogadro constant results in
+the Avogadro constant after rounding:
 
@@
 injecting a \Nan{} or \Inf{} as the result of an operation. \TF{} is
-implemented in Julia as a library and is available on JuliaHub (TODO update link):
+implemented in Julia as a library and is available on JuliaHub:
 
-\noindent\begin{center}\noindent\!\!\url{https://juliahub.com/ui/Packages/FloatTracker/dBXig}\!\!
+\noindent\begin{center}\noindent\!\!\url{https://juliahub.com/ui/Packages/TrackedFloats}\!\!
 \end{center}
